### 2020-10-18 README.Rmd
Initial commit

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "Title of your paper goes here"
```

# rrcompendium

[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh///master?urlpath=rstudio)

This repository contains the data and code for our paper:

> Rodriguez-Machado, S., Rodríguez-Cabrera, T.M., and Chakrabarty, P. (2020). _The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba_. Name of journal/book <https://doi.org/{DOI-to-paper}>

Our pre-print is online here:

> Rodriguez-Machado, S., Rodríguez-Cabrera, T.M., and Chakrabarty, P. (2020). _The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/{DOI-to-compendium}>

### How to cite

Please cite this compendium as:

> Rodriguez-Machado, S., Rodríguez-Cabrera, T.M., and Chakrabarty, P. (2020), (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/{DOI-to-compendium}>

## Contents

The **analysis** directory contains:

  - [:file\_folder: paper](/analysis/paper): R Markdown source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. It also has a rendered version,
    `paper.docx`, suitable for reading (the code is replaced by figures
    and tables in this file)
  - [:file\_folder: data](/analysis/data): Data used in the analysis.
  - [:file\_folder: figures](/analysis/figures): Plots and other
    illustrations
  - [:file\_folder:
    supplementary-materials](/analysis/supplementary-materials):
    Supplementary materials including notes and other documents
    prepared and collected during the analysis.

## How to run in your broswer or download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

The simplest way to explore the text, code and data is to click on
[binder](https://mybinder.org/v2/gh///master?urlpath=rstudio)
to open an instance of RStudio in your browser, which will have the
compendium files ready to work with. Binder uses rocker-project.org
Docker images to ensure a consistent and reproducible computational
environment. These Docker images can also be used locally.

You can download the compendium as a zip from from this URL:
[master.zip](/archive/master.zip). After unzipping:
- open the `.Rproj` file in RStudio
- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
[DESCRIPTION](/DESCRIPTION) file).
- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.Rmd")` in the R console

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### 2020-10-21 analysis/paper/rawPaper.Rmd
seqs downloaded, names ok, mine to include

---
title: "rawPaper"
author: Sheila RM
date: today
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{bash, eval = FALSE}
---

# this is the preamble
```

## Abstract

Keywords:

## Introduction

## Materials and Methods
1- text here

Tetis (19–28 mm standard length) were collected 400 m from the river mouth of Río Yumurí (Figure 1), eastern Cuba, in December, 2018. Samples were collected using hand nets (1 mm mesh size) 2 m from the riverbank. Specimens were anesthetized using MS 222 (Tricaine Methanesulphonate) and immediately preserved in 80% ethanol. Voucher specimens were deposited in the collection of the Acuario Nacional de Cuba (catalog number: ).

2- map

```{r, echo=FALSE}
chunk for generating the map.
```

3- text for dna extraction

Total DNA was extracted from caudal peduncle muscle tissues using the DNAeasy tissue kit (QIAGEN). Partial sequences of the mitochondrial gene cytochrome oxidase I (COI) were amplified through the polymerase chain reaction in 20 μL of reaction volume (one unit of GoTaq DNA polymerase-Promega, 0.2 nM of each primer, 0.2 μM dNTPs, and 1.5 mM MgCl2) following Lara et al. (2010). We used the primers COIf (5´-AAYCAYAAAGAYGGYACCCT-3´) and COIr (3´-CCTCNGGRTGNCCRAAGAAYCA-5´) (Palumbi et al., 1991).

PCR products were purified using ExoSAP-IT™ PCR Product Cleanup Reagent (Affymetrix Inc./ USB, Cleveland, OH, USA), and then sequenced on an ABI 3100 automated sequencer (Applied Biosystems). Raw sequence files were edited and aligned with the software Geneious v.9.1 (Kearse et al., 2012), and submitted to GenBank database (accession numbers #####). We used all Sicydium sequences available in GenBank as reference for recovering the taxonomic identity of the tetis. Sequences with missing data and/or carrying the same haplotype were excluded. Sequences of Awaous banana and A. ocellaris were used as outgroup. See Appendix 1 for a detailed list of species and GenBank accession numbers.

```{r, echo=FALSE}

# reading the sequences recovered during this paper

SycidiumCuba <- NA
                                                       # DOWNLOADING SEQS FROM GENBANK

# this package provides an R interface NCBI API and allows to search databases like GenBank (http://www.ncbi.nih.gov/genbank/) 
install.packages('rentrez')
library(rentrez)

# checking/setting working directory as raw_data for storing the final fasta file, which are the sequences before aligning them in the next chunk 
getwd()
# setwd is temporal, after fishining this chunk, the wd will be again /analysis/paper
setwd("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data")

Sicydium_COI <- "Sicydium[Organism] AND COI[Gene]"

# I set retmax=580 because after running this by default, I got a message with the number of hits (580 in this case)
Sicydium_COI_search <- entrez_search(db = "nuccore", term = Sicydium_COI, retmax = 580)

# given that the search is too big for the system, it needs be split in two smaller one
Sicydium_COI_seqs1 <- entrez_fetch(db = "nuccore", id = Sicydium_COI_search$ids[0:295], rettype = "fasta")
Sicydium_COI_seqs2 <- entrez_fetch(db = "nuccore", id = Sicydium_COI_search$ids[296:580], rettype = "fasta")

# GETTING OUTGROUP SEQUENCES OF AWAOUS BANANA (2) AND A. OCELLARIS (2)

Awaous_bananaCOI <- "Awaous_banana[Organism] AND COI[Gene]"
Awaous_bananaCOI_search <- entrez_search(db = "nuccore", term = Awaous_bananaCOI, retmax = 2)
Awaous_bananaCOI_seqs <- entrez_fetch(db = "nuccore", id = Awaous_bananaCOI_search$ids, rettype = "fasta")

Awaous_ocellarisCOI <- "Awaous_ocellaris[Organism] AND COI[Gene]"
Awaous_ocellarisCOI_search <- entrez_search(db = "nuccore", term = Awaous_ocellarisCOI, retmax = 2)
Awaous_ocellarisCOI_seqs <- entrez_fetch(db = "nuccore", id = Awaous_ocellarisCOI_search$ids, rettype = "fasta")

# storing ALL sequences in a single fasta file
write(Sicydium_COI_seqs1, "Sicydium_COI_seqs.fasta", sep = "\n")
write(Sicydium_COI_seqs2, "Sicydium_COI_seqs.fasta", sep = "\n", append = TRUE)   # by setting 'append = TRUE', I can keep adding seqs
write(Awaous_bananaCOI_seqs, "Sicydium_COI_seqs.fasta", sep = "\n", append = TRUE)
write(Awaous_ocellarisCOI_seqs, "Sicydium_COI_seqs.fasta", sep = "\n", append = TRUE)

                                                   # RENAMING SEQS AND SAVING THEM IN A NEW FASTA FILE

# installing necessary package for sequence manipulation
install.packages('seqinr')
library(seqinr)

# reading the file and renaming the sequences
Sicydium_COI_seqinrFormat <- read.fasta(file = "Sicydium_COI_seqs.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)

Sicydium_COI_names <- attr(Sicydium_COI_seqinrFormat, "name")      # editting species name
Sicydium_COI_names <- gsub("\\..*", "", Sicydium_COI_names)        # eliminating characters after "." using gsub
Sicydium_COI_names <- gsub("^.*\\|", "", Sicydium_COI_names)       # eliminating characters before "|" using gsub

# installing necessary package
install.packages("ape")
library(ape)

Sicydium_COI_apeFormat <- read.GenBank(Sicydium_COI_names)         # this line takes a few minutes to run

#attr(Sicydium_COI_apeFormat, "species")
#names(Sicydium_COI_apeFormat)

# creating a vector (to be added as name of the seqs) with the species, GenBank accession number, and type of gene
Sicydium_COI_GenBank_Ids <- paste(attr(Sicydium_COI_apeFormat, "species"), names(Sicydium_COI_apeFormat), sep = "_COI_")

# writing a new fasta file using the vector above 
write.fasta(sequences = Sicydium_COI_seqinrFormat, names = Sicydium_COI_GenBank_Ids, nbchar = 10, file.out = "Sicydium_COI_seqinrFormat.fasta")

                                            # ALIGNNIG/EDITTING SEQS AND GETTING THEM ALL READY FOR PHYLOGENETICS

# removing all sequences without a complete scientific name, that is, those with 'sp.'
fastafile <- read.fasta(file = "Sicydium_COI_seqinrFormat.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)

getNames <- strsplit(names(fastafile), '_')         # creating a vector splitting species names
removeSeqs <- c("sp.")                              # finding names to remove: indexes which are in the species to be removed
keepSeqs <- which(! getNames %like% removeSeqs)     # selection those without specific name in them

# saving a new fasta file with only those seqs with full scientific names
write.fasta(sequences=fastafile[keepSeqs], names=names(fastafile)[keepSeqs], file.out="Sicydium_COI_edits.fasta")

```

9- text for phylogenetic analyses

The partition scheme and nucleotide substitution model that best fitted the data was estimated using PartitionFinder 2 (Lanfear et al., 2016). The phylogenetic tree was reconstructed using the Bayesian approach using MrBayes 3.2.7a (Ronquist et al., 2012) as implemented in CIPRES (Miller et al., 2010). The Bayesian inference was based on two independent runs using four Monte Carlo Markov chains for 20 x 106 generations, with sampling every 10,000 generations. The first 25% of the sampled trees were discarded as burn-in.

```{r, echo=FALSE}

#getting all packages and libraries
#install.packages("modelTest")
#install.packages("ape")
#library(ape)

# reading the file again, only if necessary
var1 <- read.dna(sequences already alignned, format = 'fasta')    # format = can be sequential, interleaved, clustal
                                                                  # as.character = TRUE/FALSE, default is class DNAbin

# the alternatie is var1 = read.FASTA = (sequences, type = 'DNA')

#getting the class used by modelTest: phyDat or pml
var2 <- phyDat(var1, type = "DNA", levels = NULL)

# getting the best model
var3 <- modelTest(var2, tree = NULL, model = "all", multicore = TRUE, mc.cores = 2)

# getting the AICc for each competing model
# visually inspecting the AICc is better than just getting the best model automatically, because it allows to compares models with similar AICc and it is my personal opinion to choose the one with the least parameters.  
print(var3)

# storing the best model
var4 <- dist.ml(var3, model = "uno d esos")

# the alternative is to get automatically the first model from the data.frame modelTest gives and storing it as a variable...

```

10- Results

```{r, echo=FALSE}

chunk for building the tree and store it as a figure

```

11- Discussion

12- Acknowledgments

Habitat

stri

13- References

### 2020-10-21 analysis/paper/rawPaper.Rmd
var names changed-rest as before

@@
 
-getNames <- strsplit(names(fastafile), '_')         # creating a vector splitting species names
-removeSeqs <- c("sp.")                              # finding names to remove: indexes which are in the species to be removed
-keepSeqs <- which(! getNames %like% removeSeqs)     # selection those without specific name in them
+fastaNames <- strsplit(names(fastafile), '_')                  # creating a vector splitting species names
+incompleteNames <- c("sp.")                                    # finding names to remove: indexes which are in the species to be removed
+completeSeqs <- which(! fastaNames %like% incompleteNames)     # selection those without specific name in them
 
 # saving a new fasta file with only those seqs with full scientific names
-write.fasta(sequences=fastafile[keepSeqs], names=names(fastafile)[keepSeqs], file.out="Sicydium_COI_edits.fasta")
+write.fasta(sequences=fastafile[completeSeqs], names=names(fastafile)[completeSeqs], file.out="Sicydium_COI_edits.fasta")
 

### 2020-10-23 analysis/paper/rawPaper.Rmd
names & length set --- pending alignment

--- 2020-10-21 analysis/paper/rawPaper.Rmd
@@
 ```{r, echo=FALSE}
-chunk for generating the map.
+#chunk for generating the MAP.
+
+install.packages("maps")
+install.packages("mapdata")
+
+library(maps)
+library(mapdata)
+map("worldHires","Cuba", xlim=c(-88,-73), ylim=c(19,24), col="gray95", fill=TRUE)
+
 ```
@@
 # checking/setting working directory as raw_data for storing the final fasta file, which are the sequences before aligning them in the next chunk 
-getwd()
+#getwd()
 # setwd is temporal, after fishining this chunk, the wd will be again /analysis/paper
-setwd("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data")
+#setwd("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data")
+#getwd()
 
@@
 # storing ALL sequences in a single fasta file
-write(Sicydium_COI_seqs1, "Sicydium_COI_seqs.fasta", sep = "\n")
-write(Sicydium_COI_seqs2, "Sicydium_COI_seqs.fasta", sep = "\n", append = TRUE)   # by setting 'append = TRUE', I can keep adding seqs
-write(Awaous_bananaCOI_seqs, "Sicydium_COI_seqs.fasta", sep = "\n", append = TRUE)
-write(Awaous_ocellarisCOI_seqs, "Sicydium_COI_seqs.fasta", sep = "\n", append = TRUE)
+write(Sicydium_COI_seqs1, file="../data/raw_data/Sicydium_COI_seqs.fasta", sep = "\n")
+write(Sicydium_COI_seqs2, file="../data/raw_data/Sicydium_COI_seqs.fasta", sep = "\n", append = TRUE)   # 'append = TRUE', I can keep adding seqs
+write(Awaous_bananaCOI_seqs, file="../data/raw_data/Sicydium_COI_seqs.fasta", sep = "\n", append = TRUE)
+write(Awaous_ocellarisCOI_seqs, file="../data/raw_data/Sicydium_COI_seqs.fasta", sep = "\n", append = TRUE)
 
@@
 # reading the file and renaming the sequences
-Sicydium_COI_seqinrFormat <- read.fasta(file = "Sicydium_COI_seqs.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
+Sicydium_COI_seqinrFormat <- read.fasta(file ="../data/raw_data/Sicydium_COI_seqs.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
 
@@
 
-# writing a new fasta file using the vector above 
-write.fasta(sequences = Sicydium_COI_seqinrFormat, names = Sicydium_COI_GenBank_Ids, nbchar = 10, file.out = "Sicydium_COI_seqinrFormat.fasta")
+# writing a new fasta file (2nd) using the vector above 
+write.fasta(sequences = Sicydium_COI_seqinrFormat, names = Sicydium_COI_GenBank_Ids, nbchar = 10, file.out = "../data/raw_data/Sicydium_COI_seqinrFormat")
 
@@
 
-# removing all sequences without a complete scientific name, that is, those with 'sp.'
-fastafile <- read.fasta(file = "Sicydium_COI_seqinrFormat.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
+### Removing sequences of incomplete/unspecific NAMES; that is, those with 'sp.'
+nameFasta <- read.fasta(file = "../data/raw_data/Sicydium_COI_seqinrFormat", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
 
-getNames <- strsplit(names(fastafile), '_')         # creating a vector splitting species names
-removeSeqs <- c("sp.")                              # finding names to remove: indexes which are in the species to be removed
-keepSeqs <- which(! getNames %like% removeSeqs)     # selection those without specific name in them
+fastaNames <- strsplit(names(nameFasta), '_')                  # creating a vector splitting species names
+completeSeqs <- which(! grepl("sp.", fastaNames))              # selecting those with specific name in them       # 562
 
-# saving a new fasta file with only those seqs with full scientific names
-write.fasta(sequences=fastafile[keepSeqs], names=names(fastafile)[keepSeqs], file.out="Sicydium_COI_edits.fasta")
+# writing a new fasta (3rd) file with only those seqs with full scientific names
+write.fasta(sequences = nameFasta[completeSeqs], names = names(nameFasta)[completeSeqs], file.out = "../data/raw_data/Sicydium_COI_names.fasta")
+
+### Removing sequences of LENGTH shorter than 600 pb, given that the gene COI has between 627 and 658 bp.
+lengthFasta <- read.fasta(file = "../data/raw_data/Sicydium_COI_names.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
+fastaLength <- lengthFasta[which(getLength(lengthFasta) >= 600)]       # 399
+write.fasta(sequences = fastaLength, names = names(fastaLength), file.out = "../data/raw_data/Sicydium_COI_lengths.fasta")
+
+### Aligning the sequences
+
+BiocManager::install("msa")
+library(msa)
+
+alignmentFasta <- read.fasta(file = "../data/raw_data/Sicydium_COI_lengths.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
+
+length(alignmentFasta)
+
+myClustalOmegaAlignment <- msa("alignmentFasta", type = "DNA", "ClustalOmega")
+
+print(myClustalOmegaAlignment, showConsensus=FALSE, halfNrow=3)
+print(myClustalOmegaAlignment, showNames=FALSE, show="complete")
+
+### Removing most duplicated sequences
+
+### Removing the segment before the start codon TAC
 

### 2020-10-27 analysis/paper/rawPaper.Rmd
until seqs aligned --next duplicates and TAC

@@
 
-# reading the sequences recovered during this paper
+                                                      # INSTALLING PACKAGES AND LIBRARIES - if needed
+packages = c("rentrez", "ape", "seqinr", "BiocManager", "msa")
+package.check <- lapply(
+  packages,
+  FUN = function(x) {
+    if (!require(x, character.only = TRUE)) {
+      install.packages(x, dependencies = TRUE)
+      library(x, character.only = TRUE)
+    }
+  }
+)
 
-SycidiumCuba <- NA
+#if (!requireNamespace("BiocManager", quietly=TRUE))
+#install.packages("BiocManager")
+#BiocManager::install("msa")
+
                                                        # DOWNLOADING SEQS FROM GENBANK
 
-# this package provides an R interface NCBI API and allows to search databases like GenBank (http://www.ncbi.nih.gov/genbank/) 
-install.packages('rentrez')
-library(rentrez)
-
-# checking/setting working directory as raw_data for storing the final fasta file, which are the sequences before aligning them in the next chunk 
-#getwd()
-# setwd is temporal, after fishining this chunk, the wd will be again /analysis/paper
-#setwd("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data")
-#getwd()
+# rentrez: this package provides an R interface NCBI API and allows to search databases like GenBank (http://www.ncbi.nih.gov/genbank/) 
+library(rentrez)     # delete this line after finishing
 
@@
 Awaous_ocellarisCOI <- "Awaous_ocellaris[Organism] AND COI[Gene]"
-Awaous_ocellarisCOI_search <- entrez_search(db = "nuccore", term = Awaous_ocellarisCOI, retmax = 2)
+Awaous_ocellarisCOI_search <- entrez_search(db = "nuccore", term = Awaous_ocellarisCOI, retmax = 4)
 Awaous_ocellarisCOI_seqs <- entrez_fetch(db = "nuccore", id = Awaous_ocellarisCOI_search$ids, rettype = "fasta")
@@
 
-# installing necessary package for sequence manipulation
-install.packages('seqinr')
-library(seqinr)
+# seqinr: package for sequence manipulation
+
+library(seqinr)     # delete this line after finishing
 
@@
 
-Sicydium_COI_names <- attr(Sicydium_COI_seqinrFormat, "name")      # editting species name
+Sicydium_COI_names <- attr(Sicydium_COI_seqinrFormat, "name")      # editing species name
 Sicydium_COI_names <- gsub("\\..*", "", Sicydium_COI_names)        # eliminating characters after "." using gsub
@@
 
-# installing necessary package
-install.packages("ape")
-library(ape)
+library(ape)   # delete this line after finishing
 
@@
 
-                                            # ALIGNNIG/EDITTING SEQS AND GETTING THEM ALL READY FOR PHYLOGENETICS
+                                     # ALIGNIG/EDITTING SEQS AND GETTING THEM ALL READY FOR THE PHYLOGENETIC ANALYSES
 
@@
 fastaNames <- strsplit(names(nameFasta), '_')                  # creating a vector splitting species names
-completeSeqs <- which(! grepl("sp.", fastaNames))              # selecting those with specific name in them       # 562
+completeSeqs <- which(! grepl("sp.", fastaNames))              # selecting only those with specific name in them       # 562
 
@@
 ### Aligning the sequences
+# using libraries "msa" and "Biostrings"
 
-BiocManager::install("msa")
-library(msa)
+library(msa)           # delete this line after finishing
+library(Biostrings)    # delete this line after finishing
 
-alignmentFasta <- read.fasta(file = "../data/raw_data/Sicydium_COI_lengths.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
+mySequences <- readDNAStringSet("../data/raw_data/Sicydium_COI_lengths.fasta")
 
-length(alignmentFasta)
+myClustalOmegaAlignment <- msa(mySequences, "ClustalOmega")   # using algorithm Clustal Omega (Sievers et al. 2011) for aligning
+#print(myClustalOmegaAlignment, showConsensus=FALSE, showNames=TRUE, show="complete")
 
-myClustalOmegaAlignment <- msa("alignmentFasta", type = "DNA", "ClustalOmega")
+# removing samples with to many gaps at the beginning and the end of the sequences 
+colmask(myClustalOmegaAlignment) <- IRanges(start=c(1,666),end=c(13,679))                
+myClustalOmegaAlignment2 <- as(myClustalOmegaAlignment, "DNAStringSet")
+writeXStringSet(myClustalOmegaAlignment2, file="../data/raw_data/Sicydium_COI_alignment.fasta")         # fasta file of 652 bp
 
-print(myClustalOmegaAlignment, showConsensus=FALSE, halfNrow=3)
-print(myClustalOmegaAlignment, showNames=FALSE, show="complete")
-
-### Removing most duplicated sequences
+### Removing most duplicate sequences
 

### 2020-10-30 analysis/paper/rawPaper.Rmd
modelTestig here we go

@@
                                                       # INSTALLING PACKAGES AND LIBRARIES - if needed
-packages = c("rentrez", "ape", "seqinr", "BiocManager", "msa")
+packages = c("rentrez", "ape", "seqinr", "BiocManager", "modelTest")
 package.check <- lapply(
@@
 
-#if (!requireNamespace("BiocManager", quietly=TRUE))
-#install.packages("BiocManager")
-#BiocManager::install("msa")
+if (!requireNamespace("BiocManager", quietly = TRUE))
+    install.packages("BiocManager")
 
+BiocManager::install("msa")
+BiocManager::install("systemPipeR")
+library(stringr)
+library(systemPipeR)
                                                        # DOWNLOADING SEQS FROM GENBANK
@@
 
+# using library "ape"
 library(ape)   # delete this line after finishing
@@
 
-                                     # ALIGNIG/EDITTING SEQS AND GETTING THEM ALL READY FOR THE PHYLOGENETIC ANALYSES
+                                     # ALIGNIG/EDITTING SEQS AND GETTING THEM ALL READY FOR THE PHYLOGENETIC ANALYSIS
 
-### Removing sequences of incomplete/unspecific NAMES; that is, those with 'sp.'
+### REMOVING sequences of incomplete/unspecific NAMES; that is, those with 'sp.'
 nameFasta <- read.fasta(file = "../data/raw_data/Sicydium_COI_seqinrFormat", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
@@
 fastaNames <- strsplit(names(nameFasta), '_')                  # creating a vector splitting species names
-completeSeqs <- which(! grepl("sp.", fastaNames))              # selecting only those with specific name in them       # 562
+completeSeqs <- which(! grepl("sp.", fastaNames))              # selecting only those with specific name in them       # 562 seqs
 
@@
 
-### Removing sequences of LENGTH shorter than 600 pb, given that the gene COI has between 627 and 658 bp.
+### REMOVING sequences of LENGTH shorter than 600 pb, given that the gene COI has between 627 and 658 bp.
 lengthFasta <- read.fasta(file = "../data/raw_data/Sicydium_COI_names.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
-fastaLength <- lengthFasta[which(getLength(lengthFasta) >= 600)]       # 399
-write.fasta(sequences = fastaLength, names = names(fastaLength), file.out = "../data/raw_data/Sicydium_COI_lengths.fasta")
+fastaLength <- lengthFasta[which(getLength(lengthFasta) >= 600)]       
+write.fasta(sequences = fastaLength, names = names(fastaLength), file.out = "../data/raw_data/Sicydium_COI_lengths.fasta")      # 399 seqs
 
-### Aligning the sequences
-# using libraries "msa" and "Biostrings"
-
+### Aligning the sequences    # using libraries "msa" and "Biostrings"
 library(msa)           # delete this line after finishing
@@
 
-mySequences <- readDNAStringSet("../data/raw_data/Sicydium_COI_lengths.fasta")
+alignmentFasta <- readDNAStringSet("../data/raw_data/Sicydium_COI_lengths.fasta")
+ClustalOAlignment <- msa(alignmentFasta, "ClustalOmega")   # algorithm Clustal Omega (Sievers et al. 2011) used for aligning
+#print(ClustalOAlignment, showConsensus=FALSE, showNames=TRUE, show="complete")
 
-myClustalOmegaAlignment <- msa(mySequences, "ClustalOmega")   # using algorithm Clustal Omega (Sievers et al. 2011) for aligning
-#print(myClustalOmegaAlignment, showConsensus=FALSE, showNames=TRUE, show="complete")
+# REMOVING samples with to many GAPS at the beginning and the end of the sequences 
+colmask(ClustalOAlignment) <- IRanges(start=c(1,666),end=c(13,679))                
+ClustalOAlignment2 <- as(ClustalOAlignment, "DNAStringSet")
+writeXStringSet(ClustalOAlignment2, file="../data/raw_data/Sicydium_COI_alignment.fasta")         # 399 seqs of 652 bp.
 
-# removing samples with to many gaps at the beginning and the end of the sequences 
-colmask(myClustalOmegaAlignment) <- IRanges(start=c(1,666),end=c(13,679))                
-myClustalOmegaAlignment2 <- as(myClustalOmegaAlignment, "DNAStringSet")
-writeXStringSet(myClustalOmegaAlignment2, file="../data/raw_data/Sicydium_COI_alignment.fasta")         # fasta file of 652 bp
+### Removing DUPLICATE sequences    # using "seqinr" to do it
 
-### Removing most duplicate sequences
+# for not using attributes to name sequences
+duplicatesFasta <- read.fasta("../data/raw_data/Sicydium_COI_alignment.fasta", set.attributes = FALSE) 
+seqnames <- names(duplicatesFasta)
+dups <- grep(TRUE,duplicated(duplicatesFasta))
+uniqueNames <- seqnames[-dups]
+uniqueSeqs <- duplicatesFasta[-dups] 
+write.fasta(sequences = uniqueSeqs, names = uniqueNames, file.out = "../data/raw_data/Sicydium_COI_dupli.fasta")      # 197 seqs of 652 bp.
 
-### Removing the segment before the start codon TAC
+### Replacing gaps with Ns
+library(stringr)
+
+gapsFasta <- read.fasta(file = "../data/raw_data/Sicydium_COI_alignment.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
+replacingGaps <- lapply(gapsFasta, function(x) str_replace_all(x, "-", "N"))
+write.fasta(replacingGaps,names(replacingGaps),"../data/raw_data/Sicydium_COI_gaps.fasta")
+
+# GETTING the READING FRAME
+BiocManager::install("systemPipeR")    # delete this line after finishing
+library(systemPipeR)
+
+gettingRF <- function(x){
+  fasta <- readDNAStringSet("../data/raw_data/Sicydium_COI_gaps.fasta")
+  orf1 <- predORF(fasta[1:10], n=1, type="df", mode="orf", strand="sense")   # predORF() do not read seqs with gaps
+  # predicts longest ORF for sense strand in each query sequence                                                                          
+  longestSeq1 <- orf$start[orf$width == max(orf$width)]
+  codonPosition1 <- longestSeq[1]/3
+  codonConfirmation1 <- round(codonPosition1)
+  if (codonPosition1==codonConfirmation1){
+    print("The Reading Frame of the seqs is +1, no further action is required, go to modelTesting")
+    } else {
+    if (codonPosition1!=codonConfirmation1){
+      #remove first nucleotide position
+      newRF2 <- subseq(fasta, start = 2, end = tempLength)      # 651 bp
+
+      # getting new RF from the new alignment with -1 bp.
+      orf2 <- predORF(newRF2[1:10], n=1, type="df", mode="orf", strand="sense")
+
+      # checking size gain
+      longestSeq2 <- orf2$start[orf2$width == max(orf2$width)]
+      codonPosition2 <- (longestSeq2[1]+2)/3
+      codonConfirmation2 <- round(codonPosition2)
+      if (codonPosition2==codonConfirmation2){
+        writeXStringSet(newRF2, file="../data/raw_data/Sicydium_COI_newRF2.fasta")
+      print("The Reading Frame of the seqs is +2, new fasta file created, go to modelTesting")
+      } else {
+      if (codonPosition2!=codonConfirmation2){
+        #remove first nucleotide position
+        tempLength <- length(fasta[[1]])
+        newRF3 <- subseq(fasta, start = 3, end = tempLength)
+        orf3 <- predORF(newRF3[1:10], n=1, type="df", mode="orf", strand="sense") 
+        # checking size gain
+        longestSeq3 <- orf3$start[orf3$width == max(orf3$width)]
+        codonPosition3 <- (longestSeq3[1]+2)/3
+        codonConfirmation3 <- round(codonPosition3)
+        if (codonPosition3==codonConfirmation3){
+          writeXStringSet(newRF3, file="../data/raw_data/Sicydium_COI_newRF3.fasta")
+        print("The Reading Frame of the seqs is +3, new fasta file created, go to modelTesting")
+       }
+     }
+   }  
+    }
+  }
+}
+
+gettingRF(x)
 
@@
 
-# the alternatie is var1 = read.FASTA = (sequences, type = 'DNA')
+# the alternative is var1 = read.FASTA = (sequences, type = 'DNA')
 

### 2020-11-02 analysis/paper/rawPaper.Rmd
ready to modelTesting-OK

@@
 
-gapsFasta <- read.fasta(file = "../data/raw_data/Sicydium_COI_alignment.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
+gapsFasta <- read.fasta(file = "../data/raw_data/Sicydium_COI_dupli.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
 replacingGaps <- lapply(gapsFasta, function(x) str_replace_all(x, "-", "N"))
@@
   # predicts longest ORF for sense strand in each query sequence                                                                          
-  longestSeq1 <- orf$start[orf$width == max(orf$width)]
-  codonPosition1 <- longestSeq[1]/3
+  longestSeq1 <- orf1$start[orf1$width == max(orf1$width)]
+  codonPosition1 <- longestSeq1[1]/3
   codonConfirmation1 <- round(codonPosition1)
   if (codonPosition1==codonConfirmation1){
-    print("The Reading Frame of the seqs is +1, no further action is required, go to modelTesting")
+   file.rename("../data/raw_data/Sicydium_COI_gaps.fasta", "../data/raw_data/Sicydium_COI_newRF.fasta") 
+  print("The Reading Frame of these seqs is +1, no further action is required, go to modelTesting")
     } else {
     if (codonPosition1!=codonConfirmation1){
+      tempLength <- length(fasta[[1]])
       #remove first nucleotide position
       newRF2 <- subseq(fasta, start = 2, end = tempLength)      # 651 bp
-
       # getting new RF from the new alignment with -1 bp.
       orf2 <- predORF(newRF2[1:10], n=1, type="df", mode="orf", strand="sense") 
-
       # checking size gain
@@
       if (codonPosition2==codonConfirmation2){
-        writeXStringSet(newRF2, file="../data/raw_data/Sicydium_COI_newRF2.fasta")
-      print("The Reading Frame of the seqs is +2, new fasta file created, go to modelTesting")
+        writeXStringSet(newRF2, file="../data/raw_data/Sicydium_COI_newRF.fasta")
+      print("The Reading Frame of these seqs was +2, new fasta file created, go to modelTesting")
       } else {
@@
         newRF3 <- subseq(fasta, start = 3, end = tempLength)
-        orf3 <- predORF(newRF3[1:10], n=1, type="df", mode="orf", strand="sense") 
-        # checking size gain
-        longestSeq3 <- orf3$start[orf3$width == max(orf3$width)]
-        codonPosition3 <- (longestSeq3[1]+2)/3
-        codonConfirmation3 <- round(codonPosition3)
-        if (codonPosition3==codonConfirmation3){
-          writeXStringSet(newRF3, file="../data/raw_data/Sicydium_COI_newRF3.fasta")
-        print("The Reading Frame of the seqs is +3, new fasta file created, go to modelTesting")
-       }
+        writeXStringSet(newRF3, file="../data/raw_data/Sicydium_COI_newRF.fasta")
+      print("The Reading Frame of these seqs was +3, new fasta file created, go to modelTesting")
      }
@@
 #getting all packages and libraries
-#install.packages("modelTest")
-#install.packages("ape")
-#library(ape)
+BiocManager::install("modelTest")
+install.packages("ape")
+#library(modelTest)
 
-# reading the file again, only if necessary
-var1 <- read.dna(sequences already alignned, format = 'fasta')    # format = can be sequential, interleaved, clustal
+install.packages("phangorn")
+library(phangorn)
+
+gettingBestModel <- function(x){
+  var1 <- read.dna(file="../data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
+  var2 <- as.phyDat(var1)
+  var3 <- modelTest(var2, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 2)
+  return(var3)
+}
+
+gettingBestModel(x)
+
+mammals <- read.dna("mammals.dna", format="interleaved")
+mammals_phyDat <- phyDat(mammals, type = "DNA", levels = NULL)
+mammals10 <- subset(mammals_phyDat, 1:10)
+mammals10_phyDat <- phyDat(mammals10, type = "DNA", levels = NULL)
+
+var1 <- read.dna(file="../data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
+var2 <- as.phyDat(var1)
+
+var11 <- subset(var2, 1:10)
+var21 <- phyDat(var11, type = "DNA", levels = NULL)
+
+var3 <- modelTest(var21, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 2)
+
+print(var3)
+
+dna_dist <- dist.ml(var11, model="JC69")
+mammals_NJ  <- NJ(dna_dist)
+plot(mammals_NJ, main = "Neighbor Joining")
+
+fit <- pml(mammals_NJ, var11)
+print(fit)
+
+fitJC <- optim.pml(fit, model = "JC", rearrangement = "stochastic")
+bs <- bootstrap.pml(fitJC, bs=100, optNni=TRUE, multicore=TRUE, control = pml.control(trace=0))
+plotBS(midpoint(fitJC$tree), bs, p = 50, type="p")
+write.tree(bs, file="bootstrap_example.tre")
+
+# format = can be sequential, interleaved, clustal
 # as.character = TRUE/FALSE, default is class DNAbin
@@
 #getting the class used by modelTest: phyDat or pml
-var2 <- phyDat(var1, type = "DNA", levels = NULL)
-
 # getting the best model
-var3 <- modelTest(var2, tree = NULL, model = "all", multicore = TRUE, mc.cores = 2)
-
 # getting the AICc for each competing model
 # visually inspecting the AICc is better than just getting the best model automatically, because it allows to compares models with similar AICc and it is my personal opinion to choose the one with the least parameters.
-print(var3)
-
-# storing the best model
-var4 <- dist.ml(var3, model = "uno d esos")
 

### 2020-11-05 analysis/paper/rawPaper.Rmd
tree almost done

@@
 
-#getting all packages and libraries
+                                          # FINDING THE BEST NUCLEOTIDE SUSBSTITUTION MODEL
 BiocManager::install("modelTest")
 install.packages("ape")
-#library(modelTest)
 
@@
 
-gettingBestModel <- function(x){
-  var1 <- read.dna(file="../data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
-  var2 <- as.phyDat(var1)
-  var3 <- modelTest(var2, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 2)
-  return(var3)
-}
+#gettingBestModel <- function(x){
+  modelFasta <- read.dna(file="../data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
+  formatPhyDat <- as.phyDat(modelFasta)                #getting the class used by modelTest: phyDat or pml
+  modelChoice <- modelTest(formatPhyDat, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 4)
+  modelChoice[order(modelChoice$AICc),]   # getting the AICc for each competing model
+  bestmodel <- modelChoice$Model[which.min(modelChoice$AICc)]
+  #return(bestmodel)
+#}
 
-gettingBestModel(x)
-
-mammals <- read.dna("mammals.dna", format="interleaved")
-mammals_phyDat <- phyDat(mammals, type = "DNA", levels = NULL)
-mammals10 <- subset(mammals_phyDat, 1:10)
-mammals10_phyDat <- phyDat(mammals10, type = "DNA", levels = NULL)
-
-var1 <- read.dna(file="../data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
-var2 <- as.phyDat(var1)
-
-var11 <- subset(var2, 1:10)
-var21 <- phyDat(var11, type = "DNA", levels = NULL)
-
-var3 <- modelTest(var21, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 2)
-
-print(var3)
-
-dna_dist <- dist.ml(var11, model="JC69")
-mammals_NJ  <- NJ(dna_dist)
-plot(mammals_NJ, main = "Neighbor Joining")
-
-fit <- pml(mammals_NJ, var11)
-print(fit)
-
-fitJC <- optim.pml(fit, model = "JC", rearrangement = "stochastic")
-bs <- bootstrap.pml(fitJC, bs=100, optNni=TRUE, multicore=TRUE, control = pml.control(trace=0))
-plotBS(midpoint(fitJC$tree), bs, p = 50, type="p")
-write.tree(bs, file="bootstrap_example.tre")
-
-# format = can be sequential, interleaved, clustal
-# as.character = TRUE/FALSE, default is class DNAbin
-
-# the alternative is var1 = read.FASTA = (sequences, type = 'DNA')
-
-#getting the class used by modelTest: phyDat or pml
-# getting the best model
-# getting the AICc for each competing model
 # visually inspecting the AICc is better than just getting the best model automatically, because it allows to compares models with similar AICc and it is my personal opinion to choose the one with the least parameters.
@@
 # the alternative is to get automatically the first model from the data.frame modelTest gives and storing it as a variable...
+#gettingBestModel(x)        # HKY+G+I
 
@@
 
-chunk for building the tree and store it as a figure
+                                        # FINDING THE BEST NUCLEOTIDE SUSBSTITUTION MODEL
+# getting the tree
+
+library(phangorn)
+env = attr(modelChoice, "env")
+fitStart = eval(get(bestmodel, env), env)    # initial fit for the model chosen
+
+fit = optim.pml(fitStart, rearrangement = "stochastic", optGamma=TRUE, optInv=TRUE, model="HKY")     # seeking a maximum likelihood solution
+bs = bootstrap.pml(fit, bs=100, optNni=TRUE, multicore=TRUE)
+
+# verifying that the optimized tree is indeed better than the original one using standard likelihood ratio tests and AIC
+compareTrees <- anova(fitStart, fit)          # AICc(fitStart)       # AICc(fit)   
+# Both the ANOVA test (highly significant) and the AIC (lower=better) indicate that the new tree is a better model of the data than the initial one.
+
+# saving the tree in folder analysis/figures/
+bsTree <- plotBS(midpoint(fit$tree), bs, p = 75, type="p", bs.col = "blue")     # fit$tree is where the optimal tree is stored
+plotBS(bsTree, type = "p", cex = 0.2, font = 1)
+write.tree(bsTree, file="../figures/Sicydium_ML_tree.tre")                      # export tree with bootstrap values as node labels
+
+# add.scale.bar(cex = 0.7, font = 2, col = "red")
+
+#plot(bsTree, type = "phylogram", use.edge.length = FALSE, show.node.label = TRUE, label.offset = 10,  node.depth = 1)
+
+BiocManager::install("ggtree")
+library(ggtree)
+
+tree <- read.tree(file="../figures/Sicydium_ML_tree.tre")
+
+# getting the node number for each main clade
+nodeNumber <- NA
+
+ggtree(tree) + geom_text(aes(label=node), hjust=-.5, col = "blue") # looks horrible but gives me the clade number
+
+# editing/visualizing the tree
+tree2 <-ggtree(tree, size = .4) + geom_treescale(x=0, y=-1.5, offset = -4) + geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 70), size = 3, col= "black", vjust = -0.3, hjust=1.9, na.rm = FALSE) + geom_cladelabel(node=206, label="Sicydium plumieri", color="red") + geom_cladelabel(node=113, label="Sicydium brevifile", color="blue") + geom_cladelabel(node=307, label="Sicydium salvini", color="purple") + geom_cladelabel(node=314, label="Sicydium punctatum", color="brown") + geom_cladelabel(node=363, label="Sicydium buscki", color="black") + geom_cladelabel(node=312, label="Sicydium crenilabrum", color="pink") + geom_hilight(node=369, fill="green") + theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))
+
+tree2
++ ggplot2::xlim(0, 0.06) + xlim(0, 1) 
+geom_nodelab(size = 7, col= "red") + theme_tree2() + ylim(0,15)
+
+ggtree(tree) + geom_text2(aes(subset = !isTip, label=label))
+# ggtree(tree) + geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 80))
+plot(bsTree, no.margin = TRUE, cex = 0.5)
+add.scale.bar(bsTree, lwd = 1, cex = 0.7, font = 2, col = "red")
+
+# saving the tree as a png figure
+ggsave(tree2, width = 15, height = 23, units = "cm", limitsize = FALSE, file = "../figures/Sicydium_ML_tree1.png")
 

### 2020-11-05 analysis/paper/rawPaper.Rmd
tree almost done1

@@
 
-# add.scale.bar(cex = 0.7, font = 2, col = "red")
-
-#plot(bsTree, type = "phylogram", use.edge.length = FALSE, show.node.label = TRUE, label.offset = 10,  node.depth = 1)
-
 BiocManager::install("ggtree")
@@
 # editing/visualizing the tree
-tree2 <-ggtree(tree, size = .4) + geom_treescale(x=0, y=-1.5, offset = -4) + geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 70), size = 3, col= "black", vjust = -0.3, hjust=1.9, na.rm = FALSE) + geom_cladelabel(node=206, label="Sicydium plumieri", color="red") + geom_cladelabel(node=113, label="Sicydium brevifile", color="blue") + geom_cladelabel(node=307, label="Sicydium salvini", color="purple") + geom_cladelabel(node=314, label="Sicydium punctatum", color="brown") + geom_cladelabel(node=363, label="Sicydium buscki", color="black") + geom_cladelabel(node=312, label="Sicydium crenilabrum", color="pink") + geom_hilight(node=369, fill="green") + theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))
+tree2 <-ggtree(tree, size = .4) + 
+  geom_treescale(x=0, y=-1.5, offset = -4) + 
+  geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 70), size = 3, col= "black", vjust = -0.3, hjust=1.9, na.rm = FALSE) + 
+  geom_nodepoint(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 70), size = 1, col= "black") + 
+  geom_cladelabel(node=206, label="Sicydium plumieri", color="red", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=113, label="Sicydium brevifile", color="blue", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=307, label="Sicydium salvini", color="purple", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=314, label="Sicydium punctatum", color="brown", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=363, label="Sicydium buscki", color="black", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=312, label="Sicydium crenilabrum", color="orange", hjust = -0.07, fontsize = 3.5, offset = 0.004, fontface="italic") + 
+  geom_hilight(node=369, fill="green") + 
+  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))
 
 tree2
-+ ggplot2::xlim(0, 0.06) + xlim(0, 1) 
-geom_nodelab(size = 7, col= "red") + theme_tree2() + ylim(0,15)
-
-ggtree(tree) + geom_text2(aes(subset = !isTip, label=label))
-# ggtree(tree) + geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 80))
-plot(bsTree, no.margin = TRUE, cex = 0.5)
-add.scale.bar(bsTree, lwd = 1, cex = 0.7, font = 2, col = "red")
 

### 2020-11-22 analysis/paper/paper.Rmd
makefile

---
title: 'The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba'
author:
  - Sheila Rodriguez-Machado:
      email: srodr62@lsu.edu
      institute: [UofO]
      correspondence: true
  - Tomas M. Rodriguez-Cabrera:
      email: tomasmichelrodriguez@gmail.com
      institute: [UofA]
      correspondence: false
institute:
  - UofO: Museum of Natural Sciences, Louisiana State University
  - UofA: Sociedad Cubana de Zoologia
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: '/Users/sheila_ffish/Documents/rrcompendium/analysis/templates/template.docx' # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: '/Users/sheila_ffish/Documents/rrcompendium/analysis/templates/journal-of-archaeological-science.csl' # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  amphidromy; Busck's Stone-Biting Goby; COI; Gobiidae

---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

<!-- The following code chunk defines some general settings how code chunks should behave. -->

```{r setup, echo = FALSE}    
# {r settings, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)

#knitr::opts_knit$set(root.dir = '../rrcompendium/')

```

```{r}

# INSTALLING PACKAGES AND LIBRARIES - if needed
packages = c("raster", "ggplot2", "ggthemes",  
             "rentrez", "ape", "seqinr", "msa", "Biostrings", "systemPipeR", "phangorn", "ggtree")

package.check <- lapply(packages, FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE, warn.conflicts = FALSE, quietly = TRUE)
    }
  }
)

```

<!-- The actual document text starts here: -->

# Introduction

Here is a citation [@Marwick2017]

# Methods

Tetis (19–28 mm standard length) were collected 400 m from the river mouth of Río Yumurí (Figure 2.1), eastern Cuba, in December, 2018. Samples were collected using hand nets (1 mm mesh size) 2 m from the riverbank. Specimens were anesthetized using MS 222 (Tricaine Methanesulphonate) and immediately preserved in 80% ethanol. Voucher specimens were deposited in the collection of the Acuario Nacional de Cuba (catalog number: ANC-xxxx).

```{r figure_1, fig.cap="Sampling site at Yumuri river mouth, Eastern Cuba"}

# getting the map of Cuba
mydata <- getData("GADM", country = "cuba", level = 1)  # package raster
mymap <- fortify(mydata)                                # package ggplot2

# sampling site
dots <- data.frame(
  long = c(-74.293460),
  lat = c(20.298537),
  names = c("Yumuri river mouth"),
  stringsAsFactors = FALSE
  )

# main map
mainMap <-  ggplot() +
       geom_blank(data = mymap, aes(x = long, y = lat)) +
       geom_map(data = mymap, map = mymap, aes(group = group, map_id = id), fill = "gray90", color = "black", size = 0.3) +
       scale_x_continuous(limits = c(-74.51, -74.11), expand = c(0, 0)) +
       scale_y_continuous(limits = c(20.13, 20.42), expand = c(0, 0)) +
       geom_point(data = dots, aes(x = long, y = lat), color = "red", size = 4) +     
       coord_equal() +
       theme_bw()

# small square framing the zomm area
zoomOut <- data.frame(long = c(-74.49, -74.49, -74.09, -74.09, -74.49),
                   lat = c(20.13, 20.40, 20.40, 20.13, 20.13))

# inset map
insetMap <- ggplotGrob(
      ggplot() +
      geom_blank(data = mymap, aes(x = long, y = lat)) +
      geom_map(data = mymap, map = mymap, aes(group = group, map_id = id), fill = "gray95", color = "black", size = 0.3) +
      geom_path(data = zoomOut, aes(x = long, y = lat), size = 0.3) +
      scale_x_continuous(limits = c(-85.47, -73.65), expand = c(0, 0)) +
      scale_y_continuous(limits = c(19.5, 23.5), expand = c(0, 0)) +
      coord_map("polyconic") +
      theme_map() +     # package ggthemes
      theme(panel.background = element_rect(fill = NULL)) # fill = NULL or transparent, with color=NA both the background and the frame are transparent
      )

mainMap +
      annotation_custom(grob = insetMap, xmin = -74.12, xmax = -74.37,
                        ymin = 20.31, ymax = 20.42)

```

Total DNA was extracted from caudal peduncle muscle tissues using the DNAeasy tissue kit (QIAGEN). Partial sequences of the mitochondrial gene cytochrome oxidase I (COI) were amplified through the polymerase chain reaction in 20 μL of reaction volume (one unit of GoTaq DNA polymerase-Promega, 0.2 nM of each primer, 0.2 μM dNTPs, and 1.5 mM MgCl2) [@lara2010dna]. We used the primers COIf (5´-AAYCAYAAAGAYGGYACCCT-3´) and COIr (3´-CCTCNGGRTGNCCRAAGAAYCA-5´) [@palumbi1991simple].

PCR products were purified using ExoSAP-IT™ PCR Product Cleanup Reagent (Affymetrix Inc./ USB, Cleveland, OH, USA), and then sequenced on an ABI 3100 automated sequencer (Applied Biosystems). Raw sequence files were edited and aligned with the software Geneious v.9.1 [@kearse2012geneious], and submitted to GenBank database (accession numbers #####). We used sequences of the genus *Sicydium* as reference for recovering the taxonomic identity of the "tetis". Sequences of *Awaous banana* and *A. ocellaris* were used as outgroup. All non-Cuban sequences were recovered from GenBank using the R package *rentrez* [@winter2017rentrez]. Duplicated sequences and sequences with missing data were excluded. Sequences were edited using the R packages *seqinr* [@charif2007seqinr], *ape* [@paradis2015package], and *biostrings* [@pages2013package]. Final alignment was obtained with the algorithm ClustalOmega (@sievers2011fast) using the R package *msa* [@bodenhofer2015msa].

```{r gettingAPI}

                                                            # DOWNLOADING SEQS FROM GENBANK
# rentrez: this package provides an R interface NCBI API and allows to search databases like GenBank (http://www.ncbi.nih.gov/genbank/)

Sicydium_COI <- "Sicydium[Organism] AND COI[Gene]"

# I set retmax=580 because after running this by default, I got a message with the number of hits (580 in this case)
Sicydium_COI_search <- entrez_search(db = "nuccore", term = Sicydium_COI, retmax = 580)

# given that the search is too big for the system, it needs to be split into two smaller ones
Sicydium_COI_seqs1 <- entrez_fetch(db = "nuccore", id = Sicydium_COI_search$ids[0:295], rettype = "fasta")
Sicydium_COI_seqs2 <- entrez_fetch(db = "nuccore", id = Sicydium_COI_search$ids[296:580], rettype = "fasta")

# GETTING OUTGROUP SEQUENCES OF AWAOUS BANANA (2) AND A. OCELLARIS (2)

Awaous_bananaCOI <- "Awaous_banana[Organism] AND COI[Gene]"
Awaous_bananaCOI_search <- entrez_search(db = "nuccore", term = Awaous_bananaCOI, retmax = 2)
Awaous_bananaCOI_seqs <- entrez_fetch(db = "nuccore", id = Awaous_bananaCOI_search$ids, rettype = "fasta")

Awaous_ocellarisCOI <- "Awaous_ocellaris[Organism] AND COI[Gene]"
Awaous_ocellarisCOI_search <- entrez_search(db = "nuccore", term = Awaous_ocellarisCOI, retmax = 4)
Awaous_ocellarisCOI_seqs <- entrez_fetch(db = "nuccore", id = Awaous_ocellarisCOI_search$ids, rettype = "fasta")

# storing ALL sequences in a single fasta file
write(Sicydium_COI_seqs1, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n")
write(Sicydium_COI_seqs2, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)   # 'append = TRUE', I can keep adding seqs
write(Awaous_bananaCOI_seqs, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)
write(Awaous_ocellarisCOI_seqs, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)

```

```{r editingAPI}

# seqinr: package for sequence manipulation 
# reading the file and renaming the sequences

Sicydium_COI_seqinrFormat <- read.fasta(file ="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)

Sicydium_COI_names <- attr(Sicydium_COI_seqinrFormat, "name")      # editing species name
Sicydium_COI_names <- gsub("\\..*", "", Sicydium_COI_names)        # eliminating characters after "." using gsub
Sicydium_COI_names <- gsub("^.*\\|", "", Sicydium_COI_names)       # eliminating characters before "|" using gsub

# using library "ape"

Sicydium_COI_apeFormat <- read.GenBank(Sicydium_COI_names)         # this line takes a few minutes to run

#attr(Sicydium_COI_apeFormat, "species")
#names(Sicydium_COI_apeFormat)

# creating a vector (to be added as name of the seqs) with the species, GenBank accession number, and type of gene
Sicydium_COI_GenBank_Ids <- paste(attr(Sicydium_COI_apeFormat, "species"), names(Sicydium_COI_apeFormat), sep = "_COI_")

# writing a new fasta file (2nd) using the vector above 
write.fasta(sequences = Sicydium_COI_seqinrFormat, names = Sicydium_COI_GenBank_Ids, nbchar = 10,
            file.out = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_2-seqinrFormat")

                                     # EDITTING SEQS AND GETTING THEM ALL READY FOR THE PHYLOGENETIC ANALYSIS

### REMOVING sequences of incomplete/unspecific NAMES; that is, those with 'sp.'
nameFasta <- read.fasta(file = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_2-seqinrFormat",
                        seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)

fastaNames <- strsplit(names(nameFasta), '_')                  # creating a vector splitting species names
completeSeqs <- which(! grepl("sp.", fastaNames))              # selecting only those with specific name in them       # 562 seqs

# writing a new fasta (3rd) file with only those seqs with full scientific names
write.fasta(sequences = nameFasta[completeSeqs], names = names(nameFasta)[completeSeqs],
            file.out = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_3-names.fasta")

### REMOVING sequences of LENGTH shorter than 600 pb, given that the gene COI has between 627 and 658 bp.
lengthFasta <- read.fasta(file = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_3-names.fasta",
                          seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
fastaLength <- lengthFasta[which(getLength(lengthFasta) >= 600)]       
write.fasta(sequences = fastaLength, names = names(fastaLength),
            file.out = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_4-lengths.fasta")  # 401 seqs

### Removing DUPLICATE sequences
# using "seqinr" to do it 
# for not using attributes to name sequences
duplicatesFasta <- read.fasta("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_4-lengths.fasta",
                              set.attributes = FALSE) 
seqnames <- names(duplicatesFasta)
dups <- grep(TRUE,duplicated(duplicatesFasta))
uniqueNames <- seqnames[-dups]
uniqueSeqs <- duplicatesFasta[-dups] 
write.fasta(sequences = uniqueSeqs, names = uniqueNames,
            file.out = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta")     # 204 seqs

# adding samples from Cuba (n=6)
#mySamples <- read.fasta("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_Cuba.fasta")
#write.fasta(mySamples, names(mySamples),
 #           "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta", open = "a")    # 210 seqs

```

```{r alignment}
invisible({capture.output({                ### ALIGNING the sequences        # libraries "msa" and "Biostrings"

alignmentFasta <- readDNAStringSet("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta")
ClustalOAlignment <- msa(alignmentFasta, "ClustalOmega")
#print(ClustalOAlignment, showConsensus=FALSE, showNames=TRUE, show="complete")

# REMOVING samples with to many GAPS at the beginning and the end of the sequences 
colmask(ClustalOAlignment) <- IRanges(start=c(1,666),end=c(25,679))                
ClustalOAlignment2 <- as(ClustalOAlignment, "DNAStringSet")
writeXStringSet(ClustalOAlignment2,
                file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_6-alignment.fasta")   # 210 seqs -- 640 bp

### Replacing gaps at the begining and end with Ns
gapsFasta <- read.fasta(file = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_6-alignment.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
replacingGaps <- lapply(gapsFasta, function(x) gsub("-", "N", x))
write.fasta(replacingGaps,names(replacingGaps),"/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")

})})

```

```{r warning = FALSE}
invisible({capture.output({
                                          ### GETTING the READING FRAME     --- library "systemPipeR"
gettingRF <- function(x){
  fasta <- readDNAStringSet("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")
  orf1 <- predORF(fasta[1:10], n=1, type="df", mode="orf", strand="sense")   # predORF() do not read seqs with gaps
  # predicts longest ORF for sense strand in each query sequence                                                                          
  longestSeq1 <- orf1$start[orf1$width == max(orf1$width)]
  codonPosition1 <- longestSeq1[1]/3
  codonConfirmation1 <- round(codonPosition1)
  if (codonPosition1==codonConfirmation1){
   file.rename("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta", "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")                     # 640 bp
  print("The Reading Frame of these seqs is +1, no further action is required, go to modelTesting")
    } else {
    if (codonPosition1!=codonConfirmation1){
      tempLength <- length(fasta[[1]])
      #remove first nucleotide position
      newRF2 <- subseq(fasta, start = 2, end = tempLength)      # 639 bp
      # getting new RF from the new alignment with -1 bp.
      orf2 <- predORF(newRF2[1:10], n=1, type="df", mode="orf", strand="sense") 
      # checking size gain
      longestSeq2 <- orf2$start[orf2$width == max(orf2$width)]
      codonPosition2 <- (longestSeq2[1]+2)/3
      codonConfirmation2 <- round(codonPosition2)
      if (codonPosition2==codonConfirmation2){
        writeXStringSet(newRF2, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")
      print("The Reading Frame of these seqs was +2, new fasta file created, go to modelTesting")
      } else {
      if (codonPosition2!=codonConfirmation2){
        #remove first nucleotide position
        tempLength <- length(fasta[[1]])
        newRF3 <- subseq(fasta, start = 3, end = tempLength)
        writeXStringSet(newRF3, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")    # 638 bp
      print("The Reading Frame of these seqs was +3, new fasta file created, go to modelTesting")
     }
   }  
  }
 }
}

gettingRF(x)

})})

```

The nucleotide substitution model that best explained our data was estimated using the function modelTest from the R package phangorn [@schliep2011phangorn]. The phylogenetic tree was reconstructed following a maximum mikelihood approach [@felsenstein1981evolutionary] using the functions optim.pml and bootstrap.pml, also from the package phangorn. The tree was generated and visualized using the R package ggtree [@yu2017ggtree].

```{r bestModel, warning = FALSE}
                                      # FINDING THE BEST NUCLEOTIDE SUSBSTITUTION MODEL

# libraries ape and phangorn (modelTest is a function of the package phangorn)

gettingBestModel <- function(x){
  modelFasta <- read.dna(file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
  formatPhyDat <- as.phyDat(modelFasta)                #getting the class used by modelTest: phyDat or pml
  modelChoice <- modelTest(formatPhyDat, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 4)
  return(modelChoice)
}

bestModel <- gettingBestModel(x)                # getting the AICc for each competing model

# DISCLAIMER: visually inspecting the AICc is better than just getting the best model automatically, because it allows to compares models with similar AICc and it is my personal opinion to choose the one with the least parameters.  
# The alternative is to get automatically the first model that the data.frame modelTest gives and storing it as a variable. In order to get this as reproducible as possible, I chose the alternative.

```

```{r tree, include=TRUE, results = 'hide', fig.show="hide"}

                                                        # BUILINDG THE TREE
# --- library: phangorn
env = attr(bestModel[order(bestModel$AICc),], "env")    # getting the AICc for each competing model sorted
fitStart = eval(get(bestModel$Model[which.min(bestModel$AICc)], env), env)    # initial fit for the model chosen

# avoid to create a tree every time the code runs
#set.seed(2015-07-01)

fit = optim.pml(fitStart, rearrangement = "stochastic", optGamma=TRUE, optInv=FALSE, model="HKY")     # seeking a maximum likelihood solution
bs = bootstrap.pml(fit, bs=100, optNni=TRUE, multicore=TRUE, mc.cores = 4)

# saving the tree in the folder 'analysis/figures/'
bsTree <- plotBS(midpoint(fit$tree), bs, p = 75, type="p", bs.col = "blue")     # fit$tree is where the optimal tree is stored
optimalTree <- plotBS(bsTree, type = "p", cex = 0.2, font = 1)
# export tree with bootstrap values as node labels
write.tree(optimalTree, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/derived_data/Sicydium_ML_tree.tre")

```

# Results

(Figure 3.1)

hshdhdhdh

```{r inline-code2, include = TRUE, results='hide'}
# the result wont be displayed in the final document
loglikelihood <- -1*(as.numeric(fit[1]))
loglikelihood

```

```{r figure_2, fig.cap = paste("ML phylogenetic tree  (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus Sicydium (n=197). Sequences highlited in green represent samples from Cuba (n=6). Numbers on branches represent boostrap values higher than 70.")}

# Editing and visualizing the tree to represent --- library: ggtree
tree <- read.tree(file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/derived_data/Sicydium_ML_tree.tre")

# getting the node number for each main clade. This info is important because those #s are the one needed to edit each main clade/species
nodeNumber <- ggtree(tree) + geom_text(aes(label=node), hjust=-.5, col = "blue", size = 1) # looks horrible but it gives the clade number

Sicydium_ML_tree <-suppressWarnings(suppressMessages(print(ggtree(tree, size = .4) + 
  xlim(-0.02, 0.26) +
  geom_treescale(x=0, y=-1.5, offset = -4) + 
  geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 70), size = 3, col= "black", vjust = -0.3, hjust=1.9, na.rm = FALSE) + 
  geom_nodepoint(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 70), size = 1, col= "black") + 
  geom_cladelabel(node=218, label="Sicydium plumieri", color="red", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=117, label="Sicydium brevifile", color="blue", hjust = -0.05, vjust = 0.3, fontsize = 3.5, fontface="italic") + 
  geom_cladelabel(node=320, label="Sicydium salvini", color="purple", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=328, label="Sicydium punctatum", color="brown", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=379, label="Sicydium buscki", color="black", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=415, label="Sicydium crenilabrum", color="orange", hjust = -0.03, vjust = 0.3, fontsize = 3.5, offset = 0.004, fontface="italic") +
  geom_cladelabel(node=212, label="Awaous spp.", color="darkgreen", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") +
  geom_hilight(node=377, fill="green") + 
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")))))

#Sicydium_ML_tree

```

```{r demo-plot, fig.cap="A plot of random numbers"}
plot(rnorm(10))
```

Figure \@ref(fig:demo-plot) shows how we can have a caption and cross-reference for a plot

```{r demo-inline-code}
x <- round(pi, 2)
```

Here is an example of inline code `r x` in the middle of a sentence.

# Discussion

# Conclusion

Busck´s Stone-Biting Goby, *Sicydium buscki*, is a component of the fish fry “teti” fishery in Eastern Cuba. This species represents the only taxa of the orden Gobiiformes confirmed by molecular tools as part of such a fishery.
More thorough investigations and identification of the species comprising the “teti” fishery are needed. It is imperative to include them in  local management plans and to avoid overfishing this resource.

# Acknowledgements

<!-- The following line inserts a page break  -->

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

<div id="refs"></div>

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```

### 2020-11-22 analysis/install.R

@@ -1,0 +1,11 @@
+# INSTALLING PACKAGES AND LIBRARIES - if needed
+packages = c("raster", "ggplot2", "ggthemes",
+             "rentrez", "ape", "seqinr", "msa", "Biostrings", "systemPipeR", "phangorn", "ggtree")
+
+package.check <- lapply(packages, FUN = function(x) {
+  if (!require(x, character.only = TRUE)) {
+    install.packages(x, dependencies = TRUE)
+    library(x, character.only = TRUE, warn.conflicts = FALSE, quietly = TRUE)
+  }
+ }
+)

### 2020-11-23 analysis/paper/paper.Rmd
files and folders needed

@@
       install.packages(x, dependencies = TRUE)
-      library(x, character.only = TRUE, warn.conflicts = FALSE, quietly = TRUE)
+      library(x, character.only = TRUE, include = FALSE, warn.conflicts = FALSE, quietly = TRUE)
     }
@@
                                           ### GETTING the READING FRAME     --- library "systemPipeR"
-gettingRF <- function(x){
+gettingRF <- function(alignedFasta){
   fasta <- readDNAStringSet("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")
@@
 
-gettingRF(x)
+gettingRF(alignedFasta)
 
@@
 
-gettingBestModel <- function(x){
+gettingBestModel <- function(fastaORF){
   modelFasta <- read.dna(file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
@@
 
-bestModel <- gettingBestModel(x)                # getting the AICc for each competing model
+bestModel <- gettingBestModel(fastaORF)                # getting the AICc for each competing model
 
@@
 
-```{r tree, include=TRUE, results = 'hide', fig.show="hide"}
+```{r tree, include=TRUE, results = 'hide', fig.show='hide'}
 

### 2020-11-23 analysis/install.R

@@ -1,11 +1,0 @@
-# INSTALLING PACKAGES AND LIBRARIES - if needed
-packages = c("raster", "ggplot2", "ggthemes",
-             "rentrez", "ape", "seqinr", "msa", "Biostrings", "systemPipeR", "phangorn", "ggtree")
-
-package.check <- lapply(packages, FUN = function(x) {
-  if (!require(x, character.only = TRUE)) {
-    install.packages(x, dependencies = TRUE)
-    library(x, character.only = TRUE, warn.conflicts = FALSE, quietly = TRUE)
-  }
- }
-)

### 2020-11-23 install.R

@@ -1,9 +1,16 @@
-# Download needed packages
+# Installing packages for binder.
 
-install.packages("rentrez") # Download and install ...
-install.packages("ape") 
+# I did not include here "rmarkdown", "testthat", or any other because the project doesnt rely directly on them to run
+
+install.packages("raster")
+install.packages("ggplot2")
+install.packages("ggthemes")
+install.packages("rentrez")
+install.packages("ape")
 install.packages("seqinr")
-install.packages("BiocManager")
-install.packages("phangorn")
+install.packages("msa")
+install.packages("Biostrings")
 install.packages("systemPipeR")
+install.packages("phangorn")
 install.packages("ggtree")
+

### 2020-11-23 tests/testthat.R

@@ -1,0 +1,11 @@
+install.packages('knitr')
+install.packages('testthat')
+
+library(knitr)
+library(testthat)
+
+knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
+source('analysis/paper/paper.R')
+testthat::test_file('tests/testthat/test-gettingORF.R')
+testthat::test_file('tests/testthat/test-gettingBestModel.R')
+system('rm analysis/paper/paper.R')

### 2020-11-23 tests/testthat/test-gettingBestModel.R

@@ -1,0 +1,16 @@
+
+context("test-process-data")
+
+data <- readr::read_csv(system.file("testdata",
+                                    "gillespie.csv",
+                                    package = "rrcompendium"))
+
+test_that("recoding works", {
+  expect_equal(unique(recode_system_size(data)$system_size),
+               c("A. 1000 total sites", "B. 100 total sites" ))
+})
+
+test_that("data dims preserved", {
+  expect_equal(dim(recode_system_size(data)),
+               dim(data))
+})

### 2020-11-23 README.Rmd
more files and folders

@@
 # Please put your title here to include it in the file below.
-Title <- "Title of your paper goes here"
+Title <- 'The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba'
+
 ```
@@
 
-[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh///master?urlpath=rstudio)
+[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/sheilarm23/rrcompendium/master?urlpath=rstudio)
 
@@
 The simplest way to explore the text, code and data is to click on
-[binder](https://mybinder.org/v2/gh///master?urlpath=rstudio)
+[binder](https://mybinder.org/v2/gh/sheilarm23/rrcompendium/master?urlpath=rstudio)
 to open an instance of RStudio in your browser, which will have the

### 2020-11-23 R/gettingBestModel.R

@@ -1,0 +1,13 @@
+#' gettingBestModel
+#' This function provides a dataframe with the list of models of nucleotide evolution sorted by AICc (from lowest to highest)
+#'
+#' @param fastaORF takes as input a dataset of aligned sequences of the same length and known ORF
+#' @return returns a dataframe with the list of models of evolution sorted by AICc, from the lowest to the highest AICc value
+#'
+#'
+gettingBestModel <- function(fastaORF){
+  modelFasta <- read.dna(file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
+  formatPhyDat <- as.phyDat(modelFasta)                #getting the class used by modelTest: phyDat or pml
+  modelChoice <- modelTest(formatPhyDat, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 4)
+  return(modelChoice)
+}

### 2020-11-23 R/gettingRF.R

@@ -1,0 +1,44 @@
+#' gettingRF
+#' This function determines the Open Reading Frame (ORF) of a dataset of nucleotide sequences
+#'
+#' @param alignedFasta takes as input a dataset of aligned sequences of the same length
+#'
+#' As output, this function either renames the input file if the sequences are in the ORF 1, or it writes a new fasta file if the input file was in ORF 2 or 3.
+#' The output file will be of ORF 1
+#'
+gettingRF <- function(alignedFasta){
+  fasta <- readDNAStringSet("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")
+  orf1 <- predORF(fasta[1:10], n=1, type="df", mode="orf", strand="sense")   # predORF() do not read seqs with gaps
+  # predicts longest ORF for sense strand in each query sequence
+  longestSeq1 <- orf1$start[orf1$width == max(orf1$width)]
+  codonPosition1 <- longestSeq1[1]/3
+  codonConfirmation1 <- round(codonPosition1)
+  if (codonPosition1==codonConfirmation1){
+    file.rename("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta", "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")                     # 640 bp
+    print("The Reading Frame of these seqs is +1, no further action is required, go to modelTesting")
+  } else {
+    if (codonPosition1!=codonConfirmation1){
+      tempLength <- length(fasta[[1]])
+      #remove first nucleotide position
+      newRF2 <- subseq(fasta, start = 2, end = tempLength)      # 639 bp
+      # getting new RF from the new alignment with -1 bp.
+      orf2 <- predORF(newRF2[1:10], n=1, type="df", mode="orf", strand="sense")
+      # checking size gain
+      longestSeq2 <- orf2$start[orf2$width == max(orf2$width)]
+      codonPosition2 <- (longestSeq2[1]+2)/3
+      codonConfirmation2 <- round(codonPosition2)
+      if (codonPosition2==codonConfirmation2){
+        writeXStringSet(newRF2, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")
+        print("The Reading Frame of these seqs was +2, new fasta file created, go to modelTesting")
[…]

### 2020-11-23 install.R
packages for binder

@@ -1,7 +1,4 @@
-# Installing packages for binder.
-
-# I did not include here "rmarkdown", "testthat", or any other because the project doesnt rely directly on them to run
-
+# Installing packages for binder        # I did not include here "rmarkdown", "testthat", or any other because the project doesnt rely directly on them to run
 install.packages("raster")
 install.packages("ggplot2")
 install.packages("ggthemes")

### 2020-11-24 tests/testthat.R
tests done

@@ -6,6 +6,8 @@ library(testthat)
 
 knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
 source('analysis/paper/paper.R')
-testthat::test_file('tests/testthat/test-gettingORF.R')
+
+testthat::test_file('tests/testthat/test-gettingRF.R')
 testthat::test_file('tests/testthat/test-gettingBestModel.R')
+
 system('rm analysis/paper/paper.R')

### 2020-11-24 tests/testthat/test-gettingBestModel.R

@@ -1,16 +1,23 @@
 
-context("test-process-data")
+context("test-gettingBestModel")
 
-data <- readr::read_csv(system.file("testdata",
-                                    "gillespie.csv",
-                                    package = "rrcompendium"))
-
-test_that("recoding works", {
-  expect_equal(unique(recode_system_size(data)$system_size),
-               c("A. 1000 total sites", "B. 100 total sites" ))
+test1 <- test_that("input file is of a given class", {
+  expect_is(fastaRF, "DNAbin")
 })
 
-test_that("data dims preserved", {
-  expect_equal(dim(recode_system_size(data)),
-               dim(data))
+test2 <- test_that("gettingBestModel outputs an object of class modelTest", {
+  expect_is(gettingBestModel(fastaRF), "modelTest")
 })
+
+test3 <- test_that("gettingBestModel outputs a data.frame", {
+  expect_is(gettingBestModel(fastaRF), "data.frame")
+})
+
+test4 <- test_that("gettingBestModel(fastaRF) has 24 rows based on the number of model selected", {
+  expect_true(nrow(gettingBestModel(fastaRF)) == 24)
+})
+
+test5 <- test_that("gettingBestModel(fastaRF) has sorted the AICc column from lowest to highest value", {
+  expect_true(gettingBestModel(fastaRF)$AICc[1] == min(gettingBestModel(fastaRF)$AICc))
+})
+

### 2020-11-24 tests/testthat/test-gettingORF.R

@@ -1,16 +1,0 @@
-
-context("test-process-data")
-
-data <- readr::read_csv(system.file("testdata",
-                                    "gillespie.csv",
-                                    package = "rrcompendium"))
-
-test_that("recoding works", {
-  expect_equal(unique(recode_system_size(data)$system_size),
-               c("A. 1000 total sites", "B. 100 total sites" ))
-})
-
-test_that("data dims preserved", {
-  expect_equal(dim(recode_system_size(data)),
-               dim(data))
-})

### 2020-11-24 tests/testthat/test-gettingRF.R

@@ -1,0 +1,14 @@
+
+context("test-gettingRF")
+
+test1 <- test_that("gettingRF errors out when a file of class other than DNAStringSet is taken", {
+  expect_error(class(alignedFasta)!= c("DNAStringSet"))
+})
+
+test2 <- test_that("gettingRF errors out when an alignment includes gaps", {
+  expect_error(any(letterFrequency(alignedFasta, "-", as.prob=TRUE))>0)
+})
+
+
+
+

### 2020-11-24 analysis/paper/paper.Rmd
functions done

@@
 
-```{r figure_1, fig.cap="Sampling site at Yumuri river mouth, Eastern Cuba"}
+```{r mainMap, fig.cap="Sampling site at Yumuri river mouth, Eastern Cuba", cache = TRUE}
 
@@
 
-```{r gettingAPI}
+```{r gettingAPI, cache = TRUE}
 
@@
 
-```{r editingAPI}
+```{r editingAPI, cache = TRUE}
 
@@
 
-```{r alignment}
+```{r alignment, cache = TRUE}
 invisible({capture.output({                ### ALIGNING the sequences        # libraries "msa" and "Biostrings"
@@
                                           ### GETTING the READING FRAME     --- library "systemPipeR"
+
+alignedFasta <- readDNAStringSet("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")
+
 gettingRF <- function(alignedFasta){
-  fasta <- readDNAStringSet("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")
-  orf1 <- predORF(fasta[1:10], n=1, type="df", mode="orf", strand="sense")   # predORF() do not read seqs with gaps
-  # predicts longest ORF for sense strand in each query sequence                                                                          
+  if (class(alignedFasta)!= c("DNAStringSet")){
+    stop("This function only takes files of class 'DNAStringSet'. Use package 'Biostrings' to read the input file")
+  }
+  if (any(letterFrequency(alignedFasta, "-", as.prob=TRUE))>0){
+    stop("This function do not take gaps as values/strings.
+Go and replaced gaps with Ns (if the gaps are at the begining or end of the sequences) or choose a subset of sequences that do not contain gaps")
+  }
+  # predORF() do not read seqs with gaps   # predicts longest ORF for sense strand in each query sequence               
+  orf1 <- predORF(alignedFasta[1:10], n=1, type="df", mode="orf", strand="sense")                                       
   longestSeq1 <- orf1$start[orf1$width == max(orf1$width)]
@@
   codonConfirmation1 <- round(codonPosition1)
+  length(alignedFasta)
   if (codonPosition1==codonConfirmation1){
@@
     if (codonPosition1!=codonConfirmation1){
-      tempLength <- length(fasta[[1]])
+      tempLength <- length(alignedFasta[[1]])
       #remove first nucleotide position
-      newRF2 <- subseq(fasta, start = 2, end = tempLength)      # 639 bp
+      newRF2 <- subseq(alignedFasta, start = 2, end = tempLength)      # 639 bp
       # getting new RF from the new alignment with -1 bp.
@@
         #remove first nucleotide position
-        tempLength <- length(fasta[[1]])
-        newRF3 <- subseq(fasta, start = 3, end = tempLength)
+        tempLength <- length(alignedFasta[[1]])
+        newRF3 <- subseq(alignedFasta, start = 3, end = tempLength)
         writeXStringSet(newRF3, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")    # 638 bp
@@
 
-gettingBestModel <- function(fastaORF){
-  modelFasta <- read.dna(file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
-  formatPhyDat <- as.phyDat(modelFasta)                #getting the class used by modelTest: phyDat or pml
+fastaRF <- read.dna(file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
+
+gettingBestModel <- function(fastaRF){
+  if (class(fastaRF)!= c("DNAbin")){
+    stop("This function only takes files of class 'DNAbin'. Use package 'ape' to read the input file")
+  }
+  formatPhyDat <- as.phyDat(fastaRF)                #getting the class used by modelTest: phyDat or pml
   modelChoice <- modelTest(formatPhyDat, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 4)
-  return(modelChoice)
+  modelChoiceSorted <- modelChoice[order(modelChoice$AICc),]
+  return(modelChoiceSorted)
 }
 
-bestModel <- gettingBestModel(fastaORF)                # getting the AICc for each competing model
+bestModel <- gettingBestModel(fastaRF)                # getting the AICc for each competing model
 
@@
 # --- library: phangorn
-env = attr(bestModel[order(bestModel$AICc),], "env")    # getting the AICc for each competing model sorted
+env = attr(bestModel, "env")    # getting the AICc for each competing model sorted
 fitStart = eval(get(bestModel$Model[which.min(bestModel$AICc)], env), env)    # initial fit for the model chosen
@@
 # the result wont be displayed in the final document
-loglikelihood <- -1*(as.numeric(fit[1]))
+loglikelihood <- round(-1*(as.numeric(fit[1])), 2)
 loglikelihood

### 2020-11-25 R/gettingBestModel.R
functions done

@@ -1,13 +1,16 @@
 #' gettingBestModel
 #' This function provides a dataframe with the list of models of nucleotide evolution sorted by AICc (from lowest to highest)
 #'
-#' @param fastaORF takes as input a dataset of aligned sequences of the same length and known ORF
+#' @param fastaRF takes as input a dataset of aligned sequences of the same length and known ORF
 #' @return returns a dataframe with the list of models of evolution sorted by AICc, from the lowest to the highest AICc value
 #'
 #'
-gettingBestModel <- function(fastaORF){
-  modelFasta <- read.dna(file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
-  formatPhyDat <- as.phyDat(modelFasta)                #getting the class used by modelTest: phyDat or pml
+gettingBestModel <- function(fastaRF){
+  if (class(fastaRF)!= c("DNAbin")){
+    stop("This function only takes files of class 'DNAbin'. Use package 'ape' to read the input file")
+  }
+  formatPhyDat <- as.phyDat(fastaRF)                #getting the class used by modelTest: phyDat or pml
   modelChoice <- modelTest(formatPhyDat, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 4)
-  return(modelChoice)
+  modelChoiceSorted <- modelChoice[order(modelChoice$AICc),]
+  return(modelChoiceSorted)
 }

### 2020-11-25 R/gettingRF.R
functions done

@@ -1,26 +1,32 @@
 #' gettingRF
-#' This function determines the Open Reading Frame (ORF) of a dataset of nucleotide sequences
+#' This function determines the Reading Frame (RF) of a dataset of nucleotide sequences
 #'
 #' @param alignedFasta takes as input a dataset of aligned sequences of the same length
 #'
 #' As output, this function either renames the input file if the sequences are in the ORF 1, or it writes a new fasta file if the input file was in ORF 2 or 3.
-#' The output file will be of ORF 1
+#' The output file will be of RF 1
 #'
 gettingRF <- function(alignedFasta){
-  fasta <- readDNAStringSet("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")
-  orf1 <- predORF(fasta[1:10], n=1, type="df", mode="orf", strand="sense")   # predORF() do not read seqs with gaps
-  # predicts longest ORF for sense strand in each query sequence
+  if (class(alignedFasta)!= c("DNAStringSet")){
+    stop("This function only takes files of class 'DNAStringSet'. Use package 'Biostrings' to read the input file")
+  }
+  if (any(letterFrequency(alignedFasta, "-", as.prob=TRUE))>0){
+    stop("This function do not take gaps as values/strings. Go and replaced gaps with Ns (if the gaps are at the begining or end of the sequences) or choose a subset of sequences that do not contain gaps")
+  }
+  # predORF() do not read seqs with gaps   # predicts longest ORF for sense strand in each query sequence
+  orf1 <- predORF(alignedFasta[1:10], n=1, type="df", mode="orf", strand="sense")
   longestSeq1 <- orf1$start[orf1$width == max(orf1$width)]
   codonPosition1 <- longestSeq1[1]/3
   codonConfirmation1 <- round(codonPosition1)
+  length(alignedFasta)
   if (codonPosition1==codonConfirmation1){
     file.rename("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta", "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")                     # 640 bp
     print("The Reading Frame of these seqs is +1, no further action is required, go to modelTesting")
   } else {
     if (codonPosition1!=codonConfirmation1){
-      tempLength <- length(fasta[[1]])
[…]

### 2020-11-25 tests/testthat.R
removing paper.R from testthat

@@ -5,9 +5,8 @@ library(knitr)
 library(testthat)
 
 knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
-source('analysis/paper/paper.R')
 
 testthat::test_file('tests/testthat/test-gettingRF.R')
 testthat::test_file('tests/testthat/test-gettingBestModel.R')
 
-system('rm analysis/paper/paper.R')
+

### 2020-11-25 analysis/paper/paper.Rmd
testing travis2

@@
 
-gettingRF <- function(alignedFasta){
+getRF <- function(alignedFasta){
   if (class(alignedFasta)!= c("DNAStringSet")){
@@
 
-gettingRF(alignedFasta)
+getRF(alignedFasta)
 
@@
 
-gettingBestModel <- function(fastaRF){
+getBestModel <- function(fastaRF){
   if (class(fastaRF)!= c("DNAbin")){
@@
 
-bestModel <- gettingBestModel(fastaRF)                # getting the AICc for each competing model
+bestModel <- getBestModel(fastaRF)                # getting the AICc for each competing model
 

### 2020-11-25 R/getBestModel.R

@@ -1,11 +1,11 @@
-#' gettingBestModel
+#' getBestModel
 #' This function provides a dataframe with the list of models of nucleotide evolution sorted by AICc (from lowest to highest)
 #'
 #' @param fastaRF takes as input a dataset of aligned sequences of the same length and known ORF
 #' @return returns a dataframe with the list of models of evolution sorted by AICc, from the lowest to the highest AICc value
 #'
 #'
-gettingBestModel <- function(fastaRF){
+getBestModel <- function(fastaRF){
   if (class(fastaRF)!= c("DNAbin")){
     stop("This function only takes files of class 'DNAbin'. Use package 'ape' to read the input file")
   }

### 2020-11-25 R/getRF.R

@@ -1,4 +1,4 @@
-#' gettingRF
+#' getRF
 #' This function determines the Reading Frame (RF) of a dataset of nucleotide sequences
 #'
 #' @param alignedFasta takes as input a dataset of aligned sequences of the same length
@@ -6,7 +6,7 @@
 #' As output, this function either renames the input file if the sequences are in the ORF 1, or it writes a new fasta file if the input file was in ORF 2 or 3.
 #' The output file will be of RF 1
 #'
-gettingRF <- function(alignedFasta){
+getRF <- function(alignedFasta){
   if (class(alignedFasta)!= c("DNAStringSet")){
     stop("This function only takes files of class 'DNAStringSet'. Use package 'Biostrings' to read the input file")
   }

### 2020-11-25 README.Rmd
testing travis2

@@
 
-This repository contains the data and code for our paper:
+This repository contains the data and code for our future paper:
 
@@
 
-Our pre-print is online here:
+which is still a work in progress...
+
+Our pre-print will be online here:
 
@@
 
-> Rodriguez-Machado, S., Rodríguez-Cabrera, T.M., and Chakrabarty, P.
-(2020), (`r format(Sys.Date(), "%Y")`).
+> Rodriguez-Machado, S.
+(`r format(Sys.Date(), "%Y")`).
 _Compendium of R code and data for `The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba`_.
@@
 Docker images to ensure a consistent and reproducible computational
-environment. These Docker images can also be used locally.
+environment. This way will take a few minutes to be ready but you will not need to install any packages on your computer.
 

### 2020-11-25 tests/testthat.R

@@ -6,7 +6,7 @@ library(testthat)
 
 knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
 
-testthat::test_file('tests/testthat/test-gettingRF.R')
-testthat::test_file('tests/testthat/test-gettingBestModel.R')
+testthat::test_file('tests/testthat/test-getRF.R')
+testthat::test_file('tests/testthat/test-getBestModel.R')
 
-
+system('rm paper.R')

### 2020-11-25 tests/testthat/test-getBestModel.R

@@ -1,0 +1,19 @@
+
+context("test-getBestModel")
+
+test1 <- test_that("getBestModel errors out when a file of class other than 'DNAbin' is taken", {
+  expect_error(class(fastaRF)!= c("DNAbin"))
+})
+
+test2 <- test_that("getBestModel outputs an object of class modelTest", {
+  expect_is(class(getBestModel(fastaRF)), "data.frame")
+})
+
+test3 <- test_that("getBestModel(fastaRF) has 24 rows based on the number of model selected", {
+  expect_true(nrow(getBestModel(fastaRF)) == 24)
+})
+
+test4 <- test_that("getBestModel(fastaRF) has sorted the AICc column from lowest to highest value", {
+  expect_true(getBestModel(fastaRF)$AICc[1] == min(getBestModel(fastaRF)$AICc))
+})
+

### 2020-11-25 tests/testthat/test-getRF.R

@@ -1,0 +1,14 @@
+
+context("test-getRF")
+
+test1 <- test_that("getRF errors out when a file of class other than DNAStringSet is taken", {
+  expect_error(class(alignedFasta)!= c("DNAStringSet"))
+})
+
+test2 <- test_that("getRF errors out when an alignment includes gaps", {
+  expect_error(any(letterFrequency(alignedFasta, "-", as.prob=TRUE))>0)
+})
+
+
+
+

### 2020-11-25 tests/testthat/test-gettingBestModel.R

@@ -1,23 +1,0 @@
-
-context("test-gettingBestModel")
-
-test1 <- test_that("input file is of a given class", {
-  expect_is(fastaRF, "DNAbin")
-})
-
-test2 <- test_that("gettingBestModel outputs an object of class modelTest", {
-  expect_is(gettingBestModel(fastaRF), "modelTest")
-})
-
-test3 <- test_that("gettingBestModel outputs a data.frame", {
-  expect_is(gettingBestModel(fastaRF), "data.frame")
-})
-
-test4 <- test_that("gettingBestModel(fastaRF) has 24 rows based on the number of model selected", {
-  expect_true(nrow(gettingBestModel(fastaRF)) == 24)
-})
-
-test5 <- test_that("gettingBestModel(fastaRF) has sorted the AICc column from lowest to highest value", {
-  expect_true(gettingBestModel(fastaRF)$AICc[1] == min(gettingBestModel(fastaRF)$AICc))
-})
-

### 2020-11-25 tests/testthat/test-gettingRF.R

@@ -1,14 +1,0 @@
-
-context("test-gettingRF")
-
-test1 <- test_that("gettingRF errors out when a file of class other than DNAStringSet is taken", {
-  expect_error(class(alignedFasta)!= c("DNAStringSet"))
-})
-
-test2 <- test_that("gettingRF errors out when an alignment includes gaps", {
-  expect_error(any(letterFrequency(alignedFasta, "-", as.prob=TRUE))>0)
-})
-
-
-
-

### 2020-11-25 tests/testthat/test-getBestModel.R
travis3

@@ -6,14 +6,14 @@ test1 <- test_that("getBestModel errors out when a file of class other than 'DNA
 })
 
 test2 <- test_that("getBestModel outputs an object of class modelTest", {
-  expect_is(class(getBestModel(fastaRF)), "data.frame")
+  expect_is(class(getBestModel(fastaRF)), c("modelTest", "data.frame"))
 })
 
 test3 <- test_that("getBestModel(fastaRF) has 24 rows based on the number of model selected", {
-  expect_true(nrow(getBestModel(fastaRF)) == 24)
+  expect_equal(nrow(getBestModel(fastaRF)), 24)
 })
 
 test4 <- test_that("getBestModel(fastaRF) has sorted the AICc column from lowest to highest value", {
-  expect_true(getBestModel(fastaRF)$AICc[1] == min(getBestModel(fastaRF)$AICc))
+  expect_equal(getBestModel(fastaRF)$AICc[1], min(getBestModel(fastaRF)$AICc))
 })
 

### 2020-11-25 analysis/paper/paper.Rmd
travis4

@@
 
-```{r mainMap, fig.cap="Sampling site at Yumuri river mouth, Eastern Cuba", cache = TRUE}
+```{r mainMap, fig.cap="Sampling site at Yumuri river mouth, Eastern Cuba"}
 
@@
 
-```{r gettingAPI, cache = TRUE}
+```{r gettingAPI}
 
@@
 
-```{r editingAPI, cache = TRUE}
+```{r editingAPI}
 
@@
 
-```{r alignment, cache = TRUE}
+```{r alignment}
 invisible({capture.output({                ### ALIGNING the sequences        # libraries "msa" and "Biostrings"
@@
 replacingGaps <- lapply(gapsFasta, function(x) gsub("-", "N", x))
-write.fasta(replacingGaps,names(replacingGaps),"/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")
+write.fasta(replacingGaps, names(replacingGaps), "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")
 
@@
   modelChoice <- modelTest(formatPhyDat, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 4)
-  modelChoiceSorted <- modelChoice[order(modelChoice$AICc),]
+  modelChoiceSorted <- as.data.frame(modelChoice[order(modelChoice$AICc),])
   return(modelChoiceSorted)
@@
 fit = optim.pml(fitStart, rearrangement = "stochastic", optGamma=TRUE, optInv=FALSE, model="HKY")     # seeking a maximum likelihood solution
+#set.seed(123)    # getting the same tree topology each time the code runs 
 bs = bootstrap.pml(fit, bs=100, optNni=TRUE, multicore=TRUE, mc.cores = 4)
@@
 
-```{r figure_2, fig.cap = paste("ML phylogenetic tree  (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus Sicydium (n=197).
+```{r Sicydium_ML_tree, fig.cap = paste("ML phylogenetic tree  (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus Sicydium (n=197).
 Sequences highlited in green represent samples from Cuba (n=6).
-Numbers on branches represent boostrap values higher than 70.")}
+Numbers on branches represent boostrap values higher than 70.", fig.width = 11.8, fig.height = 9.84)}
 
@@
 #Sicydium_ML_tree
-
 ```

### 2020-11-26 tests/testthat/test-getBestModel.R
travis5

@@ -1,19 +1,14 @@
 
 context("test-getBestModel")
 
-test1 <- test_that("getBestModel errors out when a file of class other than 'DNAbin' is taken", {
-  expect_error(class(fastaRF)!= c("DNAbin"))
-})
+test_that("getBestModel", {
+  expect_is(getBestModel, "function")
 
-test2 <- test_that("getBestModel outputs an object of class modelTest", {
-  expect_is(class(getBestModel(fastaRF)), c("modelTest", "data.frame"))
-})
+  bestModel <- getBestModel(fastaRF)
+  expect_true(class(bestModel) == "data.frame")    # "getBestModel outputs an object of class data.frame"
+  expect_equal(nrow(bestModel), 24)  # getBestModel(fastaRF) has 24 rows based on the number of model selected"
+  expect_equal(bestModel$AICc[1], min(bestModel$AICc))  # getBestModel(fastaRF) has sorted the AICc column from lowest to highest values
 
-test3 <- test_that("getBestModel(fastaRF) has 24 rows based on the number of model selected", {
-  expect_equal(nrow(getBestModel(fastaRF)), 24)
-})
+  expect_error(getBestModel(fastaRF = c("ANTCATTTGA")))
 
-test4 <- test_that("getBestModel(fastaRF) has sorted the AICc column from lowest to highest value", {
-  expect_equal(getBestModel(fastaRF)$AICc[1], min(getBestModel(fastaRF)$AICc))
 })
-

### 2020-11-27 analysis/paper/paper.Rmd
missing final text and location

@@
 abstract: |
-  Text of abstract
+  texting texting
 keywords: |
@@
 
+editor_options: 
+  chunk_output_type: console
 ---
@@
 ```{r setup, echo = FALSE}    
-# {r settings, include = FALSE}
 knitr::opts_chunk$set(
@@
 #knitr::opts_knit$set(root.dir = '../rrcompendium/')
+
+#getwd()
 
@@
 # storing ALL sequences in a single fasta file
-write(Sicydium_COI_seqs1, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n")
-write(Sicydium_COI_seqs2, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)   # 'append = TRUE', I can keep adding seqs
-write(Awaous_bananaCOI_seqs, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)
-write(Awaous_ocellarisCOI_seqs, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)
+write(Sicydium_COI_seqs1, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n")
+write(Sicydium_COI_seqs2, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)   # 'append = TRUE', I can keep adding seqs
+write(Awaous_bananaCOI_seqs, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)
+write(Awaous_ocellarisCOI_seqs, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)
 
@@
 
-Sicydium_COI_seqinrFormat <- read.fasta(file ="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
+Sicydium_COI_seqinrFormat <- read.fasta(file ="../rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
 
@@
 # using library "ape"
-
 Sicydium_COI_apeFormat <- read.GenBank(Sicydium_COI_names)         # this line takes a few minutes to run
-
-#attr(Sicydium_COI_apeFormat, "species")
-#names(Sicydium_COI_apeFormat)
 
@@
 write.fasta(sequences = Sicydium_COI_seqinrFormat, names = Sicydium_COI_GenBank_Ids, nbchar = 10,
-            file.out = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_2-seqinrFormat")
+            file.out = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_2-seqinrFormat.fasta")
 
@@
 ### REMOVING sequences of incomplete/unspecific NAMES; that is, those with 'sp.'
-nameFasta <- read.fasta(file = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_2-seqinrFormat",
+nameFasta <- read.fasta(file = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_2-seqinrFormat.fasta",
                         seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
@@
 fastaNames <- strsplit(names(nameFasta), '_')                  # creating a vector splitting species names
-completeSeqs <- which(! grepl("sp.", fastaNames))              # selecting only those with specific name in them       # 562 seqs
+completeSeqs <- which(! grepl("sp.", fastaNames))              # selecting only those with specific name in them
 
@@
 write.fasta(sequences = nameFasta[completeSeqs], names = names(nameFasta)[completeSeqs],
-            file.out = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_3-names.fasta")
+            file.out = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_3-names.fasta")
 
 ### REMOVING sequences of LENGTH shorter than 600 pb, given that the gene COI has between 627 and 658 bp.
-lengthFasta <- read.fasta(file = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_3-names.fasta",
-                          seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
+lengthFasta <- read.fasta(file = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_3-names.fasta",
+                          seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)                     # 586 seqs
 fastaLength <- lengthFasta[which(getLength(lengthFasta) >= 600)]
 write.fasta(sequences = fastaLength, names = names(fastaLength),
-            file.out = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_4-lengths.fasta")  # 401 seqs
+            file.out = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_4-lengths.fasta")             # 401 seqs
 
@@
 # for not using attributes to name sequences
-duplicatesFasta <- read.fasta("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_4-lengths.fasta",
-                              set.attributes = FALSE) 
+duplicatesFasta <- read.fasta("../rrcompendium/analysis/data/raw_data/Sicydium_COI_4-lengths.fasta", set.attributes = FALSE)
+
 seqnames <- names(duplicatesFasta)
@@
 write.fasta(sequences = uniqueSeqs, names = uniqueNames,
-            file.out = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta")     # 204 seqs
+            file.out = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta")               # 204 seqs
 
 # adding samples from Cuba (n=6)
-#mySamples <- read.fasta("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_Cuba.fasta")
-#write.fasta(mySamples, names(mySamples),
- #           "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta", open = "a")    # 210 seqs
+mySamples <- read.fasta("../rrcompendium/analysis/data/raw_data/Sicydium_COI_Cuba.fas")
+write.fasta(mySamples, names(mySamples),
+            "../rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta", open = "a")              # 210 seqs
 
@@
 
-alignmentFasta <- readDNAStringSet("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta")
+alignmentFasta <- readDNAStringSet("../rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta")
 ClustalOAlignment <- msa(alignmentFasta, "ClustalOmega")
-#print(ClustalOAlignment, showConsensus=FALSE, showNames=TRUE, show="complete")
+# print(ClustalOAlignment, showConsensus=FALSE, showNames=TRUE, show="complete")
 
-# REMOVING samples with to many GAPS at the beginning and the end of the sequences 
+# Checking for the position of sequences that were not edited properly so they still have nucleotides different that A, G, C, T 
+nucType <- alphabetFrequency(ClustalOAlignment)
+unknownNuc <- nucType[, c(5:14)]
+unknownPosition <- which(unknownNuc > 0, arr.ind=T)
+
+# REMOVING samples with too many GAPS at the beginning and the end of the sequences and UNKNOWN nucleotides (ex.: W, K, S) 
 colmask(ClustalOAlignment) <- IRanges(start=c(1,666),end=c(25,679))
+rowmask(ClustalOAlignment) <- IRanges(start=c(1),end=c(1))
 ClustalOAlignment2 <- as(ClustalOAlignment, "DNAStringSet")
 writeXStringSet(ClustalOAlignment2,
-                file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_6-alignment.fasta")   # 210 seqs -- 640 bp
+                file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_6-alignment.fasta")           # 209 seqs -- 640 bp
 
 ### Replacing gaps at the begining and end with Ns
-gapsFasta <- read.fasta(file = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_6-alignment.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
+gapsFasta <- read.fasta(file = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_6-alignment.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE, apply.mask = FALSE)
+#getLength(gapsFasta)
 replacingGaps <- lapply(gapsFasta, function(x) gsub("-", "N", x))
-write.fasta(replacingGaps, names(replacingGaps), "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")
+write.fasta(replacingGaps, names(replacingGaps), "../rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")
 
@@
 ```{r warning = FALSE}
-invisible({capture.output({
-                                          ### GETTING the READING FRAME     --- library "systemPipeR"
+invisible({capture.output({               ### GETTING the READING FRAME     --- library "systemPipeR"
 
-alignedFasta <- readDNAStringSet("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")
+alignedFasta <- readDNAStringSet("../rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")
 
@@
   }
-  # predORF() do not read seqs with gaps   # predicts longest ORF for sense strand in each query sequence               
-  orf1 <- predORF(alignedFasta[1:10], n=1, type="df", mode="orf", strand="sense")                                       
+  orf1 <- predORF(alignedFasta[1:10], n=1, type="df", mode="orf", strand="sense")   # predORF() do not read seqs with gaps
+  # predicts longest ORF for sense strand in each query sequence                                                                          
   longestSeq1 <- orf1$start[orf1$width == max(orf1$width)]
@@
   codonConfirmation1 <- round(codonPosition1)
-  length(alignedFasta)
   if (codonPosition1==codonConfirmation1){
-   file.rename("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta", "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")                     # 640 bp
+   file.rename("../rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta", "../rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")                     # 640 bp
   print("The Reading Frame of these seqs is +1, no further action is required, go to modelTesting")
@@
       if (codonPosition2==codonConfirmation2){
-        writeXStringSet(newRF2, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")
+        writeXStringSet(newRF2, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")
       print("The Reading Frame of these seqs was +2, new fasta file created, go to modelTesting")
@@
         newRF3 <- subseq(alignedFasta, start = 3, end = tempLength)
-        writeXStringSet(newRF3, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")    # 638 bp
+        writeXStringSet(newRF3, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")    # 638 bp
       print("The Reading Frame of these seqs was +3, new fasta file created, go to modelTesting")
@@
 
-getRF(alignedFasta)
+suppressWarnings(suppressMessages(print(getRF(alignedFasta))))
 
@@
 
-fastaRF <- read.dna(file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
+fastaRF <- read.dna(file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
+
+fastaRF1 <- read.fasta(file = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_6-alignment.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE, apply.mask = FALSE)
+getLength(fastaRF1)
 
@@
 
-bestModel <- getBestModel(fastaRF)                # getting the AICc for each competing model
+bestModel <- suppressWarnings(suppressMessages(getBestModel(fastaRF)))               # getting the AICc for each competing model
 
@@
 fit = optim.pml(fitStart, rearrangement = "stochastic", optGamma=TRUE, optInv=FALSE, model="HKY")     # seeking a maximum likelihood solution
-#set.seed(123)    # getting the same tree topology each time the code runs 
+set.seed(123)    # getting the same tree topology each time the code runs 
 bs = bootstrap.pml(fit, bs=100, optNni=TRUE, multicore=TRUE, mc.cores = 4)
@@
 # export tree with bootstrap values as node labels
-write.tree(optimalTree, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/derived_data/Sicydium_ML_tree.tre")
+write.tree(optimalTree, file="../rrcompendium/analysis/data/derived_data/Sicydium_ML_tree.tre")
 
@@
 
-(Figure 3.1)
+hshdhdhdh
 
-hshdhdhdh
+```{r demo-plot1, fig.cap="A plot of random numbers"}
+plot(rnorm(10))
+```
 
@@
 loglikelihood <- round(-1*(as.numeric(fit[1])), 2)
-loglikelihood
 
@@
 
-```{r Sicydium_ML_tree, fig.cap = paste("ML phylogenetic tree  (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus Sicydium (n=197).
-Sequences highlited in green represent samples from Cuba (n=6).
-Numbers on branches represent boostrap values higher than 70.", fig.width = 11.8, fig.height = 9.84)}
+```{r Sicydium_ML_tree, fig.cap = paste("ML phylogenetic tree  (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*.
+Sequences highlited in green represent samples from Cuba.
+Numbers on branches represent boostrap values higher than 70.")}
+
+# , fig.width = 11.8, fig.height = 10
 
 # Editing and visualizing the tree to represent --- library: ggtree
-tree <- read.tree(file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/derived_data/Sicydium_ML_tree.tre")
+tree <- read.tree(file="../rrcompendium/analysis/data/derived_data/Sicydium_ML_tree.tre")
+
+# checkinf whether the tree is rooted
+rootedTree <- is.rooted(tree)   # package ape. showing the tree root: geom_rootpoint(color="pink", size=5)
 
@@
   geom_treescale(x=0, y=-1.5, offset = -4) + 
-  geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 70), size = 3, col= "black", vjust = -0.3, hjust=1.9, na.rm = FALSE) + 
+  geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 80), size = 3, col= "black", vjust = -0.3, hjust=1.9, na.rm = FALSE) + 
   geom_nodepoint(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 70), size = 1, col= "black") + 
-  geom_cladelabel(node=218, label="Sicydium plumieri", color="red", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
-  geom_cladelabel(node=117, label="Sicydium brevifile", color="blue", hjust = -0.05, vjust = 0.3, fontsize = 3.5, fontface="italic") + 
-  geom_cladelabel(node=320, label="Sicydium salvini", color="purple", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
-  geom_cladelabel(node=328, label="Sicydium punctatum", color="brown", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
-  geom_cladelabel(node=379, label="Sicydium buscki", color="black", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
-  geom_cladelabel(node=415, label="Sicydium crenilabrum", color="orange", hjust = -0.03, vjust = 0.3, fontsize = 3.5, offset = 0.004, fontface="italic") +
-  geom_cladelabel(node=212, label="Awaous spp.", color="darkgreen", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") +
-  geom_hilight(node=377, fill="green") + 
+  geom_cladelabel(node=216, label="Sicydium plumieri", color="red", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=108, label="Sicydium brevifile", color="blue", hjust = -0.01, vjust = 0.2, fontsize = 3.5, fontface="italic") + 
+  geom_cladelabel(node=408, label="Sicydium salvini", color="purple", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=356, label="Sicydium punctatum", color="brown", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=320, label="Sicydium buscki", color="black", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=407, label="Sicydium crenilabrum", color="orange", hjust = -0.03, vjust = 0.3, fontsize = 3.5, offset = 0.004, fontface="italic") +
+  geom_cladelabel(node=211, label="Awaous spp.", color="darkgreen", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") +
+  geom_hilight(node=336, fill="green") +
+  #geom_hilight(node=332, fill="green") +
+  geom_hilight(node=135, fill="green") +
+  geom_hilight(node=121, fill="green") +
+  geom_hilight(node=120, fill="green") +
+  geom_hilight(node=139, fill="green") +
   theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")))))
 
-#Sicydium_ML_tree
 ```
 
-```{r demo-plot, fig.cap="A plot of random numbers"}
+```{r demo-plot2, fig.cap="A plot of random numbers"}
 plot(rnorm(10))

### 2020-11-27 README.Rmd
makefile in README files

@@
     illustrations
-  - [:file\_folder:
-    supplementary-materials](/analysis/supplementary-materials):
-    Supplementary materials including notes and other documents
-    prepared and collected during the analysis.
 
@@
 
+Makefile:
+Whatever the option you choose to explore this compendium, you can also use the makefile (`/analysis/makefile`) in the terminal (`make all`) to compile the .docx document and then, you can use `make clean` to remove all files generated during the process.
+
 ### Licenses

### 2020-11-30 analysis/paper/paper.Rmd
paper & bib

@@
       fig_caption: yes
-      reference_docx: '/Users/sheila_ffish/Documents/rrcompendium/analysis/templates/template.docx' # Insert path for the DOCX file
+      reference_docx: '../templates/template.docx'
       pandoc_args:
@@
 bibliography: references.bib
-csl: '/Users/sheila_ffish/Documents/rrcompendium/analysis/templates/journal-of-archaeological-science.csl' # Insert path for the bib-style
+csl: '../templates/journal-of-archaeological-science.csl'
 abstract: |
-  texting texting
+  The lack of diagnostic characters to distinguish among species and even genera of gobies (Gobiidae) is still a challenge for science in the 21st century.
+However, the use of molecular tools applied to species-level identification has brought some light into the taxonomy and systematics of these fish taxa.
+The fish fry fishery in Eastern Cuba known as “Teti” represents a local fishery and there is no information about the exploitation of this natural resource outside this region.
+Fishing “Teti” has been a common practice for several years but an exponential increase of tourism in Eastern Cuba represents an additional threat to our biodiversity.
+Besides, if “Teti” refers to a single species or a species complex is still a knowledge gap in Cuba.
+The present study aims to identify the Fish Fry "Teti" Fishery in Eastern Cuba using molecular markers.
+We conducted a barcoding analysis comparing Cuban samples of gobies with samples of the genera *Sicydium* and *Awaous* available in GenBank.
+Here, we provide evidence that Busck´s Stone-Biting Goby, *Sicydium buscki*, is a component of the fish fry “Teti” fishery in Eastern Cuba.
+However, whether this fishery includes a single species or more than one species will remain as a mystery until more studies are conducted.
+Although this study represents the first effort toward and effective management, more thorough investigations and identification of the species comprising the “Teti” fishery are needed.
+It is imperative to include them in local management plans and to avoid overfishing this resource.
 keywords: |
@@
 
-#knitr::opts_knit$set(root.dir = '../rrcompendium/')
-
-#getwd()
+#knitr::opts_knit$set(root.dir = '/rrcompendium/')
 
@@
 # INSTALLING PACKAGES AND LIBRARIES - if needed
-packages = c("raster", "ggplot2", "ggthemes",  
-             "rentrez", "ape", "seqinr", "msa", "Biostrings", "systemPipeR", "phangorn", "ggtree")
+packages = c("BiocManager", "raster", "ggplot2", "ggthemes", "here",  
+             "rentrez", "ape", "seqinr", "phangorn")
 
@@
 )
+
+packagesBioconductor = c("msa", "Biostrings", "systemPipeR", "ggtree")
+
+package.Bioconductor <- lapply(packagesBioconductor, FUN = function(y) {
+    if (!require(y, character.only = TRUE)) {
+      BiocManager::install(y, dependencies = TRUE)
+      library(y, character.only = TRUE, warn.conflicts = FALSE, quietly = TRUE)
+    }
+  }
+)
+
+#if (!requireNamespace("BiocManager", quietly = TRUE))
+  #  install.packages("BiocManager")
+
+#BiocManager::install("msa")
 
@@
 
-Tetis (19–28 mm standard length) were collected 400 m from the river mouth of Río Yumurí (Figure 2.1), eastern Cuba, in December, 2018.
+Tetis (19–28 mm standard length) were collected 400 m from the river mouth of Río Yumurí (\@ref(fig:mainMap)), eastern Cuba, in December, 2018.
 Samples were collected using hand nets (1 mm mesh size) 2 m from the riverbank.
@@
 Duplicated sequences and sequences with missing data were excluded.
-Sequences were edited using the R packages *seqinr* [@charif2007seqinr], *ape* [@paradis2015package], and *biostrings* [@pages2013package].
-Final alignment was obtained with the algorithm ClustalOmega (@sievers2011fast) using the R package *msa* [@bodenhofer2015msa].
+Sequences were edited using the R packages *seqinr* [@charif2007seqinr], *ape* [@paradis2015package], and *Biostrings* [@pages2013package].
+Final alignment was obtained with the algorithm ClustalOmega [@sievers2011fast] using the R package *msa* [@bodenhofer2015msa].
 
@@
 # storing ALL sequences in a single fasta file
-write(Sicydium_COI_seqs1, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n")
-write(Sicydium_COI_seqs2, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)   # 'append = TRUE', I can keep adding seqs
-write(Awaous_bananaCOI_seqs, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)
-write(Awaous_ocellarisCOI_seqs, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)
+write(Sicydium_COI_seqs1, here(file="analysis/data/raw_data/Sicydium_COI_1-seqs.fasta"), sep = "\n")
+# 'append = TRUE', I can keep adding seqs
+write(Sicydium_COI_seqs2, here(file="analysis/data/raw_data/Sicydium_COI_1-seqs.fasta"), sep = "\n", append = TRUE)   
+write(Awaous_bananaCOI_seqs, here(file="analysis/data/raw_data/Sicydium_COI_1-seqs.fasta"), sep = "\n", append = TRUE)
+write(Awaous_ocellarisCOI_seqs, here(file="analysis/data/raw_data/Sicydium_COI_1-seqs.fasta"), sep = "\n", append = TRUE)
 
@@
 
-Sicydium_COI_seqinrFormat <- read.fasta(file ="../rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
+Sicydium_COI_seqinrFormat <- read.fasta(here(file ="analysis/data/raw_data/Sicydium_COI_1-seqs.fasta"), seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
 
@@
 write.fasta(sequences = Sicydium_COI_seqinrFormat, names = Sicydium_COI_GenBank_Ids, nbchar = 10,
-            file.out = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_2-seqinrFormat.fasta")
+            here(file.out = "analysis/data/raw_data/Sicydium_COI_2-seqinrFormat.fasta"))
 
@@
 ### REMOVING sequences of incomplete/unspecific NAMES; that is, those with 'sp.'
-nameFasta <- read.fasta(file = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_2-seqinrFormat.fasta",
+nameFasta <- read.fasta(here(file = "analysis/data/raw_data/Sicydium_COI_2-seqinrFormat.fasta"),
                         seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
@@
 write.fasta(sequences = nameFasta[completeSeqs], names = names(nameFasta)[completeSeqs],
-            file.out = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_3-names.fasta")
+            here(file.out = "analysis/data/raw_data/Sicydium_COI_3-names.fasta"))
 
 ### REMOVING sequences of LENGTH shorter than 600 pb, given that the gene COI has between 627 and 658 bp.
-lengthFasta <- read.fasta(file = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_3-names.fasta",
+lengthFasta <- read.fasta(here(file = "analysis/data/raw_data/Sicydium_COI_3-names.fasta"),
                           seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)     # 586 seqs
@@
 write.fasta(sequences = fastaLength, names = names(fastaLength),
-            file.out = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_4-lengths.fasta")             # 401 seqs
+            here(file.out = "analysis/data/raw_data/Sicydium_COI_4-lengths.fasta"))             # 401 seqs
 
@@
 # for not using attributes to name sequences
-duplicatesFasta <- read.fasta("../rrcompendium/analysis/data/raw_data/Sicydium_COI_4-lengths.fasta", set.attributes = FALSE)
+duplicatesFasta <- read.fasta(here("analysis/data/raw_data/Sicydium_COI_4-lengths.fasta"), set.attributes = FALSE)
 
@@
 write.fasta(sequences = uniqueSeqs, names = uniqueNames,
-            file.out = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta")               # 204 seqs
+            here(file.out = "analysis/data/raw_data/Sicydium_COI_5-dupli.fasta"))               # 204 seqs
 
 # adding samples from Cuba (n=6)
-mySamples <- read.fasta("../rrcompendium/analysis/data/raw_data/Sicydium_COI_Cuba.fas")
-write.fasta(mySamples, names(mySamples),
-            "../rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta", open = "a")              # 210 seqs
+mySamples <- read.fasta(here("analysis/data/raw_data/Sicydium_COI_Cuba.fas"))
+write.fasta(mySamples, names(mySamples), here("analysis/data/raw_data/Sicydium_COI_5-dupli.fasta"), open = "a")              # 210 seqs
 
@@
 
-alignmentFasta <- readDNAStringSet("../rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta")
+alignmentFasta <- readDNAStringSet(here("analysis/data/raw_data/Sicydium_COI_5-dupli.fasta"))
 ClustalOAlignment <- msa(alignmentFasta, "ClustalOmega")
@@
 nucType <- alphabetFrequency(ClustalOAlignment)
-unknownNuc <- nucType[, c(5:14)]
+unknownNuc <- nucType[, c(5:14)]  # this indexing only contains nucleotides that are not A, G, T, C, or N or gaps
 unknownPosition <- which(unknownNuc > 0, arr.ind=T)
@@
 ClustalOAlignment2 <- as(ClustalOAlignment, "DNAStringSet")
-writeXStringSet(ClustalOAlignment2,
-                file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_6-alignment.fasta")           # 209 seqs -- 640 bp
+writeXStringSet(ClustalOAlignment2, here(file="analysis/data/raw_data/Sicydium_COI_6-alignment.fasta"))            # 209 seqs -- 640 bp
 
 ### Replacing gaps at the begining and end with Ns
-gapsFasta <- read.fasta(file = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_6-alignment.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE, apply.mask = FALSE)
+gapsFasta <- read.fasta(here(file = "analysis/data/raw_data/Sicydium_COI_6-alignment.fasta"), seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE, apply.mask = FALSE)
 #getLength(gapsFasta)
 replacingGaps <- lapply(gapsFasta, function(x) gsub("-", "N", x))
-write.fasta(replacingGaps, names(replacingGaps), "../rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")
+write.fasta(replacingGaps, names(replacingGaps), here("analysis/data/raw_data/Sicydium_COI_7-gaps.fasta"))
 
@@
 
-alignedFasta <- readDNAStringSet("../rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")
+alignedFasta <- readDNAStringSet(here("analysis/data/raw_data/Sicydium_COI_7-gaps.fasta"))
 
@@
   if (codonPosition1==codonConfirmation1){
-   file.rename("../rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta", "../rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")                     # 640 bp
+   file.rename(here("analysis/data/raw_data/Sicydium_COI_7-gaps.fasta"), here("analysis/data/raw_data/Sicydium_COI_newRF.fasta")) 
   print("The Reading Frame of these seqs is +1, no further action is required, go to modelTesting")
@@
       if (codonPosition2==codonConfirmation2){
-        writeXStringSet(newRF2, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")
+        writeXStringSet(newRF2, here(file="analysis/data/raw_data/Sicydium_COI_newRF.fasta"))                    
       print("The Reading Frame of these seqs was +2, new fasta file created, go to modelTesting")
@@
         newRF3 <- subseq(alignedFasta, start = 3, end = tempLength)
-        writeXStringSet(newRF3, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")    # 638 bp
+        writeXStringSet(newRF3, here(file="analysis/data/raw_data/Sicydium_COI_newRF.fasta"))                                
       print("The Reading Frame of these seqs was +3, new fasta file created, go to modelTesting")
@@
 
-The nucleotide substitution model that best explained our data was estimated using the function modelTest from the R package phangorn [@schliep2011phangorn].
-The phylogenetic tree was reconstructed following a maximum mikelihood approach [@felsenstein1981evolutionary] using the functions optim.pml and bootstrap.pml, also from the package phangorn.
-The tree was generated and visualized using the R package ggtree [@yu2017ggtree].
+The nucleotide substitution model that best explained our data was estimated using the function modelTest from the R package *phangorn* [@schliep2011phangorn].
+The phylogenetic tree was reconstructed following a maximum likelihood approach [@felsenstein1981evolutionary] using the functions optim.pml and bootstrap.pml, also from the package *phangorn*.
+The tree was generated and visualized using the R package *ggtree* [@yu2017ggtree].
+
+This paper was completely developed in RStudio, PBC (“RStudio”) [@rstudio2020].
+R code and data are available as a research compendium for generating this paper using the following link: https://github.com/sheilarm23/rrcompendium.git
 
@@
 
-fastaRF <- read.dna(file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")
-
-fastaRF1 <- read.fasta(file = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_6-alignment.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE, apply.mask = FALSE)
-getLength(fastaRF1)
+fastaRF <- read.dna(here(file="analysis/data/raw_data/Sicydium_COI_newRF.fasta"), format = "fasta")
 
@@
 
-```{r tree, include=TRUE, results = 'hide', fig.show='hide'}
+```{r tree, include=TRUE, results = 'hide', fig.show='hide', warning = FALSE}
 
@@
 
-# avoid to create a tree every time the code runs
-#set.seed(2015-07-01)
-
-fit = optim.pml(fitStart, rearrangement = "stochastic", optGamma=TRUE, optInv=FALSE, model="HKY")     # seeking a maximum likelihood solution
+set.seed(123)
+fit = suppressWarnings(optim.pml(fitStart, rearrangement = "stochastic", optGamma=TRUE, optInv=FALSE, model="HKY"))     # seeking a maximum likelihood solution
 set.seed(123)    # getting the same tree topology each time the code runs 
@@
 # export tree with bootstrap values as node labels
-write.tree(optimalTree, file="../rrcompendium/analysis/data/derived_data/Sicydium_ML_tree.tre")
+write.tree(optimalTree, here(file="analysis/data/derived_data/Sicydium_ML_tree.tre"))
+
+```
+
+```{r loglikelihood, include = TRUE, results='hide'}
+# this result wont be displayed in the final document
+loglikelihood <- round(-1*(as.numeric(fit[1])), 2)
+
+```
+
+```{r editingTree}
+
+# , fig.width = 11.8, fig.height = 10
+
+# Editing and visualizing the tree to represent --- library: ggtree
+tree <- read.tree(here(file="analysis/data/derived_data/Sicydium_ML_tree.tre"))
+
+# checking that the tree is rooted
+rootedTree <- is.rooted(tree)   # package ape. showing the tree root: geom_rootpoint(color="pink", size=5) 
+# getting the node number for each main clade. This info is important because those #s are the one needed to edit each main clade/species
+set.seed(123)
+nodeNumber <- ggtree(tree) + geom_text(aes(label=node), hjust=-.5, col = "blue", size = 1) # looks horrible but it gives the clade number
 
@@
 
-hshdhdhdh
-
-```{r demo-plot1, fig.cap="A plot of random numbers"}
-plot(rnorm(10))
-```
-
-```{r inline-code2, include = TRUE, results='hide'}
-# the result wont be displayed in the final document
-loglikelihood <- round(-1*(as.numeric(fit[1])), 2)
-
-```
+The final alignment comprised 209 sequences, each one of 639 base pairs.
+Figure \@ref(fig:Sicydium_ML_tree) depicts the ML tree (-lnL = `r loglikelihood`) showing the relationships among all COI sequences identified as *Sicydium* spp.
+available in GenBank.
+Six monophyletic clades matching known *Sicydium* species were recovered with strong support.
+The phylogenetic reconstruction shows that the six samples from Cuba, highlighted in green, nested within *Sicydium buscki* Evermann & Clark, 1906.
 
@@
 Numbers on branches represent boostrap values higher than 70.")}
-
-# , fig.width = 11.8, fig.height = 10
-
-# Editing and visualizing the tree to represent --- library: ggtree
-tree <- read.tree(file="../rrcompendium/analysis/data/derived_data/Sicydium_ML_tree.tre")
-
-# checkinf whether the tree is rooted
-rootedTree <- is.rooted(tree)   # package ape. showing the tree root: geom_rootpoint(color="pink", size=5)
-
-# getting the node number for each main clade. This info is important because those #s are the one needed to edit each main clade/species
-nodeNumber <- ggtree(tree) + geom_text(aes(label=node), hjust=-.5, col = "blue", size = 1) # looks horrible but it gives the clade number
 
@@
 
-```{r demo-plot2, fig.cap="A plot of random numbers"}
+```{r demo-plot, fig.cap="A plot of random numbers"}
 plot(rnorm(10))
@@
 # Acknowledgements
+
+We thank the Habitat production team for their help during the field work.
+Sequencing was partially supported by a Smithsonian Tropical Research Institution Short-term Fellowship to SRM.
+In this sense, we thank advisement provided by Harilaos Lessios and the invaluable lab assistance of Eyda Gómez, Ligia and Axel Calderón.
+Additional lab work was also done with advice of Diego Elías.
+R code was developed with help of Tad Dallas, Heru Handika, and the community of Stack Overflow.
 

### 2020-12-01 analysis/paper/paper.Rmd
almost done with the paper

@@
 
+if (!requireNamespace("BiocManager", quietly = TRUE))
+  install.packages("BiocManager")
+
 packagesBioconductor = c("msa", "Biostrings", "systemPipeR", "ggtree")
@@
 )
-
-#if (!requireNamespace("BiocManager", quietly = TRUE))
-  #  install.packages("BiocManager")
-
-#BiocManager::install("msa")
 
@@
 
-Tetis (19–28 mm standard length) were collected 400 m from the river mouth of Río Yumurí (\@ref(fig:mainMap)), eastern Cuba, in December, 2018.
+Tetis (19–28 mm standard length) were collected 400 m from the river mouth of Río Yumurí (Figure \@ref(fig:mainMap)), eastern Cuba, in December, 2018.
 Samples were collected using hand nets (1 mm mesh size) 2 m from the riverbank.
@@
 nucType <- alphabetFrequency(ClustalOAlignment)
-unknownNuc <- nucType[, c(5:14)]  # this indexing only contains nucleotides that are not A, G, T, C, or N or gaps
+unknownNuc <- nucType[, c(5:14)]  # this subset only contains nucleotides that are not A, G, T, C, or N or gaps
 unknownPosition <- which(unknownNuc > 0, arr.ind=T)
@@
   formatPhyDat <- as.phyDat(fastaRF)                #getting the class used by modelTest: phyDat or pml
+  set.seed(123)
   modelChoice <- modelTest(formatPhyDat, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 4)
@@
 # --- library: phangorn
+set.seed(123)
 env = attr(bestModel, "env")    # getting the AICc for each competing model sorted
+set.seed(123)
 fitStart = eval(get(bestModel$Model[which.min(bestModel$AICc)], env), env)    # initial fit for the model chosen
@@
 # saving the tree in the folder 'analysis/figures/'
+set.seed(123)
 bsTree <- plotBS(midpoint(fit$tree), bs, p = 75, type="p", bs.col = "blue")     # fit$tree is where the optimal tree is stored
+
+set.seed(123)
 optimalTree <- plotBS(bsTree, type = "p", cex = 0.2, font = 1)
@@
 loglikelihood <- round(-1*(as.numeric(fit[1])), 2)
+loglikelihood
 
@@
 # getting the node number for each main clade. This info is important because those #s are the one needed to edit each main clade/species
-set.seed(123)
 nodeNumber <- ggtree(tree) + geom_text(aes(label=node), hjust=-.5, col = "blue", size = 1) # looks horrible but it gives the clade number
@@
 
-```{r Sicydium_ML_tree, fig.cap = paste("ML phylogenetic tree  (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*.
+```{r Sicydium_ML_tree, fig.cap=paste("ML phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*.
 Sequences highlited in green represent samples from Cuba.

### 2020-12-01 install.R
updating binder with BiocManager

@@ -1,13 +1,15 @@
-# Installing packages for binder        # I did not include here "rmarkdown", "testthat", or any other because the project doesnt rely directly on them to run
+# Installing packages for binder        # I did not include here "rmarkdown", "testthat", or any other because the project doesn't rely directly on them to run
 install.packages("raster")
 install.packages("ggplot2")
 install.packages("ggthemes")
+install.packages("here")
 install.packages("rentrez")
 install.packages("ape")
 install.packages("seqinr")
+install.packages("phangorn")
+install.packages("BiocManager")
 install.packages("msa")
 install.packages("Biostrings")
 install.packages("systemPipeR")
-install.packages("phangorn")
 install.packages("ggtree")
 

### 2020-12-01 analysis/paper/paper.Rmd
getting ready txt and tree

@@
 
-Here is a citation [@Marwick2017]
+The family Gobiidae includes almost 2,000 species of marine, freshwater, and brackish fishes worldwide [@eschmeyer2020catalog].
+This group is mostly represented by small-sized and bottom-dweller fishes found in tropical and subtropical waters.
+Such a small size along with cryptic body colors and the lack of diagnostic characters to distinguish among species and even genera is still a challenge for science in the 21st century.
+Only in the last ten years, more than a hundred gobiid species were described [@eschmeyer2020catalog].
+However, the use of molecular tools applied to species-level identification has brought some light into the taxonomy and systematic of these fish taxa [@thacker2011phylogeny, @victor2014three, @keith2017schismatogobius].
+
+Although most species in the family Gobiidae inhabit marine ecosystems, most species inhabiting freshwater or estuarine ecosystems display an amphidromous lifecycle [@myers1949usage, @bell1999overview].
+This means that adults spawn in freshwater and when eggs hatch, they are taken downstream by river flow into the ocean.
+Then, after spending from a few weeks to a few months (depending on the species), postlarvae return to rivers.
+This moment of postlarval returning is particularly fascinating because million of them gather at river mouth as they start going upstream.
+This migratory behavior, particularly, when the postlarvae are going back to rivers, supports an artisanal fishery around the world [@bell1999overview, @mcdowall2007amphidromy].
+Known as goby-fry fisheries, this is a common practice in several countries in the Indo-Pacific [@asis2013first], the Pacific [@castellanos2011goby], and the Caribbean [@neal2009factors].
+The problem with it is that during the last years insufficient management of these fisheries has led to considerable reduction in catches [@bell1999overview].
+In addition, information of the identity of the species comprising these goby-fry fisheries in many places is still incomplete [@castellanos2011goby, @asis2013first].
+
+Gobies have been poorly studied in Cuba and available information on their identification and distribution is not conclusive [@lara2010dna].
+The fish fry fishery in Eastern Cuba known as “Teti” represents a local fishery and there is no information about the exploitation of this natural resource outside this region.
+Base on other reports in the Caribbean region [@neal2009factors], some authors claimed that the gobies *Sicydium plumieri* an *Sicydium punctatum* comprises the “Teti” fisheries in Cuba, but there is no solid evidence on that yet.
+Fishing “teti” has been a common practice for several years but an exponential increase of tourism in Eastern Cuba, due to recent economic policy changes, represents an additional threat to our biodiversity.
+Besides, if “teti” refers to a single species or a species complex is still a knowledge gap in Cuba.
+The present study aims to identify the Fish Fry "Teti" Fishery in Eastern Cuba using molecular markers.
 
@@
 alignmentFasta <- readDNAStringSet(here("analysis/data/raw_data/Sicydium_COI_5-dupli.fasta"))
+set.seed(123)
 ClustalOAlignment <- msa(alignmentFasta, "ClustalOmega")
@@
   codonPosition1 <- longestSeq1[1]/3
-  codonConfirmation1 <- round(codonPosition1)
+  codonConfirmation1 <- round(codonPosition1)                         # 640 bp
   if (codonPosition1==codonConfirmation1){
@@
         tempLength <- length(alignedFasta[[1]])
-        newRF3 <- subseq(alignedFasta, start = 3, end = tempLength)
+        newRF3 <- subseq(alignedFasta, start = 3, end = tempLength)   # 638 bp
         writeXStringSet(newRF3, here(file="analysis/data/raw_data/Sicydium_COI_newRF.fasta"))                                
@@
 This paper was completely developed in RStudio, PBC (“RStudio”) [@rstudio2020].
-R code and data are available as a research compendium for generating this paper using the following link: https://github.com/sheilarm23/rrcompendium.git
+R code and data are available as a research compendium for generating this paper using the following link: https://github.com/sheilarm23/rrcompendium.git.
 
@@
 # export tree with bootstrap values as node labels
+set.seed(123)
 write.tree(optimalTree, here(file="analysis/data/derived_data/Sicydium_ML_tree.tre"))
@@
 # getting the node number for each main clade. This info is important because those #s are the one needed to edit each main clade/species
+set.seed(123)
 nodeNumber <- ggtree(tree) + geom_text(aes(label=node), hjust=-.5, col = "blue", size = 1) # looks horrible but it gives the clade number
@@
 The final alignment comprised 209 sequences, each one of 639 base pairs.
-Figure \@ref(fig:Sicydium_ML_tree) depicts the ML tree (-lnL = `r loglikelihood`) showing the relationships among all COI sequences identified as *Sicydium* spp.
+Figure 3.1 depicts the ML tree (-lnL = `r loglikelihood`) showing the relationships among all COI sequences identified as *Sicydium* spp.
 available in GenBank.

### 2020-12-02 install.R
updating packages binder

@@ -8,8 +8,9 @@ install.packages("ape")
 install.packages("seqinr")
 install.packages("phangorn")
 install.packages("BiocManager")
-install.packages("msa")
-install.packages("Biostrings")
-install.packages("systemPipeR")
-install.packages("ggtree")
+BiocManager::install("msa")
+BiocManager::install("Biostrings")
+BiocManager::install("systemPipeR")
+BiocManager::install("ggtree")
+
 

### 2020-12-02 tests/testthat.R
checking testRF

@@ -1,8 +1,12 @@
 install.packages('knitr')
 install.packages('testthat')
+if (!requireNamespace("BiocManager", quietly = TRUE))
+  install.packages("BiocManager")
+BiocManager::install("Biostrings")
 
 library(knitr)
 library(testthat)
+library(Biostrings)
 
 knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
 

### 2020-12-02 tests/testthat/test-getBestModel.R

@@ -1,10 +1,30 @@
 
 context("test-getBestModel")
 
+SeqA <- c("ANTCATTTGA")
+SeqB <- DNAString("TTGAAA-CTC-N")
+
+test1 <- test_that("getBestModel is a function", {
+  expect_is(getBestModel, "function")
+})
+
+test2 <- test_that("getRF errors out when a file of class other than DNAStringSet is taken", {
+  expect_error(getRF(Seq1))
+})
+
+test3 <- test_that("getRF errors out when an alignment includes gaps", {
+  expect_error(getRF(Seq2))
+})
+
+
+
+
+
 test_that("getBestModel", {
   expect_is(getBestModel, "function")
 
   bestModel <- getBestModel(fastaRF)
+
   expect_true(class(bestModel) == "data.frame")    # "getBestModel outputs an object of class data.frame"
   expect_equal(nrow(bestModel), 24)  # getBestModel(fastaRF) has 24 rows based on the number of model selected"
   expect_equal(bestModel$AICc[1], min(bestModel$AICc))  # getBestModel(fastaRF) has sorted the AICc column from lowest to highest values

### 2020-12-02 tests/testthat/test-getRF.R

@@ -1,12 +1,19 @@
 
 context("test-getRF")
 
-test1 <- test_that("getRF errors out when a file of class other than DNAStringSet is taken", {
-  expect_error(class(alignedFasta)!= c("DNAStringSet"))
+Seq1 <- c("ANTCATTTGA")
+Seq2 <- DNAString("TTGAAA-CTC-N")
+
+test1 <- test_that("getRF is a function", {
+  expect_is(getBestModel, "function")
 })
 
-test2 <- test_that("getRF errors out when an alignment includes gaps", {
-  expect_error(any(letterFrequency(alignedFasta, "-", as.prob=TRUE))>0)
+test2 <- test_that("getRF errors out when a file of class other than DNAStringSet is taken", {
+  expect_error(getRF(Seq1))
+})
+
+test3 <- test_that("getRF errors out when an alignment includes gaps", {
+  expect_error(getRF(Seq2))
 })
 
 

### 2020-12-02 tests/testthat/test-getRF.R
still checking getRF

@@ -5,10 +5,10 @@ Seq1 <- c("ANTCATTTGA")
 Seq2 <- DNAString("TTGAAA-CTC-N")
 
 test1 <- test_that("getRF is a function", {
-  expect_is(getBestModel, "function")
+  expect_is(getRF, "function")
 })
 
-test2 <- test_that("getRF errors out when a file of class other than DNAStringSet is taken", {
+test2 <- test_that("getRF errors out when a file of class other than -DNAStringSet- is taken", {
   expect_error(getRF(Seq1))
 })
 

### 2020-12-02 tests/testthat/test-getRF.R
still getRF

@@ -4,7 +4,7 @@ context("test-getRF")
 Seq1 <- c("ANTCATTTGA")
 Seq2 <- DNAString("TTGAAA-CTC-N")
 
-test1 <- test_that("getRF is a function", {
+test1 <- test_that("getRF", {
   expect_is(getRF, "function")
 })
 
@@ -15,7 +15,3 @@ test2 <- test_that("getRF errors out when a file of class other than -DNAStringS
 test3 <- test_that("getRF errors out when an alignment includes gaps", {
   expect_error(getRF(Seq2))
 })
-
-
-
-

### 2020-12-02 tests/testthat/test-getBestModel.R
testing functions

@@ -2,33 +2,20 @@
 context("test-getBestModel")
 
 SeqA <- c("ANTCATTTGA")
-SeqB <- DNAString("TTGAAA-CTC-N")
-
-test1 <- test_that("getBestModel is a function", {
-  expect_is(getBestModel, "function")
-})
-
-test2 <- test_that("getRF errors out when a file of class other than DNAStringSet is taken", {
-  expect_error(getRF(Seq1))
-})
-
-test3 <- test_that("getRF errors out when an alignment includes gaps", {
-  expect_error(getRF(Seq2))
-})
-
-
-
-
+SeqB <- as.DNAbin("TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC")
 
 test_that("getBestModel", {
-  expect_is(getBestModel, "function")
 
-  bestModel <- getBestModel(fastaRF)
+  #expect_is(getBestModel, "function")  # getBestModel is a function
 
-  expect_true(class(bestModel) == "data.frame")    # "getBestModel outputs an object of class data.frame"
-  expect_equal(nrow(bestModel), 24)  # getBestModel(fastaRF) has 24 rows based on the number of model selected"
-  expect_equal(bestModel$AICc[1], min(bestModel$AICc))  # getBestModel(fastaRF) has sorted the AICc column from lowest to highest values
+  expect_error(getBestModel(SeqA))     # getBestModel errors out when a file of class other than "DNAbin" is taken
 
-  expect_error(getBestModel(fastaRF = c("ANTCATTTGA")))
+  bestModel <- getBestModel(SeqB)
+
+  expect_true(class(bestModel) == "data.frame")         # getBestModel outputs an object of class data.frame
+  expect_equal(nrow(bestModel), 24)                     # getBestModel output has 24 rows based on the number of model selected
+  expect_equal(bestModel$AICc[1], min(bestModel$AICc))  # getBestModel outputs a data.frame sorted from lowest to highest AICc values
 
 })
+
+

### 2020-12-02 tests/testthat/test-getRF.R

@@ -4,14 +4,14 @@ context("test-getRF")
 Seq1 <- c("ANTCATTTGA")
 Seq2 <- DNAString("TTGAAA-CTC-N")
 
-test1 <- test_that("getRF", {
-  expect_is(getRF, "function")
-})
+#test1 <- test_that("getRF", {
+#  expect_is(getRF, "function")
+#})
 
-test2 <- test_that("getRF errors out when a file of class other than -DNAStringSet- is taken", {
+test1 <- test_that("getRF errors out when a file of class other than -DNAStringSet- is taken", {
   expect_error(getRF(Seq1))
 })
 
-test3 <- test_that("getRF errors out when an alignment includes gaps", {
+test2 <- test_that("getRF errors out when an alignment includes gaps", {
   expect_error(getRF(Seq2))
 })

### 2020-12-02 tests/testthat.R
testing okay

@@ -1,5 +1,6 @@
 install.packages('knitr')
 install.packages('testthat')
+install.packages('ape')
 if (!requireNamespace("BiocManager", quietly = TRUE))
   install.packages("BiocManager")
 BiocManager::install("Biostrings")
@@ -7,6 +8,7 @@ BiocManager::install("Biostrings")
 library(knitr)
 library(testthat)
 library(Biostrings)
+library(ape)
 
 knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
 

### 2020-12-02 tests/testthat/test-getBestModel.R

@@ -6,7 +6,7 @@ SeqB <- as.DNAbin("TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC", "TAGT
 
 test_that("getBestModel", {
 
-  #expect_is(getBestModel, "function")  # getBestModel is a function
+  expect_is(getBestModel, "function")  # getBestModel is a function
 
   expect_error(getBestModel(SeqA))     # getBestModel errors out when a file of class other than "DNAbin" is taken
 

### 2020-12-02 tests/testthat/test-getBestModel.R
checking getBestModel

@@ -6,7 +6,7 @@ SeqB <- as.DNAbin("TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC", "TAGT
 
 test_that("getBestModel", {
 
-  expect_is(getBestModel, "function")  # getBestModel is a function
+  #expect_is(getBestModel, "function")  # getBestModel is a function
 
   expect_error(getBestModel(SeqA))     # getBestModel errors out when a file of class other than "DNAbin" is taken
 

### 2020-12-02 tests/testthat/test-getBestModel.R
still getBestModel

@@ -2,20 +2,20 @@
 context("test-getBestModel")
 
 SeqA <- c("ANTCATTTGA")
-SeqB <- as.DNAbin("TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC")
+SeqB <- as.DNAbin(c("TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC"))
 
-test_that("getBestModel", {
-
-  #expect_is(getBestModel, "function")  # getBestModel is a function
-
-  expect_error(getBestModel(SeqA))     # getBestModel errors out when a file of class other than "DNAbin" is taken
-
-  bestModel <- getBestModel(SeqB)
-
-  expect_true(class(bestModel) == "data.frame")         # getBestModel outputs an object of class data.frame
-  expect_equal(nrow(bestModel), 24)                     # getBestModel output has 24 rows based on the number of model selected
-  expect_equal(bestModel$AICc[1], min(bestModel$AICc))  # getBestModel outputs a data.frame sorted from lowest to highest AICc values
-
+test1 <- test_that("getBestModel errors out when a file of class other than -DNAStringSet- is taken", {
+  expect_error(getBestModel(SeqA))
 })
 
+test2 <- test_that("getBestModel outputs an object of class data.frame", {
+  expect_true(class(getBestModel(SeqB)) == "data.frame")
+})
 
+test3 <- test_that("getBestModel output has 24 rows based on the number of model selected", {
+  expect_equal(nrow(getBestModel(SeqB)), 24)
+})
+
+test4 <- test_that("getBestModel outputs a data.frame sorted from lowest to highest AICc values", {
+  expect_equal(getBestModel(SeqB)$AICc[1], min(getBestModel(SeqB)$AICc))
+})

### 2020-12-02 tests/testthat/test-getBestModel.R
still travis

@@ -4,18 +4,22 @@ context("test-getBestModel")
 SeqA <- c("ANTCATTTGA")
 SeqB <- as.DNAbin(c("TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC"))
 
-test1 <- test_that("getBestModel errors out when a file of class other than -DNAStringSet- is taken", {
-  expect_error(getBestModel(SeqA))
+test1 <- test_that("getBestModel", {
+  expect_is(getBestModel, "function")
 })
 
-test2 <- test_that("getBestModel outputs an object of class data.frame", {
-  expect_true(class(getBestModel(SeqB)) == "data.frame")
+test2 <- test_that("getBestModel", {
+  expect_error(getBestModel(SeqA)) # errors out when a file of class other than -DNAStringSet- is taken
 })
 
-test3 <- test_that("getBestModel output has 24 rows based on the number of model selected", {
-  expect_equal(nrow(getBestModel(SeqB)), 24)
+test3 <- test_that("getBestModel", {
+  expect_true(class(getBestModel(SeqB)) == "data.frame")  # outputs an object of class data.frame
 })
 
-test4 <- test_that("getBestModel outputs a data.frame sorted from lowest to highest AICc values", {
-  expect_equal(getBestModel(SeqB)$AICc[1], min(getBestModel(SeqB)$AICc))
+test4 <- test_that("getBestModel", {
+  expect_equal(nrow(getBestModel(SeqB)), 24)   # output has 24 rows based on the number of model selected
+})
+
+test5 <- test_that("getBestModel", {
+  expect_equal(getBestModel(SeqB)$AICc[1], min(getBestModel(SeqB)$AICc))  # outputs a data.frame sorted from lowest to highest AICc values
 })

### 2020-12-02 tests/testthat/test-getRF.R

@@ -4,14 +4,14 @@ context("test-getRF")
 Seq1 <- c("ANTCATTTGA")
 Seq2 <- DNAString("TTGAAA-CTC-N")
 
-#test1 <- test_that("getRF", {
-#  expect_is(getRF, "function")
-#})
-
-test1 <- test_that("getRF errors out when a file of class other than -DNAStringSet- is taken", {
-  expect_error(getRF(Seq1))
+test1 <- test_that("getRF", {
+  expect_is(getRF, "function")
 })
 
-test2 <- test_that("getRF errors out when an alignment includes gaps", {
-  expect_error(getRF(Seq2))
+test1 <- test_that("getRF", {
+  expect_error(getRF(Seq1))   # errors out when a file of class other than -DNAStringSet- is taken
+})
+
+test2 <- test_that("getRF", {
+  expect_error(getRF(Seq2))   #  errors out when an alignment includes gaps
 })

### 2020-12-02 analysis/paper/paper.Rmd
functions name chunks

@@
 
-```{r warning = FALSE}
+```{r, getRF, warning = FALSE}
 invisible({capture.output({               ### GETTING the READING FRAME     --- library "systemPipeR"
@@
 
-```{r bestModel, warning = FALSE}
+```{r getBestModel, warning = FALSE}
                                       # FINDING THE BEST NUCLEOTIDE SUSBSTITUTION MODEL

### 2020-12-02 tests/testthat.R
still travis

@@ -1,14 +1,18 @@
 install.packages('knitr')
 install.packages('testthat')
 install.packages('ape')
+install.packages('phangorn')
 if (!requireNamespace("BiocManager", quietly = TRUE))
   install.packages("BiocManager")
 BiocManager::install("Biostrings")
+BiocManager::install("systemPipeR")
 
 library(knitr)
 library(testthat)
 library(Biostrings)
 library(ape)
+library(phangorn)
+library(systemPipeR)
 
 knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
 

### 2020-12-02 analysis/paper/paper.Rmd
updating functions arguments

@@
 
-getRF <- function(alignedFasta){
-  if (class(alignedFasta)!= c("DNAStringSet")){
+getRF <- function(x){
+  if (class(x)!= c("DNAStringSet")){
     stop("This function only takes files of class 'DNAStringSet'. Use package 'Biostrings' to read the input file")
   }
-  if (any(letterFrequency(alignedFasta, "-", as.prob=TRUE))>0){
+  if (any(letterFrequency(x, "-", as.prob=TRUE))>0){
     stop("This function do not take gaps as values/strings.
@@
   }
-  orf1 <- predORF(alignedFasta[1:10], n=1, type="df", mode="orf", strand="sense")   # predORF() do not read seqs with gaps
+  orf1 <- predORF(x[1:10], n=1, type="df", mode="orf", strand="sense")   # predORF() do not read seqs with gaps
   # predicts longest ORF for sense strand in each query sequence                                                                          
@@
     if (codonPosition1!=codonConfirmation1){
-      tempLength <- length(alignedFasta[[1]])
+      tempLength <- length(x[[1]])
       #remove first nucleotide position
-      newRF2 <- subseq(alignedFasta, start = 2, end = tempLength)      # 639 bp
+      newRF2 <- subseq(x, start = 2, end = tempLength)      # 639 bp
       # getting new RF from the new alignment with -1 bp.
@@
         #remove first nucleotide position
-        tempLength <- length(alignedFasta[[1]])
-        newRF3 <- subseq(alignedFasta, start = 3, end = tempLength)   # 638 bp
+        tempLength <- length(x[[1]])
+        newRF3 <- subseq(x, start = 3, end = tempLength)   # 638 bp
         writeXStringSet(newRF3, here(file="analysis/data/raw_data/Sicydium_COI_newRF.fasta"))                                
@@
 
-suppressWarnings(suppressMessages(print(getRF(alignedFasta))))
+suppressWarnings(suppressMessages(print(getRF(x = alignedFasta))))
 
@@
 
-getBestModel <- function(fastaRF){
-  if (class(fastaRF)!= c("DNAbin")){
+getBestModel <- function(y){
+  if (class(y)!= c("DNAbin")){
     stop("This function only takes files of class 'DNAbin'. Use package 'ape' to read the input file")
   }
-  formatPhyDat <- as.phyDat(fastaRF)                #getting the class used by modelTest: phyDat or pml
+  formatPhyDat <- as.phyDat(y)                #getting the class used by modelTest: phyDat or pml
   set.seed(123)
@@
 
-bestModel <- suppressWarnings(suppressMessages(getBestModel(fastaRF)))               # getting the AICc for each competing model
+bestModel <- suppressWarnings(suppressMessages(getBestModel(y = fastaRF)))               # getting the AICc for each competing model
 
@@
 
-```{r tree, include=TRUE, results = 'hide', fig.show='hide', warning = FALSE}
+```{r tree, echo = FALSE, include=TRUE, results = 'hide', fig.show='hide', warning = FALSE}
 
@@
 
-```{r editingTree}
+```{r editingTree, echo = FALSE}
 

### 2020-12-02 analysis/templates/template.Rmd
updating functions arguments

---
title: "My title"
author:
- author 1
- other author
date: '`r Sys.Date()`'
abstract: "My abstract \n"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Heading 1

## Heading 2

### Heading 3

#### Heading 4

##### Heading 5 - this is for the page break with MS Word

In the MS Word document, the following modifications must be made in the 'Modify Style' menu:

+ Set the font color to ‘white’ (rather than ‘Automatic’).
+ Select the smallest font size (8 rather than 11).
+ Select ‘Page break before’ in the ‘Line and Page Breaks’ tab.
+ Set the line spacing to ‘Exactly’ and ‘1 pt’ in the ‘Indents and Spacing’ tab.

After these tweaks, the ‘Heading 5’ style will no longer format a heading of level 5. Instead it will insert a very small and white (and, thus, invisible) line followed by a page break.

Now knit this doc, and edit the styles in the resulting docx file. Then rename as 'template' and use this in the yml:

```
output:
  word_document:
    reference_docx: templates/template.docx
```

Don't forget line numbers for JAS! Headings are bold, underline, nothing. And Arial Narrow 10pt for tables.

See here for more information: http://rmarkdown.rstudio.com/articles_docx.html

### 2020-12-02 R/getBestModel.R

@@ -1,15 +1,15 @@
 #' getBestModel
 #' This function provides a dataframe with the list of models of nucleotide evolution sorted by AICc (from lowest to highest)
 #'
-#' @param fastaRF takes as input a dataset of aligned sequences of the same length and known ORF
+#' @param y takes as input a dataset of aligned sequences of the same length and known ORF
 #' @return returns a dataframe with the list of models of evolution sorted by AICc, from the lowest to the highest AICc value
 #'
 #'
-getBestModel <- function(fastaRF){
-  if (class(fastaRF)!= c("DNAbin")){
+getBestModel <- function(y){
+  if (class(y)!= c("DNAbin")){
     stop("This function only takes files of class 'DNAbin'. Use package 'ape' to read the input file")
   }
-  formatPhyDat <- as.phyDat(fastaRF)                #getting the class used by modelTest: phyDat or pml
+  formatPhyDat <- as.phyDat(y)                #getting the class used by modelTest: phyDat or pml
   modelChoice <- modelTest(formatPhyDat, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 4)
   modelChoiceSorted <- modelChoice[order(modelChoice$AICc),]
   return(modelChoiceSorted)

### 2020-12-02 R/getRF.R

@@ -1,32 +1,32 @@
 #' getRF
 #' This function determines the Reading Frame (RF) of a dataset of nucleotide sequences
 #'
-#' @param alignedFasta takes as input a dataset of aligned sequences of the same length
+#' @param x takes as input a dataset of aligned sequences of the same length
 #'
 #' As output, this function either renames the input file if the sequences are in the ORF 1, or it writes a new fasta file if the input file was in ORF 2 or 3.
 #' The output file will be of RF 1
 #'
-getRF <- function(alignedFasta){
-  if (class(alignedFasta)!= c("DNAStringSet")){
+getRF <- function(x){
+  if (class(x)!= c("DNAStringSet")){
     stop("This function only takes files of class 'DNAStringSet'. Use package 'Biostrings' to read the input file")
   }
-  if (any(letterFrequency(alignedFasta, "-", as.prob=TRUE))>0){
+  if (any(letterFrequency(x, "-", as.prob=TRUE))>0){
     stop("This function do not take gaps as values/strings. Go and replaced gaps with Ns (if the gaps are at the begining or end of the sequences) or choose a subset of sequences that do not contain gaps")
   }
   # predORF() do not read seqs with gaps   # predicts longest ORF for sense strand in each query sequence
-  orf1 <- predORF(alignedFasta[1:10], n=1, type="df", mode="orf", strand="sense")
+  orf1 <- predORF(x[1:10], n=1, type="df", mode="orf", strand="sense")
   longestSeq1 <- orf1$start[orf1$width == max(orf1$width)]
   codonPosition1 <- longestSeq1[1]/3
   codonConfirmation1 <- round(codonPosition1)
-  length(alignedFasta)
+  length(x)
   if (codonPosition1==codonConfirmation1){
     file.rename("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta", "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")                     # 640 bp
     print("The Reading Frame of these seqs is +1, no further action is required, go to modelTesting")
   } else {
     if (codonPosition1!=codonConfirmation1){
-      tempLength <- length(alignedFasta[[1]])
+      tempLength <- length(x[[1]])
       #remove first nucleotide position
-      newRF2 <- subseq(alignedFasta, start = 2, end = tempLength)      # 639 bp
[…]

### 2020-12-02 tests/testthat/test-getBestModel.R

@@ -12,14 +12,16 @@ test2 <- test_that("getBestModel", {
   expect_error(getBestModel(SeqA)) # errors out when a file of class other than -DNAStringSet- is taken
 })
 
+bestModel <- getBestModel(SeqB)
+
 test3 <- test_that("getBestModel", {
-  expect_true(class(getBestModel(SeqB)) == "data.frame")  # outputs an object of class data.frame
+  expect_true(class(bestModel) == "data.frame")  # outputs an object of class data.frame
 })
 
 test4 <- test_that("getBestModel", {
-  expect_equal(nrow(getBestModel(SeqB)), 24)   # output has 24 rows based on the number of model selected
+  expect_equal(nrow(bestModel), 24)   # output has 24 rows based on the number of model selected
 })
 
 test5 <- test_that("getBestModel", {
-  expect_equal(getBestModel(SeqB)$AICc[1], min(getBestModel(SeqB)$AICc))  # outputs a data.frame sorted from lowest to highest AICc values
+  expect_equal(bestModel$AICc[1], min(bestModel$AICc))  # outputs a data.frame sorted from lowest to highest AICc values
 })

### 2020-12-02 tests/testthat/test-getRF.R

@@ -8,10 +8,10 @@ test1 <- test_that("getRF", {
   expect_is(getRF, "function")
 })
 
-test1 <- test_that("getRF", {
+test2 <- test_that("getRF", {
   expect_error(getRF(Seq1))   # errors out when a file of class other than -DNAStringSet- is taken
 })
 
-test2 <- test_that("getRF", {
+test3 <- test_that("getRF", {
   expect_error(getRF(Seq2))   #  errors out when an alignment includes gaps
 })

### 2020-12-02 tests/testthat.R
functions again

@@ -14,7 +14,10 @@ library(ape)
 library(phangorn)
 library(systemPipeR)
 
-knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
+#knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
+
+knitr::knit('R/getBestModel.R', tangle=TRUE)
+knitr::knit('R/getRF.R', tangle=TRUE)
 
 testthat::test_file('tests/testthat/test-getRF.R')
 testthat::test_file('tests/testthat/test-getBestModel.R')

### 2020-12-02 README.Rmd
updating readme file
updating licenses

@@
 
-> Rodriguez-Machado, S., Rodríguez-Cabrera, T.M., and Chakrabarty, P.
-(2020).
-_The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba_.
-Name of journal/book <https://doi.org/{DOI-to-paper}>
+> Rodriguez-Machado, S. and Rodríguez-Cabrera, T.M. (2020). _The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba_. Name of journal/book <https://doi.org/{DOI-to-paper}>
 
@@
 
-> Rodriguez-Machado, S., Rodríguez-Cabrera, T.M., and Chakrabarty, P.
+> Rodriguez-Machado, S.
+and Rodríguez-Cabrera, T.M.
 (2020).
@@
     illustrations
+  - [:file\_folder: templates](/analysis/templates): Style of citations and reference list
 

### 2020-12-02 analysis/paper/paper.Rmd
updating paper

@@
 Only in the last ten years, more than a hundred gobiid species were described [@eschmeyer2020catalog].
-However, the use of molecular tools applied to species-level identification has brought some light into the taxonomy and systematic of these fish taxa [@thacker2011phylogeny, @victor2014three, @keith2017schismatogobius].
+However, the use of molecular tools applied to species-level identification has brought some light into the taxonomy and systematic of these fish taxa [@thacker2011phylogeny; @victor2014three; @keith2017schismatogobius].
 
-Although most species in the family Gobiidae inhabit marine ecosystems, most species inhabiting freshwater or estuarine ecosystems display an amphidromous lifecycle [@myers1949usage, @bell1999overview].
+Although most species in the family Gobiidae inhabit marine ecosystems, most species inhabiting freshwater or estuarine ecosystems display an amphidromous lifecycle [@myers1949usage; @bell1999overview].
 This means that adults spawn in freshwater and when eggs hatch, they are taken downstream by river flow into the ocean.
@@
 This moment of postlarval returning is particularly fascinating because million of them gather at river mouth as they start going upstream.
-This migratory behavior, particularly, when the postlarvae are going back to rivers, supports an artisanal fishery around the world [@bell1999overview, @mcdowall2007amphidromy].
+This migratory behavior, particularly, when the postlarvae are going back to rivers, supports an artisanal fishery around the world [@bell1999overview; @mcdowall2007amphidromy].
 Known as goby-fry fisheries, this is a common practice in several countries in the Indo-Pacific [@asis2013first], the Pacific [@castellanos2011goby], and the Caribbean [@neal2009factors].
 The problem with it is that during the last years insufficient management of these fisheries has led to considerable reduction in catches [@bell1999overview].
-In addition, information of the identity of the species comprising these goby-fry fisheries in many places is still incomplete [@castellanos2011goby, @asis2013first].
+In addition, information of the identity of the species comprising these goby-fry fisheries in many places is still incomplete [@castellanos2011goby; @asis2013first].
 
@@
 This paper was completely developed in RStudio, PBC (“RStudio”) [@rstudio2020].
-R code and data are available as a research compendium for generating this paper using the following link: https://github.com/sheilarm23/rrcompendium.git.
+R code and data are available as a research compendium for generating this paper in the following link: https://github.com/sheilarm23/rrcompendium.git.
 
@@
 
-# , fig.width = 11.8, fig.height = 10
-
 # Editing and visualizing the tree to represent --- library: ggtree
@@
 nodeNumber <- ggtree(tree) + geom_text(aes(label=node), hjust=-.5, col = "blue", size = 1) # looks horrible but it gives the clade number
+nodeNumber
 
@@
 
-```{r Sicydium_ML_tree, fig.cap=paste("ML phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*.
+```{r SicydiumTree, fig.cap=paste("ML phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*.
 Sequences highlited in green represent samples from Cuba.
@@
 
-Sicydium_ML_tree <-suppressWarnings(suppressMessages(print(ggtree(tree, size = .4) + 
+# , fig.width = 11.8, fig.height = 10
+
+ggtree(tree, size = .4) + 
   xlim(-0.02, 0.26) +
@@
   geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 80), size = 3, col= "black", vjust = -0.3, hjust=1.9, na.rm = FALSE) + 
-  geom_nodepoint(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 70), size = 1, col= "black") + 
-  geom_cladelabel(node=216, label="Sicydium plumieri", color="red", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
+  geom_nodepoint(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 80), size = 1, col= "black") + 
+  geom_cladelabel(node=215, label="Sicydium plumieri", color="red", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
   geom_cladelabel(node=108, label="Sicydium brevifile", color="blue", hjust = -0.01, vjust = 0.2, fontsize = 3.5, fontface="italic") + 
   geom_cladelabel(node=408, label="Sicydium salvini", color="purple", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
-  geom_cladelabel(node=356, label="Sicydium punctatum", color="brown", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
-  geom_cladelabel(node=320, label="Sicydium buscki", color="black", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=320, label="Sicydium punctatum", color="brown", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=371, label="Sicydium buscki", color="black", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
   geom_cladelabel(node=407, label="Sicydium crenilabrum", color="orange", hjust = -0.03, vjust = 0.3, fontsize = 3.5, offset = 0.004, fontface="italic") +
   geom_cladelabel(node=211, label="Awaous spp.", color="darkgreen", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") +
-  geom_hilight(node=336, fill="green") +
-  #geom_hilight(node=332, fill="green") +
-  geom_hilight(node=135, fill="green") +
-  geom_hilight(node=121, fill="green") +
-  geom_hilight(node=120, fill="green") +
-  geom_hilight(node=139, fill="green") +
-  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")))))
+  # samples Sicydium_COI_Cuba2,4,6 were nested within this clde but the distance is too short to be displayed in the tree
+  geom_hilight(node=191, fill="green") +
+  geom_hilight(node=166, fill="green") +
+  geom_hilight(node=177, fill="green") +
+  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))
+
+# suppressWarnings(suppressMessages(print(Sicydium_ML_tree)))
 

### 2020-12-02 analysis/paper/paper.Rmd
almost done

@@
 nodeNumber <- ggtree(tree) + geom_text(aes(label=node), hjust=-.5, col = "blue", size = 1) # looks horrible but it gives the clade number
-nodeNumber
 
@@
 The final alignment comprised 209 sequences, each one of 639 base pairs.
-Figure 3.1 depicts the ML tree (-lnL = `r loglikelihood`) showing the relationships among all COI sequences identified as *Sicydium* spp.
+Figure \@ref(fig:SicydiumTree) depicts the ML tree (-lnL = `r loglikelihood`) showing the relationships among all COI sequences identified as *Sicydium* spp.
 available in GenBank.
@@
 
-```{r SicydiumTree, fig.cap=paste("ML phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*.
+```{r plot2, fig.width = 15, fig.height = 19, fig.cap=paste("ML phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*.
 Sequences highlited in green represent samples from Cuba.
@@
 
-# , fig.width = 11.8, fig.height = 10
-
-ggtree(tree, size = .4) + 
+ggtree(tree, size = .8) + 
   xlim(-0.02, 0.26) +
-  geom_treescale(x=0, y=-1.5, offset = -4) + 
-  geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 80), size = 3, col= "black", vjust = -0.3, hjust=1.9, na.rm = FALSE) + 
-  geom_nodepoint(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 80), size = 1, col= "black") + 
-  geom_cladelabel(node=215, label="Sicydium plumieri", color="red", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
-  geom_cladelabel(node=108, label="Sicydium brevifile", color="blue", hjust = -0.01, vjust = 0.2, fontsize = 3.5, fontface="italic") + 
-  geom_cladelabel(node=408, label="Sicydium salvini", color="purple", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
-  geom_cladelabel(node=320, label="Sicydium punctatum", color="brown", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
-  geom_cladelabel(node=371, label="Sicydium buscki", color="black", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
-  geom_cladelabel(node=407, label="Sicydium crenilabrum", color="orange", hjust = -0.03, vjust = 0.3, fontsize = 3.5, offset = 0.004, fontface="italic") +
-  geom_cladelabel(node=211, label="Awaous spp.", color="darkgreen", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") +
+  geom_treescale(x=0, y=-1.5, offset = -4, fontsize = 6, linesize = 0.8) + 
+  geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 82), size = 6, col= "black", vjust = -0.3, hjust=1.9, na.rm = FALSE) + 
+  geom_nodepoint(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 82), size = 2, col= "black") + 
+  geom_cladelabel(node=215, label="Sicydium plumieri", color="red", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=108, label="Sicydium brevifile", color="blue", hjust = -0.008, vjust = 0.8, fontsize = 6, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=408, label="Sicydium salvini", color="purple", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=320, label="Sicydium punctatum", color="brown", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=371, label="Sicydium buscki", color="black", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=407, label="Sicydium crenilabrum", color="orange", hjust = -0.03, vjust = 0.1, fontsize = 6, offset = 0.004, fontface="italic") +
+  geom_cladelabel(node=211, label="Awaous spp.", color="darkgreen", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") +
   # samples Sicydium_COI_Cuba2,4,6 were nested within this clde but the distance is too short to be displayed in the tree
-  geom_hilight(node=191, fill="green") +
-  geom_hilight(node=166, fill="green") +
-  geom_hilight(node=177, fill="green") +
+  geom_hilight(node=191, fill="green", alpha = 0.7, extend=.004) +
+  geom_hilight(node=166, fill="green", alpha = 0.7, extend=.004) +
+  geom_hilight(node=177, fill="green", alpha = 0.7, extend=.004) +
+  geom_hilight(node=173, fill="green", alpha = 0.7, extend=.004) +
+  geom_hilight(node=171, fill="green", alpha = 0.7, extend=.004) +
+  geom_hilight(node=172, fill="green", alpha = 0.7, extend=.004) +
   theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))

### 2020-12-02 analysis/templates/template.Rmd
almost done

@@
 date: '`r Sys.Date()`'
+institute:
+  - UofO: [UofO]
+  - UofA: [UofA]
 abstract: "My abstract \n"

### 2020-12-02 tests/testthat.R

@@ -14,10 +14,7 @@ library(ape)
 library(phangorn)
 library(systemPipeR)
 
-#knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
-
-knitr::knit('R/getBestModel.R', tangle=TRUE)
-knitr::knit('R/getRF.R', tangle=TRUE)
+knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
 
 testthat::test_file('tests/testthat/test-getRF.R')
 testthat::test_file('tests/testthat/test-getBestModel.R')

### 2020-12-03 README.Rmd
updating master.zip

@@
 )
-# Please put your title here to include it in the file below.
+
 Title <- 'The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba'
@@
 You can download the compendium as a zip from from this URL:
-[master.zip](/archive/master.zip). After unzipping:
+[master.zip](https://github.com/sheilarm23/rrcompendium/archive/master.zip). After unzipping:
 - open the `.Rproj` file in RStudio

### 2020-12-03 analysis/paper/paper.Rmd
updating master.zip

@@
 # INSTALLING PACKAGES AND LIBRARIES - if needed
-packages = c("BiocManager", "raster", "ggplot2", "ggthemes", "here",  
+packages = c("BiocManager", "raster", "ggplot2", "ggthemes", "mapproj", "here",  
              "rentrez", "ape", "seqinr", "phangorn")

### 2020-12-03 install.R
adding mapproj

@@ -2,6 +2,7 @@
 install.packages("raster")
 install.packages("ggplot2")
 install.packages("ggthemes")
+install.packages("mapproj")
 install.packages("here")
 install.packages("rentrez")
 install.packages("ape")

### 2020-12-03 analysis/paper/paper.Rmd
text and bib done

@@
 keywords: |
-  amphidromy; Busck's Stone-Biting Goby; COI; Gobiidae
+  amphidromy; Busck's Stone-Biting Goby; COI; DNA barcoding; Gobiidae
 
@@
 Only in the last ten years, more than a hundred gobiid species were described [@eschmeyer2020catalog].
-However, the use of molecular tools applied to species-level identification has brought some light into the taxonomy and systematic of these fish taxa [@thacker2011phylogeny; @victor2014three; @keith2017schismatogobius].
+However, the use of molecular tools applied to species-level identification has brought some light into the taxonomy and systematics of these fish taxa [@thacker2011phylogeny; @victor2014three; @keith2017schismatogobius].
 
 Although most species in the family Gobiidae inhabit marine ecosystems, most species inhabiting freshwater or estuarine ecosystems display an amphidromous lifecycle [@myers1949usage; @bell1999overview].
-This means that adults spawn in freshwater and when eggs hatch, they are taken downstream by river flow into the ocean.
+This means that adults spawn in freshwater and when the eggs hatch, they are taken downstream by river flow into the ocean.
 Then, after spending from a few weeks to a few months (depending on the species), postlarvae return to rivers.
-This moment of postlarval returning is particularly fascinating because million of them gather at river mouth as they start going upstream.
+This moment of postlarval returning is particularly fascinating because million of them gather at the river mouth as they start going upstream.
 This migratory behavior, particularly, when the postlarvae are going back to rivers, supports an artisanal fishery around the world [@bell1999overview; @mcdowall2007amphidromy].
 Known as goby-fry fisheries, this is a common practice in several countries in the Indo-Pacific [@asis2013first], the Pacific [@castellanos2011goby], and the Caribbean [@neal2009factors].
-The problem with it is that during the last years insufficient management of these fisheries has led to considerable reduction in catches [@bell1999overview].
-In addition, information of the identity of the species comprising these goby-fry fisheries in many places is still incomplete [@castellanos2011goby; @asis2013first].
+The problem with it is that during the last years insufficient management of these fisheries has led to a considerable reduction in catches [@bell1999overview].
+In addition, information on the identity of the species comprising these goby-fry fisheries in many places is still incomplete [@castellanos2011goby; @asis2013first].
 
@@
 The fish fry fishery in Eastern Cuba known as “Teti” represents a local fishery and there is no information about the exploitation of this natural resource outside this region.
-Base on other reports in the Caribbean region [@neal2009factors], some authors claimed that the gobies *Sicydium plumieri* an *Sicydium punctatum* comprises the “Teti” fisheries in Cuba, but there is no solid evidence on that yet.
+Base on other reports in the Caribbean region [@neal2009factors], some authors claimed that the gobies *Sicydium plumieri* and *Sicydium punctatum* comprise the “Teti” fisheries in Cuba, but there is no solid evidence on that yet.
 Fishing “teti” has been a common practice for several years but an exponential increase of tourism in Eastern Cuba, due to recent economic policy changes, represents an additional threat to our biodiversity.
@@
 Raw sequence files were edited and aligned with the software Geneious v.9.1 [@kearse2012geneious], and submitted to GenBank database (accession numbers #####).
-We used sequences of the genus *Sicydium* as reference for recovering the taxonomic identity of the "tetis".
+We used sequences of the genus *Sicydium* as reference for recovering the taxonomic identity of the "Tetis".
 Sequences of *Awaous banana* and *A.
@@
 The phylogenetic tree was reconstructed following a maximum likelihood approach [@felsenstein1981evolutionary] using the functions optim.pml and bootstrap.pml, also from the package *phangorn*.
-The tree was generated and visualized using the R package *ggtree* [@yu2017ggtree].
+The tree was edited and visualized using the R package *ggtree* [@yu2017ggtree].
 
@@
 available in GenBank.
-Six monophyletic clades matching known *Sicydium* species were recovered with strong support.
+Six clades matching known *Sicydium* species were recovered with strong support.
 The phylogenetic reconstruction shows that the six samples from Cuba, highlighted in green, nested within *Sicydium buscki* Evermann & Clark, 1906.
 
-```{r plot2, fig.width = 15, fig.height = 19, fig.cap=paste("ML phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*.
+```{r SicydiumTree, fig.width = 15, fig.height = 19, fig.cap=paste("ML phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*.
 Sequences highlited in green represent samples from Cuba.
-Numbers on branches represent boostrap values higher than 70.")}
+Numbers on branches represent boostrap values higher than 82.")}
 
@@
 
-```{r demo-plot, fig.cap="A plot of random numbers"}
-plot(rnorm(10))
-```
+# Discussion
 
-Figure \@ref(fig:demo-plot) shows how we can have a caption and cross-reference for a plot
+When it comes to nature conservation, species identification is a basic principle: what cannot be identified, will not be protected or conserved.
+Any successful management plan or effort will critically rely on a correct identification of the taxa intended to protect.
+However, when taxonomic issues exist, molecular tools have proved to be of considerable help uncovering largely hidden diversity [*e.g.*, @ardura2010dna; @kress2015dna].
+Cuban freshwater fishes are an excellent example of the importance of applying molecular tools to species conservation, where more than ten new lineages have been discovered in the last years [@lara2010dna; @garcia2011molecular; @de2014phylogeography].
+Nevertheless, some groups remain poorly studied, which prevents us from implementing appropriate conservation programs.
 
-```{r demo-inline-code}
-x <- round(pi, 2)
-```
+Freshwater gobies (Gobiiformes, Gobiidae) have been a neglected group of fishes in Cuba.
+In fact, most information about their diversity and ecology on the island is virtually unknown.
+What we do know is that given their migratory life cycle, they have been the target of an artisanal fishery based on the postlarvae returning to rivers.
+This fish fry fishery, known as “Teti” fishery, is only practiced in the easternmost part of the island, and it has been reported from four rivers: Duaba, Yumuri, Toa, and Miel.
+The main problem with "Tetis", as it has been mentioned in many other countries harvesting postlarvae, is that species (and genus)-level identification is almost impossible in the field.
 
-Here is an example of inline code `r x` in the middle of a sentence.
-
-# Discussion
+Our results suggest that the “Teti” samples collected are *Sicydium buscki*.
+This is one of the four *Sicydium* species reported in the Caribbean islands [@eschmeyer2020catalog].
+Although other species have been suggested to comprise this fishery in Cuba, this is the first report of any fish species based on molecular data.
+However, this study included a few samples from a single river.
+Other countries where fish fry fisheries also take place have reported several species of the order Gobiiformes [@castellanos2011goby; @asis2013first; @engman2019puerto].
+Thus, additional surveys including more samples and rivers are needed to have a better idea, first of the diversity of gobies in Eastern Cuba, and second, of the fish diversity comprising the fish fry “Teti” fishery.
+Although harvest mainly occurs during the last-quarter moon phase from August to October [@tetiblogger2011], an increase in tourism in Cuba is leading to a higher demand for this natural resource.
+For this reason, further studies will contribute to our knowledge of Caribbean fish assemblages, which will allow us to better work toward their conservation and propose and develop effective management actions.
 

### 2020-12-03 tests/testthat.R
travis again

@@ -1,20 +1,21 @@
 install.packages('knitr')
 install.packages('testthat')
 install.packages('ape')
-install.packages('phangorn')
+#install.packages('phangorn')
 if (!requireNamespace("BiocManager", quietly = TRUE))
   install.packages("BiocManager")
 BiocManager::install("Biostrings")
-BiocManager::install("systemPipeR")
+#BiocManager::install("systemPipeR")
 
 library(knitr)
 library(testthat)
 library(Biostrings)
 library(ape)
-library(phangorn)
-library(systemPipeR)
+#library(phangorn)
+#library(systemPipeR)
 
 knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
+source('analysis/paper/paper.R')
 
 testthat::test_file('tests/testthat/test-getRF.R')
 testthat::test_file('tests/testthat/test-getBestModel.R')

### 2020-12-03 tests/testthat.R
travissss

@@ -15,7 +15,7 @@ library(ape)
 #library(systemPipeR)
 
 knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
-source('analysis/paper/paper.R')
+source('paper.R')
 
 testthat::test_file('tests/testthat/test-getRF.R')
 testthat::test_file('tests/testthat/test-getBestModel.R')

### 2020-12-04 R/myFunctions.R
updating files for travis

@@ -1,0 +1,69 @@
+#' getRF
+#' This function determines the Reading Frame (RF) of a dataset of nucleotide sequences
+#'
+#' @param x takes as input a dataset of aligned sequences of the same length
+#'
+#' As output, this function either renames the input file if the sequences are in the ORF 1, or it writes a new fasta file if the input file was in ORF 2 or 3.
+#' The output file will be of RF 1
+#'
+getRF <- function(x){
+  if (class(x)!= c("DNAStringSet")){
+    stop("This function only takes files of class 'DNAStringSet'. Use package 'Biostrings' to read the input file")
+  }
+  if (any(letterFrequency(x, "-", as.prob=TRUE))>0){
+    stop("This function do not take gaps as values/strings. Go and replaced gaps with Ns (if the gaps are at the begining or end of the sequences) or choose a subset of sequences that do not contain gaps")
+  }
+  # predORF() do not read seqs with gaps   # predicts longest ORF for sense strand in each query sequence
+  orf1 <- predORF(x[1:10], n=1, type="df", mode="orf", strand="sense")
+  longestSeq1 <- orf1$start[orf1$width == max(orf1$width)]
+  codonPosition1 <- longestSeq1[1]/3
+  codonConfirmation1 <- round(codonPosition1)
+  length(x)
+  if (codonPosition1==codonConfirmation1){
+    file.rename("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta", "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")                     # 640 bp
+    print("The Reading Frame of these seqs is +1, no further action is required, go to modelTesting")
+  } else {
+    if (codonPosition1!=codonConfirmation1){
+      tempLength <- length(x[[1]])
+      #remove first nucleotide position
+      newRF2 <- subseq(x, start = 2, end = tempLength)      # 639 bp
+      # getting new RF from the new alignment with -1 bp.
+      orf2 <- predORF(newRF2[1:10], n=1, type="df", mode="orf", strand="sense")
+      # checking size gain
+      longestSeq2 <- orf2$start[orf2$width == max(orf2$width)]
+      codonPosition2 <- (longestSeq2[1]+2)/3
+      codonConfirmation2 <- round(codonPosition2)
+      if (codonPosition2==codonConfirmation2){
[…]

### 2020-12-04 tests/testthat.R

@@ -14,10 +14,10 @@ library(ape)
 #library(phangorn)
 #library(systemPipeR)
 
-knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
-source('paper.R')
+knitr::knit('R/myFunctions.R', tangle=TRUE)
+source('R/myFunctions.R')
 
 testthat::test_file('tests/testthat/test-getRF.R')
 testthat::test_file('tests/testthat/test-getBestModel.R')
 
-system('rm paper.R')
+#system('rm R/myFunctions.R')

### 2020-12-04 tests/testthat.R
testing better

@@ -1,7 +1,7 @@
 install.packages('knitr')
 install.packages('testthat')
 install.packages('ape')
-#install.packages('phangorn')
+install.packages('phangorn')
 if (!requireNamespace("BiocManager", quietly = TRUE))
   install.packages("BiocManager")
 BiocManager::install("Biostrings")
@@ -11,7 +11,7 @@ library(knitr)
 library(testthat)
 library(Biostrings)
 library(ape)
-#library(phangorn)
+library(phangorn)
 #library(systemPipeR)
 
 knitr::knit('R/myFunctions.R', tangle=TRUE)

### 2020-12-04 tests/testthat/test-getBestModel.R
improving getBestModel

@@ -2,7 +2,11 @@
 context("test-getBestModel")
 
 SeqA <- c("ANTCATTTGA")
-SeqB <- as.DNAbin(c("TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC", "TAGTAGGTACTGCCC"))
+
+SeqB1 <- structure(c("123", "456", "789", "101", "AATTCAATGCTCGGGAAGCAAGGAAAGCTA", "AATTCCTAAAACACCAATCAAGTTGGTGTT", "AATTCACCACCACCACTAGCATACCATCCA", "AATTCTATTGGTCATCACAATGGTGGTCCG"), .Dim = c(4L, 2L))
+SeqB2 <- t(sapply(strsplit(SeqB1[,2],""), tolower))
+rownames(SeqB2) <- SeqB1[,1]
+SeqB3 <- as.DNAbin(SeqB2)
 
 test1 <- test_that("getBestModel", {
   expect_is(getBestModel, "function")
@@ -12,7 +16,7 @@ test2 <- test_that("getBestModel", {
   expect_error(getBestModel(SeqA)) # errors out when a file of class other than -DNAStringSet- is taken
 })
 
-bestModel <- getBestModel(SeqB)
+bestModel <- getBestModel(SeqB3)
 
 test3 <- test_that("getBestModel", {
   expect_true(class(bestModel) == "data.frame")  # outputs an object of class data.frame

### 2020-12-04 R/getBestModel.R
functions done

@@ -1,16 +1,0 @@
-#' getBestModel
-#' This function provides a dataframe with the list of models of nucleotide evolution sorted by AICc (from lowest to highest)
-#'
-#' @param y takes as input a dataset of aligned sequences of the same length and known ORF
-#' @return returns a dataframe with the list of models of evolution sorted by AICc, from the lowest to the highest AICc value
-#'
-#'
-getBestModel <- function(y){
-  if (class(y)!= c("DNAbin")){
-    stop("This function only takes files of class 'DNAbin'. Use package 'ape' to read the input file")
-  }
-  formatPhyDat <- as.phyDat(y)                #getting the class used by modelTest: phyDat or pml
-  modelChoice <- modelTest(formatPhyDat, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 4)
-  modelChoiceSorted <- modelChoice[order(modelChoice$AICc),]
-  return(modelChoiceSorted)
-}

### 2020-12-04 R/getRF.R

@@ -1,50 +1,0 @@
-#' getRF
-#' This function determines the Reading Frame (RF) of a dataset of nucleotide sequences
-#'
-#' @param x takes as input a dataset of aligned sequences of the same length
-#'
-#' As output, this function either renames the input file if the sequences are in the ORF 1, or it writes a new fasta file if the input file was in ORF 2 or 3.
-#' The output file will be of RF 1
-#'
-getRF <- function(x){
-  if (class(x)!= c("DNAStringSet")){
-    stop("This function only takes files of class 'DNAStringSet'. Use package 'Biostrings' to read the input file")
-  }
-  if (any(letterFrequency(x, "-", as.prob=TRUE))>0){
-    stop("This function do not take gaps as values/strings. Go and replaced gaps with Ns (if the gaps are at the begining or end of the sequences) or choose a subset of sequences that do not contain gaps")
-  }
-  # predORF() do not read seqs with gaps   # predicts longest ORF for sense strand in each query sequence
-  orf1 <- predORF(x[1:10], n=1, type="df", mode="orf", strand="sense")
-  longestSeq1 <- orf1$start[orf1$width == max(orf1$width)]
-  codonPosition1 <- longestSeq1[1]/3
-  codonConfirmation1 <- round(codonPosition1)
-  length(x)
-  if (codonPosition1==codonConfirmation1){
-    file.rename("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta", "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")                     # 640 bp
-    print("The Reading Frame of these seqs is +1, no further action is required, go to modelTesting")
-  } else {
-    if (codonPosition1!=codonConfirmation1){
-      tempLength <- length(x[[1]])
-      #remove first nucleotide position
-      newRF2 <- subseq(x, start = 2, end = tempLength)      # 639 bp
-      # getting new RF from the new alignment with -1 bp.
-      orf2 <- predORF(newRF2[1:10], n=1, type="df", mode="orf", strand="sense")
-      # checking size gain
-      longestSeq2 <- orf2$start[orf2$width == max(orf2$width)]
-      codonPosition2 <- (longestSeq2[1]+2)/3
-      codonConfirmation2 <- round(codonPosition2)
-      if (codonPosition2==codonConfirmation2){
[…]

### 2020-12-04 R/myFunctions.R

@@ -64,6 +64,7 @@ getBestModel <- function(y){
   }
   formatPhyDat <- as.phyDat(y)                #getting the class used by modelTest: phyDat or pml
   modelChoice <- modelTest(formatPhyDat, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 4)
-  modelChoiceSorted <- modelChoice[order(modelChoice$AICc),]
+  modelChoiceSorted <- as.data.frame(modelChoice[order(modelChoice$AICc),])
   return(modelChoiceSorted)
 }
+

### 2020-12-04 analysis/paper/paper.Rmd
pdf output

@@
 output: 
-    bookdown::word_document2:
+    pdf_document:
+      toc: FALSE
       fig_caption: yes
-      reference_docx: '../templates/template.docx'
+      citation_package: biblatex
+      latex_engine: xelatex
       pandoc_args:
@@
 
-#knitr::opts_knit$set(root.dir = '/rrcompendium/')
-
 ```
 
-```{r}
+```{r, echo = TRUE}
 
@@
 
-Tetis (19–28 mm standard length) were collected 400 m from the river mouth of Río Yumurí (Figure \@ref(fig:mainMap)), eastern Cuba, in December, 2018.
+Tetis (19–28 mm standard length) were collected 400 m from the river mouth of Río Yumurí (Figure 1), eastern Cuba, in December, 2018.
 Samples were collected using hand nets (1 mm mesh size) 2 m from the riverbank.
@@
 
-```{r mainMap, fig.cap="Sampling site at Yumuri river mouth, Eastern Cuba"}
+```{r mainMap, fig.cap="Sampling site at Yumuri river mouth, Eastern Cuba", echo = TRUE}
 
@@
 
-```{r gettingAPI}
+```{r gettingAPI, echo = TRUE}
 
@@
 
-```{r editingAPI}
+```{r editingAPI, echo = TRUE}
 
@@
 
-```{r alignment}
+```{r alignment, echo = TRUE}
 invisible({capture.output({                ### ALIGNING the sequences        # libraries "msa" and "Biostrings"
@@
 
-```{r, getRF, warning = FALSE}
+```{r, getRF, warning = FALSE, echo = TRUE}
 invisible({capture.output({               ### GETTING the READING FRAME     --- library "systemPipeR"
@@
 
-```{r getBestModel, warning = FALSE}
+```{r getBestModel, warning = FALSE, echo = TRUE}
                                       # FINDING THE BEST NUCLEOTIDE SUSBSTITUTION MODEL
@@
 
-```{r tree, echo = FALSE, include=TRUE, results = 'hide', fig.show='hide', warning = FALSE}
+```{r tree, echo = FALSE, include=TRUE, results = 'hide', fig.show='hide', warning = FALSE, echo = TRUE}
 
@@
 
-```{r loglikelihood, include = TRUE, results='hide'}
+```{r loglikelihood, include = TRUE, results='hide', echo = TRUE}
 # this result wont be displayed in the final document
@@
 
-```{r editingTree, echo = FALSE}
+```{r editingTree, echo = FALSE, fig.keep='none', echo = TRUE}
 
@@
 The final alignment comprised 209 sequences, each one of 639 base pairs.
-Figure \@ref(fig:SicydiumTree) depicts the ML tree (-lnL = `r loglikelihood`) showing the relationships among all COI sequences identified as *Sicydium* spp.
+Figure 2 depicts the ML tree (-lnL = `r loglikelihood`) showing the relationships among all COI sequences identified as *Sicydium* spp.
 available in GenBank.
@@
 Sequences highlited in green represent samples from Cuba.
-Numbers on branches represent boostrap values higher than 82.")}
+Numbers on branches represent boostrap values higher than 82."), echo = TRUE}
 
@@
 \newpage
-
-# References
-
-<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
-
-<div id="refs"></div>
-
-\newpage
-
-### Colophon
-
-This report was generated on `r Sys.time()` using the following computational environment and dependencies:
-
-```{r colophon, cache = FALSE}
-# which R packages and versions?
-if ("devtools" %in% installed.packages()) devtools::session_info()
-```
-
-The current Git commit details are:
-
-```{r}
-# what commit is this file at? 
-if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
-```

### 2020-12-04 README.Rmd
functions done
margins ok

@@
 
-[![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/sheilarm23/rrcompendium/master?urlpath=rstudio)
-
 This repository contains the data and code for our future paper:
 
-> Rodriguez-Machado, S. and Rodríguez-Cabrera, T.M. (2020). _The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba_. Name of journal/book <https://doi.org/{DOI-to-paper}>
+> Rodriguez-Machado, S. and Rodríguez-Cabrera, T.M. (2020). _The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba_.
 
@@
 Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`.
-Online at <https://doi.org/{DOI-to-compendium}>
 
@@
 Accessed `r format(Sys.Date(), "%d %b %Y")`.
-Online at <https://doi.org/{DOI-to-compendium}>
 
@@
 
-## How to run in your broswer or download and run locally
+## How to download and run locally
 
@@
 
-The simplest way to explore the text, code and data is to click on
-[binder](https://mybinder.org/v2/gh/sheilarm23/rrcompendium/master?urlpath=rstudio)
-to open an instance of RStudio in your browser, which will have the
-compendium files ready to work with. Binder uses rocker-project.org
-Docker images to ensure a consistent and reproducible computational
-environment. This way will take a few minutes to be ready but you will not need to install any packages on your computer.
-
 You can download the compendium as a zip from from this URL:
@@
 [DESCRIPTION](/DESCRIPTION) file).
-- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.Rmd")` in the R console
+- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.pdf`, or run `rmarkdown::render("analysis/paper/paper.Rmd")` in the R console
 
-Makefile:
-Whatever the option you choose to explore this compendium, you can also use the makefile (`/analysis/makefile`) in the terminal (`make all`) to compile the .docx document and then, you can use `make clean` to remove all files generated during the process.
+MAKEFILE:
+Whatever option you choose to explore, and run, this compendium, you can also use the makefile (`/analysis/makefile`) in the terminal (`make all`) to compile the .docx document and then, you can use `make clean` to remove all files generated during the process.
 

### 2020-12-04 analysis/paper/paper.Rmd
chunk size & margins

@@
 output: 
-    pdf_document:
+    bookdown::pdf_document2:
       toc: FALSE
       fig_caption: yes
-      citation_package: biblatex
+      #citation_package: biblatex
       latex_engine: xelatex
@@
 However, whether this fishery includes a single species or more than one species will remain as a mystery until more studies are conducted.
-Although this study represents the first effort toward and effective management, more thorough investigations and identification of the species comprising the “Teti” fishery are needed.
+Although this study represents the first effort toward an effective management, more thorough investigations and identification of the species comprising the “Teti” fishery are needed.
 It is imperative to include them in local management plans and to avoid overfishing this resource.
@@
 knitr::opts_chunk$set(
+  tidy.opts=list(width.cutoff=60),
+  tidy=TRUE,
   collapse = TRUE,
@@
 
-Tetis (19–28 mm standard length) were collected 400 m from the river mouth of Río Yumurí (Figure 1), eastern Cuba, in December, 2018.
+Tetis (19–28 mm standard length) were collected 400 m from the river mouth of Río Yumurí (Figure \@ref(fig:fig1)), eastern Cuba, in December, 2018.
 Samples were collected using hand nets (1 mm mesh size) 2 m from the riverbank.
@@
 
-```{r mainMap, fig.cap="Sampling site at Yumuri river mouth, Eastern Cuba", echo = TRUE}
+```{r fig1, fig.cap="Sampling site at Yumuri river mouth, Eastern Cuba", echo = TRUE}
 
@@
        geom_blank(data = mymap, aes(x = long, y = lat)) +
-       geom_map(data = mymap, map = mymap, aes(group = group, map_id = id), fill = "gray90", color = "black", size = 0.3) +
+       geom_map(data = mymap, map = mymap, aes(group = group, map_id = id),
+                fill = "gray90", color = "black", size = 0.3) +
        scale_x_continuous(limits = c(-74.51, -74.11), expand = c(0, 0)) +
@@
       geom_blank(data = mymap, aes(x = long, y = lat)) +
-      geom_map(data = mymap, map = mymap, aes(group = group, map_id = id), fill = "gray95", color = "black", size = 0.3) +
+      geom_map(data = mymap, map = mymap, aes(group = group, map_id = id), fill = "gray95",
+               color = "black", size = 0.3) +
       geom_path(data = zoomOut, aes(x = long, y = lat), size = 0.3) +
@@
       theme_map() +     # package ggthemes
-      theme(panel.background = element_rect(fill = NULL)) # fill = NULL or transparent, with color=NA both the background and the frame are transparent
+# fill = NULL or transparent, with color=NA both the background and the frame are transparent
+      theme(panel.background = element_rect(fill = NULL)) 
       )
@@
 Awaous_ocellarisCOI_search <- entrez_search(db = "nuccore", term = Awaous_ocellarisCOI, retmax = 4)
-Awaous_ocellarisCOI_seqs <- entrez_fetch(db = "nuccore", id = Awaous_ocellarisCOI_search$ids, rettype = "fasta")
+Awaous_ocellarisCOI_seqs <- entrez_fetch(db = "nuccore",
+                                         id = Awaous_ocellarisCOI_search$ids, rettype = "fasta")
 
@@
 # 'append = TRUE', I can keep adding seqs
-write(Sicydium_COI_seqs2, here(file="analysis/data/raw_data/Sicydium_COI_1-seqs.fasta"), sep = "\n", append = TRUE)   
-write(Awaous_bananaCOI_seqs, here(file="analysis/data/raw_data/Sicydium_COI_1-seqs.fasta"), sep = "\n", append = TRUE)
-write(Awaous_ocellarisCOI_seqs, here(file="analysis/data/raw_data/Sicydium_COI_1-seqs.fasta"), sep = "\n", append = TRUE)
+write(Sicydium_COI_seqs2,
+      here(file="analysis/data/raw_data/Sicydium_COI_1-seqs.fasta"), sep = "\n", append = TRUE)   
+write(Awaous_bananaCOI_seqs,
+      here(file="analysis/data/raw_data/Sicydium_COI_1-seqs.fasta"), sep = "\n", append = TRUE)
+write(Awaous_ocellarisCOI_seqs,
+      here(file="analysis/data/raw_data/Sicydium_COI_1-seqs.fasta"), sep = "\n", append = TRUE)
 
@@
 
-Sicydium_COI_seqinrFormat <- read.fasta(here(file ="analysis/data/raw_data/Sicydium_COI_1-seqs.fasta"), seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
+Sicydium_COI_seqinrFormat <- read.fasta(
+  here(file ="analysis/data/raw_data/Sicydium_COI_1-seqs.fasta"),
+  seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
 
-Sicydium_COI_names <- attr(Sicydium_COI_seqinrFormat, "name")      # editing species name
-Sicydium_COI_names <- gsub("\\..*", "", Sicydium_COI_names)        # eliminating characters after "." using gsub
-Sicydium_COI_names <- gsub("^.*\\|", "", Sicydium_COI_names)       # eliminating characters before "|" using gsub
+# editing species name
+Sicydium_COI_names <- attr(Sicydium_COI_seqinrFormat, "name") 
+# eliminating characters after "." using gsub
+Sicydium_COI_names <- gsub("\\..*", "", Sicydium_COI_names)        
+# eliminating characters before "|" using gsub
+Sicydium_COI_names <- gsub("^.*\\|", "", Sicydium_COI_names)
 
-# using library "ape"
-Sicydium_COI_apeFormat <- read.GenBank(Sicydium_COI_names)         # this line takes a few minutes to run
+# using library "ape" --- this line takes a few minutes to run
+Sicydium_COI_apeFormat <- read.GenBank(Sicydium_COI_names)
 
 # creating a vector (to be added as name of the seqs) with the species, GenBank accession number, and type of gene
-Sicydium_COI_GenBank_Ids <- paste(attr(Sicydium_COI_apeFormat, "species"), names(Sicydium_COI_apeFormat), sep = "_COI_")
+Sicydium_COI_GenBank_Ids <- paste(attr(Sicydium_COI_apeFormat, "species"),
+                                  names(Sicydium_COI_apeFormat),
+                                  sep = "_COI_")
 
 # writing a new fasta file (2nd) using the vector above 
-write.fasta(sequences = Sicydium_COI_seqinrFormat, names = Sicydium_COI_GenBank_Ids, nbchar = 10,
+write.fasta(sequences = Sicydium_COI_seqinrFormat,
+            names = Sicydium_COI_GenBank_Ids, nbchar = 10,
             here(file.out = "analysis/data/raw_data/Sicydium_COI_2-seqinrFormat.fasta"))
@@
 fastaNames <- strsplit(names(nameFasta), '_')       # creating a vector splitting species names
-completeSeqs <- which(! grepl("sp.", fastaNames))              # selecting only those with specific name in them
+completeSeqs <- which(! grepl("sp.", fastaNames))   # selecting only those with specific name
 
@@
 fastaLength <- lengthFasta[which(getLength(lengthFasta) >= 600)]
-write.fasta(sequences = fastaLength, names = names(fastaLength),
+write.fasta(sequences = fastaLength,
+            names = names(fastaLength),
             here(file.out = "analysis/data/raw_data/Sicydium_COI_4-lengths.fasta"))             # 401 seqs
@@
 uniqueSeqs <- duplicatesFasta[-dups]
-write.fasta(sequences = uniqueSeqs, names = uniqueNames,
+
+write.fasta(sequences = uniqueSeqs,
+            names = uniqueNames,
             here(file.out = "analysis/data/raw_data/Sicydium_COI_5-dupli.fasta"))               # 204 seqs
@@
 mySamples <- read.fasta(here("analysis/data/raw_data/Sicydium_COI_Cuba.fas"))
-write.fasta(mySamples, names(mySamples), here("analysis/data/raw_data/Sicydium_COI_5-dupli.fasta"), open = "a")              # 210 seqs
+write.fasta(mySamples,
+            names(mySamples),
+            here("analysis/data/raw_data/Sicydium_COI_5-dupli.fasta"), open = "a")              # 210 seqs
 
@@
 
-alignmentFasta <- readDNAStringSet(here("analysis/data/raw_data/Sicydium_COI_5-dupli.fasta"))
+alignmentFasta <- readDNAStringSet(here(
+  "analysis/data/raw_data/Sicydium_COI_5-dupli.fasta"))
 set.seed(123)
@@
 nucType <- alphabetFrequency(ClustalOAlignment)
-unknownNuc <- nucType[, c(5:14)]  # this subset only contains nucleotides that are not A, G, T, C, or N or gaps
+unknownNuc <- nucType[, c(5:14)]  # only nucleotides that are not A, G, T, C, or N or gaps
 unknownPosition <- which(unknownNuc > 0, arr.ind=T)
@@
 ClustalOAlignment2 <- as(ClustalOAlignment, "DNAStringSet")
-writeXStringSet(ClustalOAlignment2, here(file="analysis/data/raw_data/Sicydium_COI_6-alignment.fasta"))            # 209 seqs -- 640 bp
+writeXStringSet(ClustalOAlignment2,
+                here("analysis/data/raw_data/Sicydium_COI_6-alignment.fasta"))            # 209 seqs -- 640 bp
 
 ### Replacing gaps at the begining and end with Ns
-gapsFasta <- read.fasta(here(file = "analysis/data/raw_data/Sicydium_COI_6-alignment.fasta"), seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE, apply.mask = FALSE)
+gapsFasta <- read.fasta(here(file = "analysis/data/raw_data/Sicydium_COI_6-alignment.fasta"),
+                        seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE, apply.mask = FALSE)
 #getLength(gapsFasta)
 replacingGaps <- lapply(gapsFasta, function(x) gsub("-", "N", x))
-write.fasta(replacingGaps, names(replacingGaps), here("analysis/data/raw_data/Sicydium_COI_7-gaps.fasta"))
+write.fasta(replacingGaps, names(replacingGaps),
+            here("analysis/data/raw_data/Sicydium_COI_7-gaps.fasta"))
 
@@
   if (class(x)!= c("DNAStringSet")){
-    stop("This function only takes files of class 'DNAStringSet'. Use package 'Biostrings' to read the input file")
+    stop("This function only takes files of class 'DNAStringSet'.
+         Use package 'Biostrings' to read the input file")
   }
@@
     stop("This function do not take gaps as values/strings.
-Go and replaced gaps with Ns (if the gaps are at the begining or end of the sequences) or choose a subset of sequences that do not contain gaps")
+         Go and replaced gaps with Ns (if the gaps are at the begining
+         or end of the sequences)
+         or choose a subset of sequences that do not contain gaps")
   }
-  orf1 <- predORF(x[1:10], n=1, type="df", mode="orf", strand="sense")   # predORF() do not read seqs with gaps
+  orf1 <- predORF(x[1:10], n=1, type="df", mode="orf", strand="sense")
+  # predORF() do not read seqs with gaps
   # predicts longest ORF for sense strand in each query sequence                                                                          
@@
   if (codonPosition1==codonConfirmation1){
-   file.rename(here("analysis/data/raw_data/Sicydium_COI_7-gaps.fasta"), here("analysis/data/raw_data/Sicydium_COI_newRF.fasta")) 
-  print("The Reading Frame of these seqs is +1, no further action is required, go to modelTesting")
+   file.rename(here("analysis/data/raw_data/Sicydium_COI_7-gaps.fasta"),
+               here("analysis/data/raw_data/Sicydium_COI_newRF.fasta")) 
+  print("The Reading Frame of these seqs is +1, go to ModelTesting")
     } else {
@@
       if (codonPosition2==codonConfirmation2){
-        writeXStringSet(newRF2, here(file="analysis/data/raw_data/Sicydium_COI_newRF.fasta"))                    
-      print("The Reading Frame of these seqs was +2, new fasta file created, go to modelTesting")
+        writeXStringSet(newRF2, 
+                        here("analysis/data/raw_data/Sicydium_COI_newRF.fasta"))                    
+      print("The Reading Frame of these seqs was +2,
+            new fasta file created, go to modelTesting")
       } else {
@@
         newRF3 <- subseq(x, start = 3, end = tempLength)   # 638 bp
-        writeXStringSet(newRF3, here(file="analysis/data/raw_data/Sicydium_COI_newRF.fasta"))                                
-      print("The Reading Frame of these seqs was +3, new fasta file created, go to modelTesting")
+        writeXStringSet(newRF3, 
+                        here("analysis/data/raw_data/Sicydium_COI_newRF.fasta"))                                
+      print("The Reading Frame of these seqs was +3,
+            new fasta file created, go to modelTesting")
      }
@@
   if (class(y)!= c("DNAbin")){
-    stop("This function only takes files of class 'DNAbin'. Use package 'ape' to read the input file")
+    stop("This function only takes files of class 'DNAbin'.
+         Use package 'ape' to read the input file")
   }
@@
 
-bestModel <- suppressWarnings(suppressMessages(getBestModel(y = fastaRF)))               # getting the AICc for each competing model
+bestModel <- suppressWarnings(suppressMessages(getBestModel(y = fastaRF)))
 
@@
 
-```{r tree, echo = FALSE, include=TRUE, results = 'hide', fig.show='hide', warning = FALSE, echo = TRUE}
+```{r tree, echo = TRUE, include=TRUE, results = 'hide', fig.show='hide', warning = FALSE}
 
@@
 set.seed(123)
-fitStart = eval(get(bestModel$Model[which.min(bestModel$AICc)], env), env)    # initial fit for the model chosen
+# initial fit for the model chosen
+fitStart = eval(get(bestModel$Model[which.min(bestModel$AICc)], env), env)
 
@@
 set.seed(123)
-bsTree <- plotBS(midpoint(fit$tree), bs, p = 75, type="p", bs.col = "blue")     # fit$tree is where the optimal tree is stored
+# fit$tree is where the optimal tree is stored
+bsTree <- plotBS(midpoint(fit$tree), bs, p = 75, type="p", bs.col = "blue")
 
@@
 
-```{r editingTree, echo = FALSE, fig.keep='none', echo = TRUE}
+```{r editingTree, echo = FALSE, echo = TRUE}
 
@@
 # checking that the tree is rooted
-rootedTree <- is.rooted(tree)   # package ape. showing the tree root: geom_rootpoint(color="pink", size=5) 
+# package ape. showing the tree root: geom_rootpoint(color="pink", size=5) 
+rootedTree <- is.rooted(tree)   
 # getting the node number for each main clade. This info is important because those #s are the one needed to edit each main clade/species
@@
 The final alignment comprised 209 sequences, each one of 639 base pairs.
-Figure 2 depicts the ML tree (-lnL = `r loglikelihood`) showing the relationships among all COI sequences identified as *Sicydium* spp.
+Figure \@ref(fig:fig2) depicts the ML tree (-lnL = `r loglikelihood`) showing the relationships among all COI sequences identified as *Sicydium* spp.
 available in GenBank.
@@
 
-```{r SicydiumTree, fig.width = 15, fig.height = 19, fig.cap=paste("ML phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*.
+```{r fig2, fig.width = 15, fig.height = 19, fig.cap=paste("ML phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*.
 Sequences highlited in green represent samples from Cuba.
@@
   geom_cladelabel(node=211, label="Awaous spp.", color="darkgreen", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") +
-  # samples Sicydium_COI_Cuba2,4,6 were nested within this clde but the distance is too short to be displayed in the tree
   geom_hilight(node=191, fill="green", alpha = 0.7, extend=.004) +
@@
   theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))
-
-# suppressWarnings(suppressMessages(print(Sicydium_ML_tree)))
 
@@
 \newpage
+
+# References
+
+<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->
+
+<div id="refs"></div>
+
+\newpage
+
+### Colophon
+
+This report was generated on `r Sys.time()` using the following computational environment and dependencies:
+
+```{r colophon, cache = FALSE}
+# which R packages and versions?
+if ("devtools" %in% installed.packages()) devtools::session_info()
+```
+
+The current Git commit details are:
+
+```{r}
+# what commit is this file at? 
+if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
+```

### 2020-12-04 tests/testthat.R
testthat matching travis

@@ -14,10 +14,10 @@ library(ape)
 library(phangorn)
 #library(systemPipeR)
 
-knitr::knit('R/myFunctions.R', tangle=TRUE)
-source('R/myFunctions.R')
+knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
+source('analysis/paper/paper.R')
 
 testthat::test_file('tests/testthat/test-getRF.R')
 testthat::test_file('tests/testthat/test-getBestModel.R')
 
-#system('rm R/myFunctions.R')
+system('rm analysis/paper/paper.R')

### 2020-12-04 tests/testthat.R
paper.R missing

@@ -15,9 +15,9 @@ library(phangorn)
 #library(systemPipeR)
 
 knitr::knit('analysis/paper/paper.Rmd', tangle=TRUE)
-source('analysis/paper/paper.R')
+source('paper.R')
 
 testthat::test_file('tests/testthat/test-getRF.R')
 testthat::test_file('tests/testthat/test-getBestModel.R')
 
-system('rm analysis/paper/paper.R')
+system('rm paper.R')

### 2021-09-06 analysis/paper/paper.Rmd
pdf done
paperLastVersion092021

--- 2020-12-03 analysis/paper/paper.Rmd
@@
 ---
-title: 'The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba'
+title: 'Identity of fish fry from the “Tetí” fishery in Eastern Cuba'
 author:
@@
       correspondence: false
+  - Prosanta Chakrabarty:
+      email: prosanta@lsu.edu
+      institute: [UofO]
+      correspondence: true
 institute:
@@
 bibliography: references.bib
-csl: '../templates/journal-of-archaeological-science.csl'
+csl: '../templates/cell-structure-and-function.csl'
 abstract: |
-  The lack of diagnostic characters to distinguish among species and even genera of gobies (Gobiidae) is still a challenge for science in the 21st century.
-However, the use of molecular tools applied to species-level identification has brought some light into the taxonomy and systematics of these fish taxa.
-The fish fry fishery in Eastern Cuba known as “Teti” represents a local fishery and there is no information about the exploitation of this natural resource outside this region.
-Fishing “Teti” has been a common practice for several years but an exponential increase of tourism in Eastern Cuba represents an additional threat to our biodiversity.
-Besides, if “Teti” refers to a single species or a species complex is still a knowledge gap in Cuba.
-The present study aims to identify the Fish Fry "Teti" Fishery in Eastern Cuba using molecular markers.
-We conducted a barcoding analysis comparing Cuban samples of gobies with samples of the genera *Sicydium* and *Awaous* available in GenBank.
-Here, we provide evidence that Busck´s Stone-Biting Goby, *Sicydium buscki*, is a component of the fish fry “Teti” fishery in Eastern Cuba.
-However, whether this fishery includes a single species or more than one species will remain as a mystery until more studies are conducted.
-Although this study represents the first effort toward and effective management, more thorough investigations and identification of the species comprising the “Teti” fishery are needed.
-It is imperative to include them in local management plans and to avoid overfishing this resource.
+  Fisheries of juvenile and larval fishes (i.e., fish-fry fisheries) often exploit the migratory behavior of diadromous species on their return to rivers.
+Although these fisheries are common worldwide, fish-fry fisheries are subject to insufficient management and the identity of the target species is frequently unknown.
+The artisanal fish-fry fishery in Cuba is known as “Tetí”, and is restricted to the eastern region of the country.
+Fishing “Tetí” has been practiced for hundreds of years, but an increase in tourism has elevated demand in recent years and may pose a threat to the conservation of this natural resource.
+It remains unclear if “Tetí” refers to one or multiple species.
+The present study aims to identify the fish fry that make up the “Tetí” fishery in Eastern Cuba with genetic data.
+We first used a standard nucleotide BLAST to identify the sequences.
+Then, we recovered the phylogenetic position of these samples based on a maximum likelihood phylogenetic analysis that also used available sequences from GenBank.
+Here, we provide evidence that Busck´s Stone-Biting Goby, Sicydium buscki, is a component of the “Tetí” fishery in Cuba.
+However, whether this fishery includes one or more species remains a mystery until additional samples can be analyzed.
+However, with this species identification the first efforts towards effective management can be taken to manage this fishery.
+Collecting additional scientific and fishing data from the “Tetí” fishery and including them in local management plans will be an important strategy for avoiding overfishing this resource.
 keywords: |
@@
 )
-
 #knitr::opts_knit$set(root.dir = '/rrcompendium/')
-
 ```
@@
 ```{r}
-
 # INSTALLING PACKAGES AND LIBRARIES - if needed
 packages = c("BiocManager", "raster", "ggplot2", "ggthemes", "mapproj", "here",  
-             "rentrez", "ape", "seqinr", "phangorn")
-
+             "rentrez", "ape", "seqinr", "phangorn",
+             "adegenet", "pegas")
 package.check <- lapply(packages, FUN = function(x) {
@@
 )
-
 if (!requireNamespace("BiocManager", quietly = TRUE))
@@
 
-The family Gobiidae includes almost 2,000 species of marine, freshwater, and brackish fishes worldwide [@eschmeyer2020catalog].
-This group is mostly represented by small-sized and bottom-dweller fishes found in tropical and subtropical waters.
-Such a small size along with cryptic body colors and the lack of diagnostic characters to distinguish among species and even genera is still a challenge for science in the 21st century.
-Only in the last ten years, more than a hundred gobiid species were described [@eschmeyer2020catalog].
-However, the use of molecular tools applied to species-level identification has brought some light into the taxonomy and systematics of these fish taxa [@thacker2011phylogeny; @victor2014three; @keith2017schismatogobius].
+The family Gobiidae includes almost 2,000 species of marine, freshwater, and brackish water fishes worldwide [@eschmeyer2020catalog].
+This group is mostly represented by small and bottom-dwelling fishes found in tropical and subtropical waters.
+Their small size along with cryptic body colors and the paucity of clear diagnostic characters to distinguish among taxa is a substantial challenge.
+However, the use of molecular tools applied to species-level identification has shed light on the taxonomy of these fishes [@keith2011phylogeny; @thacker2011phylogeny; @victor2014three; @kuang2018phylogenomic] and in the last ten years, more than one hundred gobiid species have been described [@eschmeyer2020catalog].
 
-Although most species in the family Gobiidae inhabit marine ecosystems, most species inhabiting freshwater or estuarine ecosystems display an amphidromous lifecycle [@myers1949usage; @bell1999overview].
-This means that adults spawn in freshwater and when the eggs hatch, they are taken downstream by river flow into the ocean.
-Then, after spending from a few weeks to a few months (depending on the species), postlarvae return to rivers.
-This moment of postlarval returning is particularly fascinating because million of them gather at the river mouth as they start going upstream.
-This migratory behavior, particularly, when the postlarvae are going back to rivers, supports an artisanal fishery around the world [@bell1999overview; @mcdowall2007amphidromy].
-Known as goby-fry fisheries, this is a common practice in several countries in the Indo-Pacific [@asis2013first], the Pacific [@castellanos2011goby], and the Caribbean [@neal2009factors].
-The problem with it is that during the last years insufficient management of these fisheries has led to a considerable reduction in catches [@bell1999overview].
-In addition, information on the identity of the species comprising these goby-fry fisheries in many places is still incomplete [@castellanos2011goby; @asis2013first].
+Although most gobies inhabit marine ecosystems, some species inhabiting freshwater or estuarine ecosystems display an amphidromous lifecycle [@myers1949usage; @bell1999overview]: meaning that adults spawn in freshwater and the post-hatch larvae are carried downstream by rivers into the ocean.
+After spending weeks to months (depending on the species) in these marine habitats, postlarval stage individuals return to rivers.
+This period of postlarval return is particularly fascinating because million of them gather at the river mouth before starting their journey upstream.
+This migratory behavior supports artisanal fisheries worldwide [@bell1999overview; @mcdowall2007amphidromy].
+Goby-fry fisheries are common practice in several countries in the Indo-Pacific [@asis2013first], the Pacific [@castellanos2011goby], and the Caribbean (@erdman1961notes; @kwaketal2016).
+One problem with this style of fishery is that insufficient management has led to a considerable reduction in catch [@bell1999overview].
+In addition, information on the identity of the species comprising these goby-fry fisheries is still incomplete in many places [@castellanos2011goby; @asis2013first].
 
 Gobies have been poorly studied in Cuba and available information on their identification and distribution is not conclusive [@lara2010dna].
-The fish fry fishery in Eastern Cuba known as “Teti” represents a local fishery and there is no information about the exploitation of this natural resource outside this region.
-Base on other reports in the Caribbean region [@neal2009factors], some authors claimed that the gobies *Sicydium plumieri* and *Sicydium punctatum* comprise the “Teti” fisheries in Cuba, but there is no solid evidence on that yet.
-Fishing “teti” has been a common practice for several years but an exponential increase of tourism in Eastern Cuba, due to recent economic policy changes, represents an additional threat to our biodiversity.
-Besides, if “teti” refers to a single species or a species complex is still a knowledge gap in Cuba.
-The present study aims to identify the Fish Fry "Teti" Fishery in Eastern Cuba using molecular markers.
+The fish-fry “Tetí” fishery in Eastern Cuba represents a local artisanal fishery and there is no scientific information about the exploitation of this natural resource outside of the region.
+Based on other reports from the Caribbean (*e.g.*, @watson2000), some authors have claimed that the gobies *Sicydium plumieri* and *Sicydium punctatum* make up the “Tetí” fishery in Cuba, but evidence for those identications are lacking.
+“Tetí” has been historically exploited as the main ingredient of “traditional and exotic” dishes [@pichardoytapia1862; A.
+Hartman, pers.
+comm., August 01, 2011).
+However, recent changes to Cuban economic policies have promoted considerable growth in tourism [@rutty2019tourism], leading to increased fisheries pressure.
+Annual data series show increases in the number of visitors and infrastructure development in Baracoa [@baracoaStats2015; @baracoaStats2019], where this fishery takes place.
+Such actions along with the lack of fishing statistics might represent a threat to the conservation of these fish.
+This study aims to identify fish-fry species involved in the “Tetí” fishery in Eastern Cuba using molecular data.
 
@@
 
-Tetis (19–28 mm standard length) were collected 400 m from the river mouth of Río Yumurí (Figure \@ref(fig:mainMap)), eastern Cuba, in December, 2018.
-Samples were collected using hand nets (1 mm mesh size) 2 m from the riverbank.
-Specimens were anesthetized using MS 222 (Tricaine Methanesulphonate) and immediately preserved in 80% ethanol.
-Voucher specimens were deposited in the collection of the Acuario Nacional de Cuba (catalog number: ANC-xxxx).
+Nine “Tetís” (19–28 mm standard length) (Figure \@ref(fig:tetiPhoto)) were collected 400 m from the river mouth of the Río Yumurí (Figure \@ref(fig:mainMap)), eastern Cuba, in December 2018.
+Specimens were collected using hand nets (1 mm mesh size) 2 m from the riverbank, anesthetized using MS 222 (Tricaine methanesulphonate), and immediately preserved in 80% ethanol.
+Voucher specimens were deposited in the collection of the Acuario Nacional de Cuba (catalog number: ANC-12.2.454).
 
-```{r mainMap, fig.cap="Sampling site at Yumuri river mouth, Eastern Cuba"}
+```{r tetiPhoto, fig.cap="Lateral view of goby fry “Tetí” captured in Río Yumurí, Cuba. Photo by Tomás M. Rodríguez-Cabrera."}
+
+```
+
+```{r mainMap, fig.cap="Map of Cuba indicating where the fish-fry “Tetí” fishery has been reported. The outer circle represents the sampling locality of this study."}
 
@@
 mymap <- fortify(mydata)                                # package ggplot2
-
 # sampling site
@@
        theme_bw()
-
 # small square framing the zomm area
@@
                    lat = c(20.13, 20.40, 20.40, 20.13, 20.13))
-
 # inset map
@@
       )
-
 mainMap +
@@
 
-Total DNA was extracted from caudal peduncle muscle tissues using the DNAeasy tissue kit (QIAGEN).
+Total DNA was extracted from caudal peduncle muscle tissues using the DNeasy tissue kit (QIAGEN).
 Partial sequences of the mitochondrial gene cytochrome oxidase I (COI) were amplified through the polymerase chain reaction in 20 μL of reaction volume (one unit of GoTaq DNA polymerase-Promega, 0.2 nM of each primer, 0.2 μM dNTPs, and 1.5 mM MgCl2) [@lara2010dna].
@@
 PCR products were purified using ExoSAP-IT™ PCR Product Cleanup Reagent (Affymetrix Inc./ USB, Cleveland, OH, USA), and then sequenced on an ABI 3100 automated sequencer (Applied Biosystems).
-Raw sequence files were edited and aligned with the software Geneious v.9.1 [@kearse2012geneious], and submitted to GenBank database (accession numbers #####).
-We used sequences of the genus *Sicydium* as reference for recovering the taxonomic identity of the "Tetis".
-Sequences of *Awaous banana* and *A.
-ocellaris* were used as outgroup.
-All non-Cuban sequences were recovered from GenBank using the R package *rentrez* [@winter2017rentrez].
+Raw sequence files were edited and aligned with the software Geneious v.9.1 [@kearse2012geneious] and submitted to GenBank (accession numbers: MZ848370– MZ848378).
+We used a standard nucleotide BLAST (BLASTN 2.11.0; @zhangetal2000), optimized for highly similar sequences (MegaBLAST; @morgulis2008database) (Query IDs: 1237–1245), and recovered the identity of the Cuban sequences with 99–100% identity.
+
+We also aimed to recover the phylogenetic position of the fish-fry samples.
+For this analysis, all non-Cuban sequences were retrieved from GenBank using the R package *rentrez* [@winter2017rentrez] (Appendix 1).
 Duplicated sequences and sequences with missing data were excluded.
-Sequences were edited using the R packages *seqinr* [@charif2007seqinr], *ape* [@paradis2015package], and *Biostrings* [@pages2013package].
-Final alignment was obtained with the algorithm ClustalOmega [@sievers2011fast] using the R package *msa* [@bodenhofer2015msa].
+The dataset was edited and aligned using the R packages *seqinr* [@charif2007seqinr], *ape* [@paradis2015package], and *Biostrings* [@pages2013package].
+The final alignment was obtained with the algorithm ClustalOmega [@sievers2011fast] using the R package *msa* [@bodenhofer2015msa].
 
 ```{r gettingAPI}
-
                                                             # DOWNLOADING SEQS FROM GENBANK
 # rentrez: this package provides an R interface NCBI API and allows to search databases like GenBank (http://www.ncbi.nih.gov/genbank/) 
-
 Sicydium_COI <- "Sicydium[Organism] AND COI[Gene]"
-
 # I set retmax=580 because after running this by default, I got a message with the number of hits (580 in this case)
 Sicydium_COI_search <- entrez_search(db = "nuccore", term = Sicydium_COI, retmax = 580)
-
 # given that the search is too big for the system, it needs to be split into two smaller ones
@@
 Sicydium_COI_seqs2 <- entrez_fetch(db = "nuccore", id = Sicydium_COI_search$ids[296:580], rettype = "fasta")
-
 # GETTING OUTGROUP SEQUENCES OF AWAOUS BANANA (2) AND A. OCELLARIS (2)
-
 Awaous_bananaCOI <- "Awaous_banana[Organism] AND COI[Gene]"
@@
 Awaous_bananaCOI_seqs <- entrez_fetch(db = "nuccore", id = Awaous_bananaCOI_search$ids, rettype = "fasta")
-
 Awaous_ocellarisCOI <- "Awaous_ocellaris[Organism] AND COI[Gene]"
@@
 Awaous_ocellarisCOI_seqs <- entrez_fetch(db = "nuccore", id = Awaous_ocellarisCOI_search$ids, rettype = "fasta")
-
 # storing ALL sequences in a single fasta file
@@
 ```{r editingAPI}
-
 # seqinr: package for sequence manipulation 
 # reading the file and renaming the sequences
-
 Sicydium_COI_seqinrFormat <- read.fasta(here(file ="analysis/data/raw_data/Sicydium_COI_1-seqs.fasta"), seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
-
 Sicydium_COI_names <- attr(Sicydium_COI_seqinrFormat, "name")      # editing species name
@@
 Sicydium_COI_names <- gsub("^.*\\|", "", Sicydium_COI_names)       # eliminating characters before "|" using gsub
-
 # using library "ape"
 Sicydium_COI_apeFormat <- read.GenBank(Sicydium_COI_names)         # this line takes a few minutes to run
-
 # creating a vector (to be added as name of the seqs) with the species, GenBank accession number, and type of gene
 Sicydium_COI_GenBank_Ids <- paste(attr(Sicydium_COI_apeFormat, "species"), names(Sicydium_COI_apeFormat), sep = "_COI_")
-
 # writing a new fasta file (2nd) using the vector above 
@@
             here(file.out = "analysis/data/raw_data/Sicydium_COI_2-seqinrFormat.fasta"))
-
                                      # EDITTING SEQS AND GETTING THEM ALL READY FOR THE PHYLOGENETIC ANALYSIS
-
 ### REMOVING sequences of incomplete/unspecific NAMES; that is, those with 'sp.'
@@
                         seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
-
 fastaNames <- strsplit(names(nameFasta), '_')                  # creating a vector splitting species names
 completeSeqs <- which(! grepl("sp.", fastaNames))              # selecting only those with specific name in them  
-
 # writing a new fasta (3rd) file with only those seqs with full scientific names
@@
             here(file.out = "analysis/data/raw_data/Sicydium_COI_3-names.fasta"))
-
 ### REMOVING sequences of LENGTH shorter than 600 pb, given that the gene COI has between 627 and 658 bp.
@@
             here(file.out = "analysis/data/raw_data/Sicydium_COI_4-lengths.fasta"))             # 401 seqs
-
 ### Removing DUPLICATE sequences
@@
 duplicatesFasta <- read.fasta(here("analysis/data/raw_data/Sicydium_COI_4-lengths.fasta"), set.attributes = FALSE) 
-
 seqnames <- names(duplicatesFasta)
@@
             here(file.out = "analysis/data/raw_data/Sicydium_COI_5-dupli.fasta"))               # 204 seqs
-
-# adding samples from Cuba (n=6)
-mySamples <- read.fasta(here("analysis/data/raw_data/Sicydium_COI_Cuba.fas"))
+# adding samples from Cuba (n=9)
+mySamples <- read.fasta(here("analysis/data/raw_data/Sicydium_COI_Cuba_n=9.fas"))
 write.fasta(mySamples, names(mySamples), here("analysis/data/raw_data/Sicydium_COI_5-dupli.fasta"), open = "a")              # 210 seqs
@@
 invisible({capture.output({                ### ALIGNING the sequences        # libraries "msa" and "Biostrings"
-
 alignmentFasta <- readDNAStringSet(here("analysis/data/raw_data/Sicydium_COI_5-dupli.fasta"))
@@
 # print(ClustalOAlignment, showConsensus=FALSE, showNames=TRUE, show="complete") 
-
 # Checking for the position of sequences that were not edited properly so they still have nucleotides different that A, G, C, T 
@@
 unknownPosition <- which(unknownNuc > 0, arr.ind=T)
-
 # REMOVING samples with too many GAPS at the beginning and the end of the sequences and UNKNOWN nucleotides (ex.: W, K, S) 
@@
 writeXStringSet(ClustalOAlignment2, here(file="analysis/data/raw_data/Sicydium_COI_6-alignment.fasta"))            # 209 seqs -- 640 bp
-
 ### Replacing gaps at the begining and end with Ns
@@
 write.fasta(replacingGaps, names(replacingGaps), here("analysis/data/raw_data/Sicydium_COI_7-gaps.fasta"))
-
 })})
-
 ```
@@
 alignedFasta <- readDNAStringSet(here("analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")) 
-
 getRF <- function(x){
@@
 }
-
 suppressWarnings(suppressMessages(print(getRF(x = alignedFasta))))
-
 })})
-
 ```
 
-The nucleotide substitution model that best explained our data was estimated using the function modelTest from the R package *phangorn* [@schliep2011phangorn].
+The nucleotide substitution model that best explained our data was inferred using the function modelTest from the R package *phangorn* [@schliep2011phangorn].
 The phylogenetic tree was reconstructed following a maximum likelihood approach [@felsenstein1981evolutionary] using the functions optim.pml and bootstrap.pml, also from the package *phangorn*.
 The tree was edited and visualized using the R package *ggtree* [@yu2017ggtree].
+Sequences of *Awaous banana* and *A.
+ocellaris* were used as outgroups.
+Genetic diversity (the number of haplotypes as well as haplotype and nucleotide diversity) was estimated using the package *pegas* [@paradis2010pegas].
 
-This paper was completely developed in RStudio, PBC (“RStudio”) [@rstudio2020].
-R code and data are available as a research compendium for generating this paper in the following link: https://github.com/sheilarm23/rrcompendium.git.
+All analyses and figures here were developed in RStudio, PBC (“RStudio”) [@rstudio2020].
+R code and data are available as a research compendium in the following link: https://github.com/sheilarm23/rrcompendium.git.
 
@@
                                       # FINDING THE BEST NUCLEOTIDE SUSBSTITUTION MODEL
-
 # libraries ape and phangorn (modelTest is a function of the package phangorn)
-
 fastaRF <- read.dna(here(file="analysis/data/raw_data/Sicydium_COI_newRF.fasta"), format = "fasta")
-
 getBestModel <- function(y){
@@
 bestModel <- suppressWarnings(suppressMessages(getBestModel(y = fastaRF)))               # getting the AICc for each competing model
-
 # DISCLAIMER: visually inspecting the AICc is better than just getting the best model automatically, because it allows to compares models with similar AICc and it is my personal opinion to choose the one with the least parameters.
 # The alternative is to get automatically the first model that the data.frame modelTest gives and storing it as a variable. In order to get this as reproducible as possible, I chose the alternative.
-
 ```
@@
 ```{r tree, echo = FALSE, include=TRUE, results = 'hide', fig.show='hide', warning = FALSE}
-
                                                         # BUILINDG THE TREE
@@
 fitStart = eval(get(bestModel$Model[which.min(bestModel$AICc)], env), env)    # initial fit for the model chosen
-
 set.seed(123)
@@
 bs = bootstrap.pml(fit, bs=100, optNni=TRUE, multicore=TRUE, mc.cores = 4)
-
 # saving the tree in the folder 'analysis/figures/'
@@
 bsTree <- plotBS(midpoint(fit$tree), bs, p = 75, type="p", bs.col = "blue")     # fit$tree is where the optimal tree is stored
-
 set.seed(123)
@@
 ```{r editingTree, echo = FALSE}
-
 # Editing and visualizing the tree to represent --- library: ggtree
 tree <- read.tree(here(file="analysis/data/derived_data/Sicydium_ML_tree.tre"))
-
 # checking that the tree is rooted
@@
 set.seed(123)
-nodeNumber <- ggtree(tree) + geom_text(aes(label=node), hjust=-.5, col = "blue", size = 1) # looks horrible but it gives the clade number
+nodeNumber <- ggtree(tree) + geom_text(aes(label=node), hjust=-.5, col = "blue", size = 1) # looks horrible but it gives the node number
+
+```
+
+```{r genetic diversity, echo = FALSE}
+# uses packages "pegas", which depends on "adegenet"
+
+diversityData <- read.dna(here("analysis/data/raw_data/Sicydium_COI_Cuba_n=9.fas"), format = "fasta")
+diversityToUse <- haplotype(diversityData)
+
+haploDiv <- round(hap.div(diversityToUse, variance = TRUE), 4)
+nucleDiv <- round(nuc.div(diversityToUse, variance = TRUE), 4)
+
+# to see the network: net <- haploNet(diversityToUse) and plot(net)
 
@@
 
-The final alignment comprised 209 sequences, each one of 639 base pairs.
-Figure \@ref(fig:SicydiumTree) depicts the ML tree (-lnL = `r loglikelihood`) showing the relationships among all COI sequences identified as *Sicydium* spp.
-available in GenBank.
-Six clades matching known *Sicydium* species were recovered with strong support.
-The phylogenetic reconstruction shows that the six samples from Cuba, highlighted in green, nested within *Sicydium buscki* Evermann & Clark, 1906.
+The final alignment included 212 sequences, each one of 639 base pairs.
+Figure \@ref(fig:SicydiumTree depicts the ML tree (-lnL = `r loglikelihood`) showing the relationships among all COI sequences identified as *Sicydium* spp.
+available from GenBank.
+Six clades corresponding to known *Sicydium* species were recovered with moderate to strong support.
+This phylogenetic reconstruction shows that the nine samples from Cuba, highlighted in green, nested within *Sicydium buscki* Evermann & Clark, 1906.
+Six haplotypes were identified from the Cuban samples, which showed high haplotype diversity (`r haploDiv[1]` ± `r haploDiv[2]`) and low nucleotide diversity (`r nucleDiv[1]` ± `r nucleDiv[2]`).
 
-```{r SicydiumTree, fig.width = 15, fig.height = 19, fig.cap=paste("ML phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*.
-Sequences highlited in green represent samples from Cuba.
-Numbers on branches represent boostrap values higher than 82.")}
+```{r SicydiumTree, fig.width = 15, fig.height = 19, fig.cap=paste("Maximum likelihood phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among partial COI sequences of the genus *Sicydium*.
+Sequences highlighted in green represent samples from Cuba.
+Numbers on branches represent bootstrap values higher than 82.")}
 
@@
   geom_treescale(x=0, y=-1.5, offset = -4, fontsize = 6, linesize = 0.8) + 
-  geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 82), size = 6, col= "black", vjust = -0.3, hjust=1.9, na.rm = FALSE) + 
+  geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 82), size = 4, col= "black", vjust = -0.3, hjust=2.3, na.rm = FALSE) + 
   geom_nodepoint(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 82), size = 2, col= "black") + 
-  geom_cladelabel(node=215, label="Sicydium plumieri", color="red", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") + 
-  geom_cladelabel(node=108, label="Sicydium brevifile", color="blue", hjust = -0.008, vjust = 0.8, fontsize = 6, offset = 0.002, fontface="italic") + 
-  geom_cladelabel(node=408, label="Sicydium salvini", color="purple", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") + 
-  geom_cladelabel(node=320, label="Sicydium punctatum", color="brown", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") + 
-  geom_cladelabel(node=371, label="Sicydium buscki", color="black", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") + 
-  geom_cladelabel(node=407, label="Sicydium crenilabrum", color="orange", hjust = -0.03, vjust = 0.1, fontsize = 6, offset = 0.004, fontface="italic") +
-  geom_cladelabel(node=211, label="Awaous spp.", color="darkgreen", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") +
+  geom_cladelabel(node=218, label="Sicydium plumieri", color="red", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=212, label="Sicydium brevifile", color="blue", hjust = -0.3, vjust = 0.08, fontsize = 6, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=320, label="Sicydium salvini", color="purple", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=330, label="Sicydium punctatum", color="brown", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=381, label="Sicydium buscki", color="black", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") + 
+  geom_cladelabel(node=328, label="Sicydium crenilabrum", color="orange", hjust = -0.001, vjust = 0.2, fontsize = 6, offset = 0.002, fontface="italic") +
+  geom_cladelabel(node=214, label="Awaous spp.", color="darkgreen", hjust = -0.07, fontsize = 6, offset = 0.002, fontface="italic") +
   # samples Sicydium_COI_Cuba2,4,6 were nested within this clde but the distance is too short to be displayed in the tree
@@
   theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm"))
+rotate(tree_view = NULL, 329)
+# suppressWarnings(suppressMessages(print(Sicydium_ML_tree)))
 
-# suppressWarnings(suppressMessages(print(Sicydium_ML_tree)))
+### Code to start working on Table 1
+ # Final version imported (and edited) as an external file (excel)
+
+### sicydiumTable <- setNames(data.frame(do.call(rbind, strsplit(as.character(nameTable$names.nameTable.), split="_"))), c("genus", "species", "gene", "accession number"))
+# nameTable: data.frame with the names of the sequences (all 212 seqs) got from the final fasta file (#7)
+# names.nameTable.: name of the only column of that df
 
@@
 
-When it comes to nature conservation, species identification is a basic principle: what cannot be identified, will not be protected or conserved.
-Any successful management plan or effort will critically rely on a correct identification of the taxa intended to protect.
-However, when taxonomic issues exist, molecular tools have proved to be of considerable help uncovering largely hidden diversity [*e.g.*, @ardura2010dna; @kress2015dna].
-Cuban freshwater fishes are an excellent example of the importance of applying molecular tools to species conservation, where more than ten new lineages have been discovered in the last years [@lara2010dna; @garcia2011molecular; @de2014phylogeography].
-Nevertheless, some groups remain poorly studied, which prevents us from implementing appropriate conservation programs.
+In terms of nature conservation, species identification is a basic starting point: what cannot be identified cannot be protected or conserved.
+Any successful management plan or effort will rely on a correct identification of the taxon needing protection.
+However, when taxonomic issues exist, molecular tools have proved to be of considerable help uncovering hidden diversity [*e.g.*, @ardura2010dna; @kress2015dna].
+Cuban freshwater fishes are an excellent example of the importance of applying molecular tools to species conservation: more than ten new lineages across three families have been discovered in the last ten years [@lara2010dna; @garcia2011molecular; @de2014phylogeography].
+Nevertheless, some groups remain poorly studied, which has prevented the implementation of appropriate conservation programs.
 
-Freshwater gobies (Gobiiformes, Gobiidae) have been a neglected group of fishes in Cuba.
-In fact, most information about their diversity and ecology on the island is virtually unknown.
-What we do know is that given their migratory life cycle, they have been the target of an artisanal fishery based on the postlarvae returning to rivers.
-This fish fry fishery, known as “Teti” fishery, is only practiced in the easternmost part of the island, and it has been reported from four rivers: Duaba, Yumuri, Toa, and Miel.
-The main problem with "Tetis", as it has been mentioned in many other countries harvesting postlarvae, is that species (and genus)-level identification is almost impossible in the field.
+Freshwater gobies (Gobiiformes: Gobiidae) have been a neglected group of fishes in Cuba and most information about their diversity and ecology on the island is virtually unknown.
+What we do know is that given their migratory life cycle, gobies have been the target of an artisanal fishery based on postlarvae returning to rivers on the easternmost part of the island.
+Our results suggest that Busck’s Goby, *Sicydium buscki*, is one of the species included in the “Tetí” fishery.
+*Sicydium buscki* is one of the four species of amphidromous goby reported in the Greater Antilles (@watson2000; @engmanetal2019; @eschmeyer2020catalog) and it may be that these other species (*S.
+gilberti*, *S.
+plumieri*, *S.
+punctatum*) are part of the “Tetí” fishery but without evidence supporting an alternative we must consider *S.
+buscki* as the primary candidate for the identity of this fishery.
+Information provided by local fishermen suggests that they harvest the fry mainly from four rivers: the Duaba, Yumurí, Toa, and Miel, during the last-quarter moon phase from August to October.
 
-Our results suggest that the “Teti” samples collected are *Sicydium buscki*.
-This is one of the four *Sicydium* species reported in the Caribbean islands [@eschmeyer2020catalog].
-Although other species have been suggested to comprise this fishery in Cuba, this is the first report of any fish species based on molecular data.
-However, this study included a few samples from a single river.
-Other countries where fish fry fisheries also take place have reported several species of the order Gobiiformes [@castellanos2011goby; @asis2013first; @engman2019puerto].
-Thus, additional surveys including more samples and rivers are needed to have a better idea, first of the diversity of gobies in Eastern Cuba, and second, of the fish diversity comprising the fish fry “Teti” fishery.
-Although harvest mainly occurs during the last-quarter moon phase from August to October [@tetiblogger2011], an increase in tourism in Cuba is leading to a higher demand for this natural resource.
-For this reason, further studies will contribute to our knowledge of Caribbean fish assemblages, which will allow us to better work toward their conservation and propose and develop effective management actions.
-
-# Conclusion
-
-Busck´s Stone-Biting Goby, *Sicydium buscki*, is a component of the fish fry “teti” fishery in Eastern Cuba.
-This species represents the only taxa of the orden Gobiiformes confirmed by molecular tools as part of such a fishery.
-More thorough investigations and identification of the species comprising the “teti” fishery are needed.
-It is imperative to include them in  local management plans and to avoid overfishing this resource.
+Although other species have been mentioned as potentially being part of the “Tetí” fishery, this is the first report of any fish species being identified based on molecular data.
+Other countries where fish-fry fisheries also take place have reported more than one species of goby (Gobiidae) and/or sleeper making up the fishery (Eleotridae) [@castellanos2011goby; @asis2013first; @engman2019puerto].
+Thus, additional surveys including more samples from different localities are needed to uncover the diversity of gobies in Eastern Cuba and the fish diversity comprising this particular fishery.
+An increase in tourism in Cuba, possibly leading to a higher demand for this natural resource, is a relatively new threat faced by the native fishes.
+They are also threatened by the introduction of exotic species, habitat fragmentation, and water pollution [@monitoreoCuba].
+All this makes the present study a valuable tool providing basic information about the fish fauna of the region.
+More thorough investigations and identification of species comprising the “Tetí” fishery will allow us to better work toward their conservation and propose effective management actions.
 
@@
 
-We thank the Habitat production team for their help during the field work.
-Sequencing was partially supported by a Smithsonian Tropical Research Institution Short-term Fellowship to SRM.
-In this sense, we thank advisement provided by Harilaos Lessios and the invaluable lab assistance of Eyda Gómez, Ligia and Axel Calderón.
-Additional lab work was also done with advice of Diego Elías.
-R code was developed with help of Tad Dallas, Heru Handika, and the community of Stack Overflow.
+We thank the Habitat Production Team for their help during fieldwork.
+Sequencing was supported by a Smithsonian Tropical Research Institute (STRI) Short-Term Fellowship to SRM.
+We are thankful for the advice and support provided by Harilaos Lessios (STRI) and the valuable lab assistance of Eyda Gómez, Ligia Calderón, and Axel Calderón (STRI).
+We would also like to thank two anonymous reviewers who made useful suggestions that improved the manuscript.
+R code was developed with the help of Tad Dallas, Heru Handika, and the community at http://stackoverflow.com.
 
@@
 if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())
-```

### 2021-09-07 analysis/paper/paper.Rmd
updatingMS&Ref

@@
 Then, we recovered the phylogenetic position of these samples based on a maximum likelihood phylogenetic analysis that also used available sequences from GenBank.
-Here, we provide evidence that Busck´s Stone-Biting Goby, Sicydium buscki, is a component of the “Tetí” fishery in Cuba.
+Here, we provide evidence that Busck´s Stone-Biting Goby, *Sicydium buscki*, is a component of the “Tetí” fishery in Cuba.
 However, whether this fishery includes one or more species remains a mystery until additional samples can be analyzed.
@@
 Raw sequence files were edited and aligned with the software Geneious v.9.1 [@kearse2012geneious] and submitted to GenBank (accession numbers: MZ848370– MZ848378).
-We used a standard nucleotide BLAST (BLASTN 2.11.0; @zhangetal2000), optimized for highly similar sequences (MegaBLAST; @morgulis2008database) (Query IDs: 1237–1245), and recovered the identity of the Cuban sequences with 99–100% identity.
+We used a standard nucleotide BLAST (BLASTN 2.11.0; @zhang2000greedy), optimized for highly similar sequences (MegaBLAST; @morgulis2008database) (Query IDs: 1237–1245), and recovered the identity of the Cuban sequences with 99–100% identity.
 
@@
 An increase in tourism in Cuba, possibly leading to a higher demand for this natural resource, is a relatively new threat faced by the native fishes.
-They are also threatened by the introduction of exotic species, habitat fragmentation, and water pollution [@monitoreoCuba].
+They are also threatened by the introduction of exotic species, habitat fragmentation, and water pollution [@monitoreoCuba2017peces].
 All this makes the present study a valuable tool providing basic information about the fish fauna of the region.

### 2021-09-08 analysis/paper/paper.Rmd
uploadingAppendix

@@
              "rentrez", "ape", "seqinr", "phangorn",
-             "adegenet", "pegas")
+             "adegenet", "pegas",
+             "knitr")
 package.check <- lapply(packages, FUN = function(x) {
@@
 Their small size along with cryptic body colors and the paucity of clear diagnostic characters to distinguish among taxa is a substantial challenge.
-However, the use of molecular tools applied to species-level identification has shed light on the taxonomy of these fishes [@keith2011phylogeny; @thacker2011phylogeny; @victor2014three; @kuang2018phylogenomic] and in the last ten years, more than one hundred gobiid species have been described [@eschmeyer2020catalog].
+However, the use of molecular tools applied to species-level identification has shed light on the taxonomy of these fishes [@thacker2011phylogeny; @victor2014three; @keith2017schismatogobius; @kuang2018phylogenomic] and in the last ten years, more than one hundred gobiid species have been described [@eschmeyer2020catalog].
 
@@
 This migratory behavior supports artisanal fisheries worldwide [@bell1999overview; @mcdowall2007amphidromy].
-Goby-fry fisheries are common practice in several countries in the Indo-Pacific [@asis2013first], the Pacific [@castellanos2011goby], and the Caribbean (@erdman1961notes; @kwaketal2016).
+Goby-fry fisheries are common practice in several countries in the Indo-Pacific [@asis2013first], the Pacific [@castellanos2011goby], and the Caribbean [@erdman1961notes; @kwaketal2016].
 One problem with this style of fishery is that insufficient management has led to a considerable reduction in catch [@bell1999overview].
@@
 The fish-fry “Tetí” fishery in Eastern Cuba represents a local artisanal fishery and there is no scientific information about the exploitation of this natural resource outside of the region.
-Based on other reports from the Caribbean (*e.g.*, @watson2000), some authors have claimed that the gobies *Sicydium plumieri* and *Sicydium punctatum* make up the “Tetí” fishery in Cuba, but evidence for those identications are lacking.
+Based on other reports from the Caribbean [*e.g.*, @watson2000], some authors have claimed that the gobies *Sicydium plumieri* and *Sicydium punctatum* make up the “Tetí” fishery in Cuba, but evidence for those identications are lacking.
 “Tetí” has been historically exploited as the main ingredient of “traditional and exotic” dishes [@pichardoytapia1862; A.
 Hartman, pers.
-comm., August 01, 2011).
+comm., August 01, 2011].
 However, recent changes to Cuban economic policies have promoted considerable growth in tourism [@rutty2019tourism], leading to increased fisheries pressure.
@@
 ```{r tetiPhoto, fig.cap="Lateral view of goby fry “Tetí” captured in Río Yumurí, Cuba. Photo by Tomás M. Rodríguez-Cabrera."}
+knitr::include_graphics("analysis/figures/Fig1_Rodriguez-Machado2.jpg")
 
@@
 Raw sequence files were edited and aligned with the software Geneious v.9.1 [@kearse2012geneious] and submitted to GenBank (accession numbers: MZ848370– MZ848378).
-We used a standard nucleotide BLAST (BLASTN 2.11.0; @zhang2000greedy), optimized for highly similar sequences (MegaBLAST; @morgulis2008database) (Query IDs: 1237–1245), and recovered the identity of the Cuban sequences with 99–100% identity.
+We used a standard nucleotide BLAST [BLASTN 2.11.0; @zhang2000greedy], optimized for highly similar sequences [MegaBLAST; @morgulis2008database] (Query IDs: 1237–1245), and recovered the identity of the Cuban sequences with 99–100% identity.
 
@@
 Duplicated sequences and sequences with missing data were excluded.
-The dataset was edited and aligned using the R packages *seqinr* [@charif2007seqinr], *ape* [@paradis2015package], and *Biostrings* [@pages2013package].
+The dataset was edited and aligned using the R packages *seqinr* [@charif2007seqinr], *ape* [@paradis2019ape], and *Biostrings* [@pages2013package].
 The final alignment was obtained with the algorithm ClustalOmega [@sievers2011fast] using the R package *msa* [@bodenhofer2015msa].
@@
 Our results suggest that Busck’s Goby, *Sicydium buscki*, is one of the species included in the “Tetí” fishery.
-*Sicydium buscki* is one of the four species of amphidromous goby reported in the Greater Antilles (@watson2000; @engmanetal2019; @eschmeyer2020catalog) and it may be that these other species (*S.
+*Sicydium buscki* is one of the four species of amphidromous goby reported in the Greater Antilles [@watson2000; @engman2019puerto; @eschmeyer2020catalog] and it may be that these other species (*S.
 gilberti*, *S.
@@
 
+## Appendix
+
+```{r appendix, echo = FALSE}
+
+# This sections requires packages here and knitr [library(here), library(knitr)]
+
+appendixCSV <- read.csv(here("/analysis/data/derived_data/Appendix_Rodriguez-Machado_et_al.csv"))
+
+appendixDF <- data.frame(lapply(appendixCSV, function(x) {gsub("_", " ", x)}))
+
+knitr::kable(appendixDF, caption = "Appendix 1.
+Species, localities, and accession number information of the COI sequences included in this study.
+All sequences were retrieved from GenBank.
+References were included when locality information is not available on GenBank.
+N: total number of sequences per species.")
+
+```
+
+<!-- The following line inserts a page break  -->
+
+\newpage
+
 ### Colophon
